### 2009-07-17 SC09/abstract.tex
Adding the SC paper in its source form.
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@2 979fb63a-54b4-4514-8d67-9ab8ea1b59df

\vspace{-.5cm}
\abstract{
\scribble{Rewordings. Eliminated last paragraph. Changed definition of \\upfs to use 'file system' instead of 'file system'. Changed National Science Foundation site to PSC}
\em{
Parallel applications running across thousands of processors must protect
themselves from inevitable component failures. Many applications 
insulate themselves from failures by checkpointing, a
process in which they save their state to persistent storage. Following a
failure, they can resume computation using this state.
For many applications, saving this state into a shared single file
is most convenient.  With such an approach, the size of writes are often 
small and not aligned with file system boundaries. Unfortunately for these
applications, this preferred data layout results in pathologically poor
performance from the underlying file system which is optimized for
large, aligned writes to non-shared files.
\if 0
% Milo's changes which I didn't incorporate
process in which they periodically save their state to persistent storage. Following a
failure, they can retrieve the most recently saved state and resume computation from that
point. For many applications, saving this state into a shared single file, often
formatted for storage by complex libraries such as NetCDF or HDF5,
is most convenient.  With such an approach, the size of individual write system calls are often 
small and not aligned with file system boundaries especially when the application cannot statically define the order of state variables, as with adaptive mesh refinement app. Unfortunately for these
applications, this write access pattern can result in pathologically poor
performance from an underlying file system, itself often optimized for
\fi

To address this fundamental mismatch, we have developed a parallel
log-structured file system, \plfs, which is positioned between the applications
and the \upfs. \plfs\ remaps an application's write access pattern to be
optimized for the underlying file system. Through testing on Panasas
ActiveScale Storage System and IBM's General Parallel File System at Los Alamos
National Lab and on Lustre at Pittsburgh Supercomputer Center, we have seen
that this layer of indirection and reorganization can reduce checkpoint time by
up to several orders of magnitude for several important benchmarks and real
applications.

We expect that \plfs\ can improve the checkpoint bandwidth for any large
parallel application that writes to a single file. The expected improvement is
especially large for those applications doing unaligned or random IO, patterns
which have become increasingly prevalent recently due to the wide-spread
adoption of complex formatting libraries such as NetCDF and HDF5.

\if 0
% here's the 150 word limited version we submitted to SC
Parallel applications running across thousands of processors must protect themselves from inevitable system failures. Many applications insulate themselves from failures by checkpointing. For many applications, checkpointing into a shared single file is most convenient. With such an approach, the size of writes are often small and not aligned with file system boundaries. Unfortunately for these applications, this preferred data layout results in pathologically poor performance from the underlying file system which is optimized for large, aligned writes to non-shared files. To address this fundamental mismatch, we have developed a virtual parallel log structured file system, PLFS. PLFS remaps an application’s preferred data layout into one which is optimized for the underlying file system. Through testing on PanFS, Lustre, and GPFS, we have seen that this layer of indirection and reorganization can reduce checkpoint time by an order of magnitude for several important benchmarks and real applications.
\fi
}
}

### 2009-07-17 SC09/arch-fig.tex
Adding the SC paper in its source form.
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@2 979fb63a-54b4-4514-8d67-9ab8ea1b59df

\if 0
% here's a way to try to make it just be one column wide but it didn't
% work for me, it made it one columnwide but it moved it to the end of the
%document
\begin{wrapfigure}{r}{0.5\textwidth}[tb]
    \begin{center}
        \psfig{file=figs/architecture.eps,width=0.48\textwidth}
    \end{center}
% we can also try to use SCfigure for captions on the side, I tried that
% and it looks sort of weird.  But I'm also not happy with the default
% which makes it way too big and leaves too much white space..
\begin{SCfigure}
    \centering
    \includegraphics[width=0.5\textwidth]{figs/architecture.eps}
\fi

\singlespace
\begin{figure}[tb]
    \centering
    \includegraphics[width=0.7\textwidth]{figs/architecture.eps}
    \mycaption{fig-arch}{Data Reorganization.}{
This figure depicts how PLFS reorganizes an N-1 strided checkpoint file onto
the \upfs.  A parallel application consisting of six
processes on three compute nodes is represented by the top three boxes.  Each 
box represents a compute node, a circle is a process, and the three small boxes
below each process represent the state of that process. The processes
create a new file on \plfs\ called \textit{checkpoint1}, causing  \plfs\ in
turn to
create a \Term{container} structure on the underlying parallel file system.
The container consists of a top-level directory also called
\textit{checkpoint1} and several
sub-directories to store the application's data.  For each
process opening the file, \plfs\ creates a \Term{data file} within one of the
sub-directories, it also creates one \Term{index file} within that same 
sub-directory 
which is shared by all processes on a compute node.
For each write, \plfs\ appends the data to the corresponding data file 
and appends a record into the appropriate index file.  This record
contains the length of the write, its logical offset, and a pointer
to its physical offset within the data file to which it was appended.
To satisfy reads, \plfs\ aggregates these index files to create a lookup
table for the logical file.  Also shown in this figure are the
\textit{access} file, which is used to store ownership and privilege
information about the logical file, and the \textit{openhosts} and
\textit{metadata} sub-directories which are used to cache metadata in order
to improve query time (\eg\ a \syscall{stat} call). 
}
\end{figure}
\doublespace

### 2009-07-17 SC09/arch.tex
Adding the SC paper in its source form.
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@2 979fb63a-54b4-4514-8d67-9ab8ea1b59df

\section{Design of an Interposition Layer}
\label{arch}

\input{arch-fig}

We start with the hypothesis that an interposition layer can transparently
rearrange an N-1 checkpoint pattern into an N-N pattern and thereby decrease
checkpoint time by taking advantage of the increased bandwidth achievable via
an N-N pattern. To test this hypothesis, we have developed one such
interposition layer, \plfs, designed specifically for large parallel N-1
checkpoint files. The basic architecture is illustrated in
Figure~\ref{fig-arch}. \plfs\ was prototyped with FUSE~\cite{fuse}, a 
framework for running stackable file systems~\cite{usenix00fist} in 
non-privileged mode.

\plfs\ is a virtual file system situated between the parallel application and
an \upfs\ responsible for the actual data storage. As \plfs\ is a virtual file
system, it leverages many of the services provided by the \upfs\ such as
redundancy, high availability, and a globally distributed data store.  This
frees \plfs\ to focus on just one specialized task: rearranging application
data so the N-1 write pattern is better suited for the \upfs.  In the remainder
of this paper,  we refer to \plfs\ generally to mean this virtual file system
which itself is comprised of a set of \plfs\ servers running across a compute
system bound together by an \upfs; when we refer to a {\em specific} \plfs\ we
mean just one of these servers.

\scribble{Cut last sentence --Milo}

% Although \plfs\ looks and feels like a
%full-fledged parallel file system, in actuality it is merely a collection of
%independent virtual file systems running on each compute node that appear
%parallel only because they exploit the parallelism of the underlying
%file system.

%As multiple processes in a parallel application open an N-1 checkpoint file,

\subsection{Basic operation}
The basic operation of \plfs\ is as follows. For every logical \plfs\ file
created, \plfs\ creates a \Term{container} structure on the \upfs. Internally,
the basic structure of a container is a hierarchical directory tree consisting
of a single top-level directory and multiple sub-directories that appears to
users; \plfs\ builds a logical view of a single file from this container
structure in a manner similar to the core idea of Apple
bundles~\cite{bundles} in Mac OS X.  Multiple processes opening the same
logical file for writing share the container although each open gets a unique
\Term{data file} within the container into which all of its writes are
appended. By giving each writing process in a parallel application access to a
non-shared data file, \plfs\ converts an N-1 write access pattern into a N-N
write access pattern.  When the process writes to the file, the write is
appended to its data file and a record identifying the write is appended to an
index file (described below).

\scribble{Added words, Mac OS X citation}

\subsubsection{Reading from \plfs}
\label{arch-read}
Rearranging data in this manner should improve write bandwidths, but
it also introduces additional complexity for reads. In order to read 
the logical file, \plfs\ maintains an \Term{index file} for each compute
node which records the logical offset and length of each write. \plfs\ can
then construct a \Term{global index} by aggregating the multiple index files 
into an offset lookup table.

One difficultly in constructing this global index stems from concurrent
processes that may write the same offset at the same time.  These processes
cannot know which will be the ultimate writer, so they need to synchronize to
determine the appropriate order. This synchronization needs to be exposed to
the file system to be effective~\cite{Gibson95thescotch}. If a logical
clock~\cite{lamport1978} was associated with these synchronizations, its value
could be written to index files so that the merge process could correctly
determine write ordering. Since parallel applications synchronize with a
barrier called by all processes, a simple count of synchronization calls could
be sufficient. In practice checkpoints do not experience overlapping writes, so
at this time \plfs\ has not implemented an overlap conflict resolution scheme.

\scribble{changed language on 'serialization of index writes'. We actually only
serialize the index offset updates --Milo}

One interesting nuance is that \plfs\ has a data file for every process but
only a single index file per compute node shared by all processes on that node.
Sharing an index file is easy; by the time \plfs\ sees writes, they have been
merged by the operating system into a single memory space. The operating
systems sends all writes to a single \plfs\ process which ensures index records
are correctly, chronologically appended.  Having a single index greatly reduces
the number of files in a container since current LANL applications run up to
sixteen processes on a node; on the next LANL supercomputer, this could be up
to sixty-four. We tried reducing the number of data files in the same manner.
Write bandwidth was not affected, but reads were slowed for \Term{uniform
restarts} in which reading processes access the file in the same access pattern
as it was written.  The pattern of a single reader accessing a single data file
sequentially lends itself very well to prefetching.  However, due to timing
differences between the write and read phases, multiple processes in a uniform
restart may not always read from a shared file in sequential order.

Having described the basic operation of \plfs, we now present some of its
implementation in finer detail.  Although there are many interposition
techniques available (~\cite{bypass} includes a survey), we have selected FUSE
for several reasons. Because FUSE allows \plfs\ to be accessible via a standard
file system interface, applications can use it without modification and files on
\plfs\ can be accessed by the entire suite of existing tools such as
\textit{ls}, \textit{diff}, and \textit{cp}. In addition to providing user
transparency, using FUSE dramatically simplified our development effort. A file
system in userspace is significantly easier to develop than a kernel file
system and is more portable as well. However, this convenience is not free as
FUSE does add some overhead as shown in Section~\ref{eval}.

\subsubsection{Container implementation}

\scribble{I think this part was superfluous so I cut the LD\_PRELOAD stuff}
%Finally, the simplified interface
%of the FUSE API means developers need only implement a small subset of
%function calls as opposed to the much larger set that would be required if
%we had chosen to interpose using an \textit{LD\_PRELOAD} library. As will
%be shown however in Section~\ref{eval}, this convenience is not free as FUSE
%does add some small amount of overhead.

Because \plfs\ leverages the \upfs\ as much as possible, we give the container
the same logical name as the \plfs\ file. This allows \plfs\ to pass a
\syscall{readdir} system call directly to the \upfs\ and return its result to
the application without any translation.  \plfs\ also handles \syscall{mkdir}
system calls in the same way (\ie\ without any translation). The implication of
leveraging the \upfs\ in this way is that \plfs\ requires some other mechanism
by which to distinguish between regular directories and containers in order to
implement the \syscall{stat} system call correctly. As the \textit{SUID} bit is
rarely used on directories and yet is allowed to be set by the underlying file
system, \plfs\ sets this bit on containers; this does mean, however, that
\plfs\ must disallow setting this bit on a regular directory.

As we discussed previously, parallel applications do synchronized
checkpointing; the implication for \plfs\ is that multiple processes running on
multiple compute nodes writing an N-1 checkpoint file will cause
\plfs\ on each compute node to attempt to create the same container
concurrently on the \upfs. The difficulty arises because each \plfs\ must first
\syscall{stat} that path to see whether the path is available, whether that
container already exists, or whether there is a regular directory at that
location. Ideally, each \plfs\ could \syscall{stat} the path and, when the
location is empty, atomically create a
directory with the \textit{SUID} bit set. Unfortunately, the \syscall{mkdir}
system call ignores the \textit{SUID} bit; each \plfs\ must therefore first
create a directory and then set the \textit{SUID} bit. Doing this
naively results in a race condition: if one \plfs\ \syscall{stats} the path
after another has made the directory but before it has set the \textit{SUID}
bit, then the \syscall{stat} will indicate that there is a regular directory in
that location. The application issuing the open of the logical file will then
receive an error incorrectly indicating that there is a directory already at
that location. To avoid this race condition, each \plfs\ first makes a
hidden directory with a unique name, set its \textit{SUID} bit, and then
atomically \syscall{renames} it to the original container name.

\subsection{Metadata operations}
\label{arch-meta}

Metadata operations against a file include accessing its permissions (including
SUID), its capacity, the offset of its last byte and the timestamp of its last
update. For a directory on \plfs, these are provided by the underlying
file system. But for a \plfs\ file which is constructed from a container like
the example in Figure~\ref{fig-arch}, these metadata operations have to be
computed from the many underlying files within the container.

Because the SUID bit on the container itself has been overloaded to indicate
that the directory is not a directory at all, but rather a container, it cannot
be also used to indicate if the user has set SUID on the \plfs\ file
represented by the container. Instead we use a file inside the container, the
\Term{access file}, to represent the appropriate SUID and the rest of the
permissions associated with the container. For example, where \syscall{chmod}
and \syscall{chgrp} are directed at the logical file, they are applied to the
access file within the container.

Capacity for the logical file is the sum of the capacities of the files inside
the container.  The last update timestamp is the maximum of the last update
timestamps.  And the offset of the last byte is the maximum logical offset
recorded in any of the index files.

Computing these sums and maximums with every \syscall{stat} call on a \plfs\
file is expensive. Our strategy for speeding this up is to cache recently
computed values in the \Term{metadata subdirectory}. To make this cache as
effective as possible we have each FUSE process cache into this metadata
subdirectory any information it has in its in memory data structures when the
last writer on that node closes the file. On this close, the
FUSE process creates a file named H.L.B.T, where H is the node's hostname, L is
the last offset recorded on that node, B is the sum of the capacity of all
files that this node manages, and T is the maximum timestamp among these files.

When no process has this container open for write, a stat call on the container
is implemented by issuing a \syscall{readdir} on the metadata subdirectory,
then reporting the maximum of the Ls for last byte offset, the maximum of the
Ts for modification timestamp and the sum of the Bs for capacity.

If one or more processes has the container open for writing, then the
corresponding cached metadata values could be stale. \plfs\ clients therefore
create a file in the \Term{openhosts subdirectory} named by its hostname when
one or more processes on that node have that file open for writing, and
then deleting this file once all \syscall{opens} have been closed.
\syscall{stat} must then do a \syscall{readdir} on openhosts as well to
discover if any node has the file open for writing, and thereby determine which
metadata cache entries might be stale.

When there are hostname files in the openhosts subdirectory, the node that is
executing the \syscall{stat} call could read the contents of the index files
associated with those hostnames in order to find the largest logical offset,
and then combine this with the metadata cache files that are not stale.

In the experience of the HPC community, \syscall{stat'ing} an open file is
almost always done by a user trying to monitor the progress of their job. What
they want is an inexpensive probe showing progress, not an expensive
instantaneously correct value~\cite{hec-posix}. Following this logic, \plfs\
does not read and proccess the index files associated with hostnames that have
the container open for write. Instead it assumes that files are not sparse
(\ie\ every byte has been written) and sums the sizes of all data files within
a container to estimate the last offset of the \plfs\ file.  Because writes to
each data are always simply appended, this estimation will monotonically
increase as additional data is written into the file, allowing users to monitor
progress. When the container is closed, the metadata subdirectory contains
fully correct cached values, and full accuracy is provided at all times when
the container has no processes writing it.

### 2009-07-17 SC09/conclude.tex
Adding the SC paper in its source form.
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@2 979fb63a-54b4-4514-8d67-9ab8ea1b59df

\section{Conclusion}
\label{conclude}

Large systems experience failures.  To protect themselves against these
failures, parallel applications running on these systems save their progress by
checkpointing.  Unfortunately for many of these applications, their preferred
checkpointing patterns are extremely challenging for the \upfs\ and impose
severe bandwidth limitations.  In this paper, we have developed \plfs\ to
demonstrate how a simple interposition layer can transparently rearrange these
challenging patterns and improve checkpoint bandwidth by several orders of
magnitude.

The parallel file system attached to Roadrunner is the largest LANL has ever
had; testing it has revealed that the challenges of N-1 patterns are severely
exacerbated at this scale.  Given current bandwidths, we know of no current N-1
application at LANL that can effectively checkpoint across the full width of
Roadrunner.  \plfs\ allows them to do so.

### 2009-07-17 SC09/eval-graphs.tex
Adding the SC paper in its source form.
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@2 979fb63a-54b4-4514-8d67-9ab8ea1b59df

% eval figures all in one big table
\begin{figure}[t!]
\centering
\begin{tabular}{ccc}
    \tablefigure{MPI-IO Test on PanFS}{panfs}
    &
    \tablefigure{MPI-IO Test on GPFS}{gpfs}
    &
    \tablefigure{MPI-IO Test on Lustre}{lustre}
    \\
    \tablefigure{LANL Anonymous 1}{lanl1}
    &
    \tablefigure{LANL Anonymous 2}{lanl2}
    &
    \tablefigure{LANL Anonymous 3}{lanl3}
    \\
    \tablefigure{PatternIO}{pattern}
    &
    \tablefigure{QCD}{qcd}
    &
    \tablefigure{BTIO}{btio}
    \\
    \tablefigure{FLASH IO}{flash}
    &
    \tablefigure{SUMMARY}{summary}
    &
    \tablefigure{Chombo IO}{chombo}
\end{tabular}
\mycaption{fig-eval}{Experimental Results.}{
The three graphs in the top row are the same graphs that were presented
earlier in Figure~\ref{fig-motivation}, except now they have an additional
line showing how PLFS allows an N-1 checkpoint to achieve most, if not all,
of the bandwidth available to an N-N checkpoint. 
The bar graph in the center of the bottom row consolidates these results and
shows a pair of bars for each, showing both the relative minimum and the
maximum speedups achieved across the set of experiments.  Due to radically
different configurations for these various experiments, the axes for 
these graphs are not consistent.  The relative comparison within each graph
should be obvious; absolute values can be ascertained by reading the axes.
}
\end{figure}

### 2009-07-17 SC09/eval.tex
Adding the SC paper in its source form.
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@2 979fb63a-54b4-4514-8d67-9ab8ea1b59df

\section{Evaluation}
\label{eval}

\input{eval-graphs}

We present the results of our experimental evaluation in Figure~\ref{fig-eval}.
Eleven of these twelve graphs present one experiment each. The twelfth,
Figure~\ref{eval:summary}, presents a summary. In the majority of these graphs,
the write bandwidth is shown on the \yaxis\ in \MBs\ as a function of the
number of processes. We will note it in the text for those few graphs for
which we deviate from this general configuration. The write bandwidth that we
report is whatever is reported by the particular benchmark; whenever possible,
we report the most conservative value (\ie\ we include open and close times in
our write times, and we either barrier after close or we use the time reported
by the slowest writer). Finally we have attempted to run multiple iterations
for each experiment; where applicable, the standard deviation is therefore
included.

\subsection{MPI-IO Test}

The top three graphs, 
Figures~\ref{eval:panfs},~\ref{eval:gpfs},~and~\ref{eval:lustre}, present the
results of our study using the LANL synthetic checkpoint tool, \Term{MPI-IO
Test}~\cite{mpi-io-test}, on three different parallel file systems, PanFS, GPFS,
and Lustre. For each of these graphs, the size of each \syscall{write} was
47001 bytes (a small, unaligned number observed in actual applications to be
particularly problematic for file systems). \syscall{Writes} were
issued until two minutes had elapsed. Although this is atypical since 
applications tend to write a fixed amount of data instead of writing for
a fixed amount of time, we have observed that this allows representative 
bandwidth measurements with a predictable runtime.

There are several things to notice in these graphs. The first is that these
are the same three graphs that we presented in Figure~\ref{fig-motivation}
except that we have now added a third line to each. The three lines show the
bandwidth achieved by writing an N-N pattern directly to the \upfs, the
bandwidth achieved by writing an N-1 pattern directly to the \upfs, and the
third line is the bandwidth achieved by writing an N-1 pattern {\em indirectly}
to the \upfs\ through \plfs.

These graphs illustrate how the performance discrepancy between N-N
and N-1 checkpoint patterns is common across PanFS, GPFS,
and Lustre. Remember, as was discussed in Section~\ref{motivation}, switching
to N-N is not a viable option for many applications which are inextricably wed
to an N-1 pattern and are resigned to the attendant loss of bandwidth.
Fortunately, as is evidenced by these graphs, \plfs\ allows these applications
to retain their preferred N-1 pattern while achieving most, if not all, of the
bandwidth available to an N-N pattern. Particularly for the PanFS results,
which were run on our Roadrunner supercomputer, \plfs\ achieves the full
bandwidth of an N-N pattern (\ie\ up to about 31 \GBs). In fact, for several
of the points, an N-1 pattern on \plfs\ actually outperforms an N-N pattern
written directly to PanFS. Although we have yet to fully investigate the exact
reason for this, there are several reasons why this could be the case. The
first is that \plfs\ rearranges writes into a log structured pattern so an N-N
pattern which incurs seeks could do worse than a \plfs\ pattern which appends
only. Secondly, the structure of the \plfs\ container spreads data across
multiple sub-directories within a top-level directory whereas a typical N-N
pattern confines all N files into only a single parent directory.

Although \plfs\ does improve the bandwidth of N-1 patterns on GPFS and Lustre,
the improvement is not as large as it is on PanFS. This is because the scale
of the experiments on PanFS are 200 times larger than on the other two
platforms. In the inset in Figure~\ref{eval:panfs} at the extreme low values
for the number of processes, we see that \plfs\ does not scale N-1 bandwidth
as fast as N-N scales for PanFS as well. This is due to overheads incurred by
both FUSE and \plfs; these overheads limit the total bandwidth achieved by any
single compute node relative to N-N. For HPC systems at extreme scale this
limitation does not matter since aggregate bandwidth across multiple nodes is
more relevant than the bandwidth from a single node. In this case the real
limiting factor is the total bandwidth capacity of the storage fabric, which
generally saturates when using only a small fraction of compute nodes.
However, with poorly-behaved IO patterns (\ie\ N-1), even very large jobs may
not reach these bandwidth limits because they will be more severely
constrained by file system limitations as exemplified by
Figure~\ref{fig-motivation}. \plfs\ is designed to remove these file system
limitations, so that jobs can achieve higher bandwidth and reach the
same limitation of the storage fabric which is constraining N-N bandwidth.

An unusual feture of the inset graph in Figure~\ref{eval:panfs} is that there
are localized bottlenecks about 700 processes.  This is due to the
architecture of Roadrunner and it's connection to PanFS.  Roadrunner is split
into seventeen sub-clusters, \Term{CUs}, each of which can run 720 proceses.
The CUs are fully connected via an Infiniband fat tree, however, in order to
minimize network infrastructure costs, the storage system is partitioned into
multiple sub-nets. Each CU currently has six specialized IO nodes, one for
each sub-net; these six IO nodes are the storage bandwidth bottleneck.
Therefore, as a job grows within a CU, it quickly saturates the IO node
bandwidth; when it grows into another CU, its bandwidth increases sharply due
to gaining six additional IO nodes: this explains the "stair-step" behavior of
the N-N and \plfs\ lines in this graph.

We do see this same behavior in our GPFS and Lustre graphs, but due to the
very small size of our test systems, we do not have enough compute nodes to
saturate the available storage bandwidth and thus neither the N-N nor the
\plfs\ lines in these graphs reach a storage bandwidth bottleneck. We will
examine the FUSE and PLFS overheads in greater detail in
Section~\ref{overhead}.

\subsection{Real LANL Applications}

The second row in Figure~\ref{fig-eval}, consisting of
Figures~\ref{eval:lanl1},~\ref{eval:lanl2},~and~\ref{eval:lanl3}, shows the
results of using \plfs\ to improve the bandwidth of three important LANL
applications which use N-1 checkpoint patterns. The application
shown in Figure~\ref{eval:lanl1} is the application whose developers augmented
their checkpoint routine with a new checkpoint library called bulkio, which
aggregates and buffers writes into larger, contiguous writes more friendly to
the \upfs. Therefore we show four lines in this graph: one for the old
checkpoint method which uses MPI-IO routines to directly write data to PanFS, a
line for the bulkio method writing directly to PanFS, and a second line for
each of these techniques writing indirectly to PanFS via \plfs. Notice that
\plfs\ is not able to improve the bulkio technique which loses a significant
amount of bandwidth due to aggregating data and reducing parallelism by
reducing the number of writers. However, when the older MPI-IO technique is
used with \plfs, \plfs\ is able to exploit the additional parallelism possible
by using all of the processes as writers. This data was collected on a small,
Opteron-only, \rrz. The next two graphs show similar results; using \plfs\ to
rearrange an N-1 pattern yields significantly improved bandwidths.
Figure~\ref{eval:lanl2} was also run on the \rrz, whereas
Figure~\ref{eval:lanl3} was run on Roadrunner itself; this can be seen by the
much larger scale (on both the \yaxis\ and the \xaxis) for
Figure~\ref{eval:lanl3}. Note that hese are extremely important applications
at LANL; it has been estimated that they account for more than half of the
total computing cycles at LANL. Although these applications must remain
anonymous, traces of their IO are available~\cite{plfs-maps}.

\subsection{NERSC's PatternIO Benchmark}

Figure~\ref{eval:pattern} presents the data derived from our measurements taken
using NERSC's PatternIO benchmark~\cite{nersc} which plots write bandwidth as a
function of \syscall{write} size. Notice that this deviates from the other
graphs in Figure~\ref{fig-eval} which plot write bandwidth as a function of the
number of processes. For this experiment, run on the \rrz, the number of
processes was set at a constant 512. This graph also is a scatter plot instead
of using lines with standard deviations. The points for writing directly to
PanFS show three interesting slopes all converging at about 1 GB/s on the far
right side of the graph. The highest of these three regions shows the
performance of \syscall{writes} when the write size is both block aligned and
aligned on file system boundaries. The middle region is when the write size is
block aligned only and the lowest region is when the write size is not aligned
at all. {\em This graph demonstrates that \plfs\ allows an application to
achieve a consistently high level of write bandwidth regardless of the size, or
alignment, of its \syscall{writes}.}

\subsection{Other Benchmarks}

The remainder of the graphs show various other parallel IO benchmarks with an
N-1 checkpoint pattern. QCD~\cite{qcd-io} in Figure~\ref{eval:qcd} shows a
large improvement using \plfs\ but this improvement actually {\em degrades} as
the number of processes increases. This is because the amount of data written
by QCD is a small constant value and the overhead due to container creation
incurred in the \syscall{open} becomes proportionally greater as the amount of
data written by each process decreases. BTIO~\cite{btio-io} in
Figure~\ref{eval:btio} also shows a large improvement but we currently only
have one data point; we will have more in the final version.
FLASH~IO~\cite{flash-io} and Chombo~IO\cite{chombo-io}, shown respectively in
Figures~\ref{eval:flash}~and~\ref{eval:chombo}, both show improvement due to
\plfs\ which scales nicely with an increasing number of processes. FLASH was
run on Roadrunner whereas Chombo was run on the \rrz; both were built with the
HDF5 library~\cite{hdf5}.

\subsection{Summary}

A summary of the benchmark results, excluding the LANL and NERSC synthetics, is
presented in Figure~\ref{eval:summary}. This summary shows a pair of bars for
each benchmark; one for both the worst-case and best-case speedups for \plfs.
Only in one case did \plfs\ actually degrade performance: a slowdown of three
percent for LANL 1 when it was run on just a single node. Although we did test
on single nodes to be thorough, single node sized jobs are extremely rare in
HPC. The relevant results from reasonably sized jobs, showed at least a
doubling of performance for all studied benchmarks. QCD and BTIO experienced a
single order of magnitude improvement. The best result was observed for FLASH
which was improved in the best case by two orders of magnitude.

[Note to the reviewer: Only FLASH and LANL3 were run at large-scale using both
\plfs\ and the \upfs. We will have fuller results for the final experiment and
we expect at least two orders of magnitude improvement for all of them except
QCD.]

\subsection{Overhead}
\label{overhead}

\input{overhead-fig}

As was seen in
Figures~\ref{eval:panfs},~\ref{eval:gpfs},~and~\ref{eval:lustre}, N-1 patterns
written through \plfs\ only match the bandwidth achieved by N-N patterns
written directly to the \upfs\ once the storage system bandwidth is saturated.
For small number of processes, \plfs\ cannot match the performance of N-N
(nonetheless, it does still improve bandwidth over a direct N-1 pattern). This
overhead is measured in Figure~\ref{fig-overhead} which shows results as
measured on LANL's GPFS system. In order to measure the both overhead incurred
by FUSE as well as any additional overhead incurred by \plfs, we developed a
second FUSE file system, \Term{\noopfs}.  \noopfs\ does no extra work, merely
redirecting all IO to the \upfs (\ie\ GPFS). For those readers familar with
FUSE, please note that \noopfs\ caches the file descriptor created in the
\syscall{open} into the opaque FUSE file handle pointer, uses it for subsequent
\syscall{writes} and \syscall{reads}, and closes it in the \syscall{flush}.

Figure~\ref{fig-overhead} is almost the same as Figure~\ref{eval:gpfs} which
compares the bandwidths measured for N-N directly to GPFS, N-1 directly to
GPFS, and N-1 indirectly to GPFS written through \plfs. The difference here is
that several new measurements have been added. The first is for N-N written
indirectly to GPFS through \noopfs. This line is significantly lower than the
line for N-N written directly to GPFS; the delta between these lines is the
overhead incurred due to FUSE, which is approximately 20\%. The next
measurement added is running an N-N workload through \plfs; this shows the
additional overhead incurred by \plfs, approximately 10\%. 
The delta beween that line and the existing N-1 through \plfs\
measurements show additional overhead which we believe is due to serializations
within FUSE due to multiple processes accessing the same path within FUSE's
file table (approximately another 10\% loss). For completeness, we also
measured the bandwidth achieved by an N-1 workload written through \noopfs;
therefore, the bottom two lines on the graph provide another view of the
overhead lost to FUSE.

Although this graph shows a total overhead cost of about 40 to 50\%, this loss
of potential bandwidth is not unduly concerning. As was seen in
Figure~\ref{eval:panfs}, the loss of potential bandwidth due to this overhead
disappears for reasonable HPC processor counts. Even with a limited per-node
bandwidth, a relatively small number of nodes writing through \plfs\ is able to
saturate the storage bandwidth.

\subsection{Beyond Writing}
Although \plfs\ is designed primarily for {\em writing} checkpoints,
checkpoint files are still occassionally read.
We must therefore weigh improvements in 
write bandwidth against possible degradations in other operations.

\subsubsection{Read Bandwidth}
\input{read-graphs.tex}

To ensure \plfs\ does not improve write bandwidth at the expense of read
bandwidth, we ran a set of read experiments on \rrz\ which are shown in
Figure~\ref{fig-reading}. We first created two sets of 20~\GB\ files, each
written by a different number of writers; all writes were 47001 bytes (in other
words, increasing numbers of writers issued decreasing numbers of writes). One
set was created directly on PanFS; the other indirectly through \plfs. In all
cases, reads were performed on different nodes than those where the
corresponding data was written and caches were always flushed between
successive reads.

We measured the time to read these files using three different read
scenarios. One emulates a \Term{uniform restart} in which the number of
processes resuming the application is the same as the number that wrote the
last checkpoint. In this case, each rank within the parallel job reads the
same data in the same order as the previous writer of that same rank. In
contrast, a \Term{non-uniform restart} is one in which the number of
readers is different from the number of writers and the read offsets are
therefore not aligned with the previous writes. In this case, we emulate the
scenario where an application resumes after a failure by simply running on the
newly reduced number of nodes. The number of reads is not affected by the size
of the job for typical N-1 checkpoints. Each read is extracting a region for a
particular variable within the simulation. The size of this region depends on
the number of processes within the job. Each gets $1/Nth$ of the region.
Specifically, for a non-uniform restart in which the number of processes, $N$,
has been reduced by $M$, the size of each read will be $N/(N-M)$ times the size
of the original writes. The third read scenario we emulated is a typical
scenario in which a relatively, small fixed number of processes reads a
checkpoint in order to save it onto an archive system. For these measurements,
we used LANL's archive copy utility with fours readers each running on their
own node; each reader read just one contiguous 5~\GB\ region by issuing 
sequential 1~\MB\ reads.

\if 0
% tempting to say this but we said at the beginning that we wouldn't really
% talk about segmented anymore
The archive copy uses an N-1 segmented pattern to
read a checkpoint that was written with N-1 strided.
\fi

The results of these experiments can be seen in Figure~\ref{fig-reading}. In
order to allow comparison across the three experiments, the ranges of the axes
have been made consistent. The \yaxis\ shows read bandwidth as a function of
the number of writers who created the checkpoint. We can easily see from these
graphs that the highest bandwidth is achieved using \plfs\ in the uniform
restart scenario. This is not surprising: each reader moves sequentially
through just one, nonshared file, a pattern easily improved through prefetching
performed by the \upfs. The bandwidth decreases here due to the decreasing
amount of data read as the number of readers increases. With less data read by
each, the open times begin to dominate, and there is less potential for
prefetching. The very low bandwith observed when the checkpoint is stored
directly on PanFS is also not surprising due to its layout of data within RAID
groups. Each contiguous \GB\ is spread only across a single RAID group (in this
case consisting of just eight storage devices from a pool of around one
hundred). The nature of the N-1 pattern and the size of the reads means that
all readers will almost always be reading within just a single RAID group. In
addition to limiting the bandwidth to only a few storage devices, it also
reduces the bandwidth from each of them due to overloading them with
non-sequential requests.

The results for the non-uniform restart can be similarly explained. The PanFS
results and explanation are essentially the same. The results for \plfs\ are
also very similar; better than PanFS due to spreading the reads across all
storage devices and not quite as good as the \plfs\ results for the uniform
restart. The difference between the uniform and non-uniform results for \plfs\
is only seen for small numbers of readers in the area where \plfs\ was helped
by prefetch in the uniform experiment. Since the readers are reading different
offsets than were written in the non-uniform restart, they will read multiple
data files instead of just reading a single one. The \upfs, unsurprising,
does not identify this admittedly strange pattern and therefore there is no
prefetch benefit. Only when there is no longer any prefetch benefit, 
do these results converge.

Although we are pleased to see that \plfs\ also does relatively well for the
archive copy experiment, we do not yet fully understand all of these results.
We can think of no reason why the bandwidths should be this low and we assumed
that PanFS would easily outperform \plfs\ due to having contiguous data within
its RAID groups instead of having data spread across multiple data files
within a \plfs\ container.  However, we are not surprised to see the read
bandwidths drop as the number of writers increases.  Increasing numbers of
writers results in a data layout with fewer large contiguous regions of data
for both \plfs\ and PanFS; therefore, the read bandwidth will suffer
accordingly.

%this behavior has been
%independently observed thrice that we know of: once by us, and twice by two
%groups of students at CMU.

\subsubsection{Metadata Queries}

As we discussed in Section~\ref{arch-meta}, there are currently two techniques
used to discover the metadata for a \plfs\ file. For a file being currently
written, metadata is discovered by \syscall{stat'ing} individual data files.
When a file is closed, metadata information is cached as file names within a
specific subdirectory inside the container. Thereafter, metadata information
can be discovered merely by issuing a \syscall{readdir}. Obviously, the first
technique is much slower; if our primary interest was in optimizing metadata
query rates, than we could cache this metadata following every write. However,
since \plfs\ is designed for checkpoint writing, we do not consider this
technique.

Figure~\ref{fig-metadata} compares these two times against the time it takes to
query a closed file written directly to the \upfs, PanFS. We have not yet
measured the time to query an open file on PanFS. For this experiment,
conducted on LANL's \rrz, we created two sets of 20~\GB\ files each written by
a different number of writers all issuing 47001 byte-sized writes. One set was
created directly on PanFS; the other indirectly through \plfs.  As the graph
shows, the \syscall{stat} times for closed files on \plfs\ and PanFS are
approximately the same.  However, as expected, the time to query a \plfs\ file
open for writing is greater than to query a closed file.

%plateaus for \plfs may be due to \Term{statahead} in the PanFS client.
%We suspect this is due to the small number of writers available on
%the \rrz; we hope to rerun this experiment on Roadrunner itself.

\input{metadata-fig.tex}

% What's up with the weird data for the stat time on an open PLFS file?
% I thought maybe it had to do with the number of subdirectories so I
% reran it with 999 subdirectories and it didn't change. Not sure now.

### 2009-07-17 SC09/future.tex
Adding the SC paper in its source form.
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@2 979fb63a-54b4-4514-8d67-9ab8ea1b59df

\section{Current Status and Future Work}
\label{future}

\scribble{I removed the 'cute' part of this. Sorry, john}
We initially developed \plfs\ in order to test our hypothesis that an
interposition layer could rearrange checkpoint patterns such that the
convenience of N-1 patterns could be preserved while achieving the bandwidth of
N-N.  However, after achieving such large improvements with real LANL
applications, we were compelled to harden it 
into a production file system. Consisting of about three thousand lines of
code, \plfs\ is currently a permanent mount point on Roadrunner where several
applications have begun evaluating whether it is appropriate for their write 
patterns.

%Since \plfs\ consists of only three thousand
%source lines of code and is easily ported to all Linux architectures, we
%decided~\footnotemark that the costs of maintaining it are easily outweighed by
%the benefits.
%\footnotetext{were persuaded by management}

Although \plfs\ works transparently with unmodified applications, in practice
some applications may need to make minor changes.  Applications that open
\plfs\ files in read-write mode will find that reading from these files can be
very slow.  As we discussed in Section~\ref{arch-read}, reads in \plfs\ are
handled by reading and aggregating the index files.  When files are opened in
write-read mode, this process must be repeated for every \syscall{read} since
intervening \syscall{writes} may have occurred.  Although the vast majority of
HPC applications are able to read in read-only mode, we do plan to investigate
removing this limitation in the future for those few applications who must occasionally
read while writing.  One possibility is to introduce metadata
servers which can synchronously maintain an aggregated index in memory.

Originally intended merely to address N-1 challenges, \plfs\, as currently
designed, also has the potential to address N-N challenges.  One way that
\plfs\ can reduce N-N checkpointing times is by reducing disk seeks through its
use of log-structured writing.  However, we have yet to measure the
frequency of non-sequential IO within N-N checkpoints.

\scribble{Added gigaplus reference to the end. --Milo}
Another challenge of an N-N pattern is the
overwhelming metadata pressure resulting from the simultaneous creation of tens
of thousands of files within a single directory.  Currently HPC parallel file
systems do distribute metadata across multiple metadata servers; however they do so
at the granularity of a \Term{volume} or a directory (\ie\ all files within a
directory share a single metadata server).  \plfs\ can easily refine this
granularity by distributing container subdirectories 
across multiple metadata servers.  At the moment, \plfs\ only creates container
structures for regular files; directories are simply created directly on the
\upfs.  By extending \plfs\ to create a similar container structure for
directories, we believe that \plfs\ can effectively address this N-N challenge
as well, in a manner similar to~\cite{gigaplus}.

### 2009-07-17 SC09/intro.tex
Adding the SC paper in its source form.
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@2 979fb63a-54b4-4514-8d67-9ab8ea1b59df

\section{Introduction}

\scribble{Need to define "its class" --Milo}
% computers are large, they have failures
In June 2008, Los Alamos National Labs (LANL), in partnership with IBM, broke
the petaflop barrier and claimed the top spot on the Top 500 list~\cite{top500}
with the Roadrunner supercomputer~\cite{roadrunner}.  Due to its unique hybrid
architecture, Roadrunner has only 3060 nodes, which is a relatively small
number for a supercomputer of its class.  Unfortunately, even at this size,
component failures are a frequent event.  These failures are particularly
problematic as many applications at LANL, and other High Performance Computing
(HPC) sites, have long run times in excess of days, weeks, and even months.

% checkpoint is defense against failure, drives parallel file systems, is
% challengin
Typically these applications protect themselves against failure by periodically
\Term{checkpointing} their progress by saving the state of the application to
persistent storage. After a failure the application can then restart from the most recent 
checkpoint.  Due to the difficulty of reconstructing a consistent
image from multiple asynchronous checkpoints~\cite{lamport1978}, HPC
applications checkpoint synchronously (\ie\ following a barrier).  Synchronous
checkpointing, however, does not eliminate the complexity; it merely shifts it to
the parallel file system which now must coordinate simultaneous access from
thousands of compute nodes.  Even using optimal checkpoint frequencies~\cite{daly-optimal}, 
checkpointing has become the driving workload for
parallel file systems and the challenge it imposes grows with each successively
larger supercomputer~\cite{pcl:99:me,schroeder2007}.  The difficulty of this
challenge can vary greatly depending on the particular pattern of checkpointing
chosen by the application.

% our approach
In this paper, we describe different checkpointing patterns, and show how some
result in very poor storage performance on three of the major HPC
parallel file systems: PanFS, GPFS, and Lustre.  We then posit that an
interposition layer inserted into the existing storage stack can rearrange this
problematic access pattern to achieve much better performance from the
\upfs.  To test this hypothesis, we have developed \plfs, a Parallel
Log-structured File System, which is one such interposition layer.  We present measurements using \plfs\ on several synthetic benchmarks and real applications at multiple HPC supercomputing centers.  The
results confirm our hypothesis: writing to the \upfs\ through
\plfs\ improves checkpoint bandwidth for all tested applications and
benchmarks and on all three studied parallel file systems; in some cases,
bandwidth is raised by several orders of magnitude.

\input{summary-fig}

As we shall discuss, and is summarized in Figure~\ref{fig-page1}, PLFS is
already showing very large speed ups for widely used HPC benchmarks and
important real HPC codes, and at extreme scale.

% the stupid mandatory list of the rest of the paper
The rest of the paper is organized as follows.  We present more detailed background and
motivation in Section~\ref{motivation}, describe our design in
Section~\ref{arch} and our evaluation in Section~\ref{eval}.  We present
related work in Section~\ref{related}, current status and future work in
Section~\ref{future}, and finally we conclude in Section~\ref{conclude}.

### 2009-07-17 SC09/metadata-fig.tex
Adding the SC paper in its source form.
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@2 979fb63a-54b4-4514-8d67-9ab8ea1b59df

\begin{SCfigure}
    \centering
    \includegraphics[width=0.4\textwidth]{data/stat/stat.eps}
    \mycaption{fig-metadata}{Metadata Queries.}{
        This graph compares the time to \syscall{stat} a 20~\GB\ file
        on \plfs\ with the time to \syscall{stat} a comparable
        file on PanFS.    
        The \yaxis, which is in logscale,
        shows the \syscall{stat} time 
        as a function of the number of writers who created that
        file.  We compare the PanFS \syscall{stat} rate to the
        \plfs\ rate for files opened for writing, and to the 
        \plfs\ rate for closed files.
        \vspace{1cm}
    }
\end{SCfigure}

### 2009-07-17 SC09/motivation.tex
Adding the SC paper in its source form.
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@2 979fb63a-54b4-4514-8d67-9ab8ea1b59df

\section{Background}
\label{motivation}

% Why do we care about checkpointing
For large parallel simulations which can run for several months, restarting a
calculation from the beginning is not viable, since the mean time to interrupt
for today's supercomputers is measured in days or hours and not
months~\cite{google-failure,schroeder-failure}.  As a result, applications 
periodically save
the current state of their computation to persistent storage.  Later, after a
failure is encountered, the application can restart from the last saved
checkpoint.

% N-N compared to N-1

\scribble{Lots of tweaking of final sentence} 
From a file system perspective,
there are two basic checkpointing patterns: ~\Term{N-N} and ~\Term{N-1}.  An
N-N checkpoint is one in which each of ~\Math{N} processes writes to a unique
file, for a total of ~\Math{N} files written.  An N-1 checkpoint differs in
that all of ~\Math{N} processes write to a single shared file.  Applications
using N-N checkpoints usually write sequentially to each file, an access
pattern ideally suited to parallel file systems. Conversely, applications using
N-1 checkpoint files typically organize the collected state of all N processes
in some application specific, canonical order, often resulting in 
small, unaligned, interspersed writes.

Some N-1 checkpoint files are logically the equivalent of concatenating the
files of an N-N checkpoint (\ie\ each process has its own unique region
within the shared file).  This is referred to as an ~\Term{N-1 segmented}
checkpoint file and is extremely rare in practice.  More common is an
~\Term{N-1 strided} checkpoint file in which the processes write multiple small
regions at many different offsets within the file; these offsets are typically
not aligned with file system block boundaries~\cite{plfs-maps}.  N-1 strided
checkpointing applications often make roughly synchronous progress such that
all the processes tend to write to the same region within the file concurrently, and collectively this region sweeps across the file.
These three patterns, N-N, N-1 segmented, and N-1 strided, are
illustrated in Figure~\ref{fig-patterns}.  Since N-1 segmented is a rare pattern
in practice, hereafter we consider only N-1 strided and we refer to it with the
shorthand N-1.

\input{patterns-fig}
\scribble{Should probably explain or remove 'false sharing' in figure}
\scribble{If those Argonne graphs were decipherable in B-W, I'd like to include
two of them to visualize the difference between strided and non.}

% file system challenges for N-N and N-1
The file system challenge for an N-N workload is the concurrent creation of
thousands of files which are typically within a single directory.  An N-1
workload can be even more challenging however for several different reasons
which may depend on the particular parallel file system or the underlying
\RAID\ protection scheme.  In the Panasas ActiveScale parallel
file system~\cite{Welch04managingscalability} (PanFS), for example, small strided
writes within a parity stripe must serialize in order to maintain consistent
parity.  In the Lustre parallel file system~\cite{lustre}, writes to N-1
checkpoint files which are not aligned on file system block boundaries cause
serious performance slowdowns as well~\cite{nersc}.  Although both
N-N and N-1 patterns pose challenges, it has been our observation, as well
as that of many others~\cite{grider-hec,adios,nersc,zest,Thakur99datasieving},
that the challenges of N-1 checkpointing are more difficult. 
Applications using N-1 patterns consistently achieve significantly less
bandwidth than do those using an N-N pattern.

Figure~\ref{fig-motivation} presents experimental data validating this discrepancy: An N-1 checkpoint pattern receives
only a small fraction of the bandwidth achieved by an N-N pattern on PanFS,
GPFS, and Lustre and does not scale with increased numbers of nodes.  The 
PanFS experiment, run on LANL's Roadrunner
supercomputer, shows a maximum observed N-N bandwidth of 31~\GBs\ compared
to a maximum observed N-1 bandwidth of less than 100~\MBs. 
Although we show PanFS results
using their default RAID-5 configuration, PanFS has also a RAID-10
configuration which reduces the implicit sharing caused by N-1 patterns when
two writers both need to update the same parity.  While this solution improves
scaling and offers much higher N-1 write bandwidth without sacrificing
reliability, it does so by writing every byte twice, a scheme that, at best,
can achieve only approximately half of the write bandwidth of N-N on RAID-5.
\plfs, however, as will be shown in Section~\ref{eval}, can get much closer.

The GPFS and Lustre experiments were run on much smaller systems. The GPFS
experiment was run using an archive attached to Roadrunner using its 
nine, quad-core, file transfer nodes.  The Lustre experiment was run using five
client machines, each with eight cores, and twelve Lustre servers.  All three
file systems exhibit similar behavior; N-N bandwidths are consistently higher
than N-1 by at least an order of magnitude.

% why all applications not N-N
Since N-N checkpointing derives higher bandwidth than N-1, the obvious path to
faster checkpointing is for application developers to rewrite existing N-1
checkpointing applications to do N-N checkpointing instead.  Additionally, all
new applications should be written to take advantage of the higher bandwidth
available to N-N checkpointing.  Although some developers have gone this route,
many continue to prefer an N-1 pattern even though its disadvantages are well
understood.  There are several advantages to N-1 checkpointing that appeal to
parallel application developers.  One, a single file is much easier to manage
and to archive.  Two, N-1 files usually organize data into an application
specific canonical order that commonly aggregates related data together in
contiguous regions, making visualization of intermediate state simple and
efficient.  Additionally, following a failure, a restart on a different number
of compute nodes is easier to code as the checkpoint format is independent of
the number of processes that captured the checkpoint; conversely, gathering the
appropriate regions from multiple files or from multiple regions within a
single file is more complicated.

\input{motivation-graphs}

% a list of some apps that do N-1

Essentially these developers have once again shifted complexity to the parallel
file system for their own convenience.  This is not unreasonable; it has long
been the province of computer systems to make computing more convenient for its
users.  Many important applications have made this choice.  Of the twenty-three
applications listed on the Parallel I/O Benchmarks page~\cite{pio-benchmarks},
at least ten have an N-1 pattern; two major applications at LANL use an N-1
checkpointing pattern as do at least two of the eight applications chosen to
run on Roadrunner during its initial stabilization phase.  N-1 checkpointing is
very important to these applications.  For example, at the core of one of
LANL's N-1 applications is a twenty-year old Fortran checkpointing library. 
About a decade ago, in response to a growing clamor about the
limitations of N-1 checkpointing bandwidth, developers for this application
augmented their checkpointing library with fifteen thousand lines of code.
However, instead of changing the application to write an N-N pattern, they
added a prefer to the IO routiones in which interprocess communication is used to aggregate and buffer writes.
Although they did not improve the performance to match that of other 
applications using an N-N checkpoint pattern, 
this effort was considered a success as they did improve the N-1
performance by a factor of two to three. This new checkpointing library,
called \Term{bulkio}, has been maintained over the past decade and ported to
each new successive supercomputer at LANL~\cite{bent-personal-bulk}. 
Furthermore, N-1 patterns continue to be developed anew in many new
applications.  High level data formatting libraries such as Parallel
NetCDF~\cite{pnetcdf} and HDF5~\cite{hdf5} offer convenience to application
developers who simply describes the logical format of their data and need no
longer consider how that data is physically organized in the file system.  Once
again this convenience merely shifts the complexity to the \upfs\ 
since these libraries use an N-1 pattern.

% Milo's change which I didn't like
%More generally, parallel application developers employ data formatting libraries such as NetCDF~\cite{pnetcdf} and HDF5~\cite{hdf5} to simplify programming, but these extra layers of mapping can lead to decreased locality and sequentiality when checkpointing a logically related set of variables. N-1 write access patterns can appear in new applications this way.

### 2009-07-17 SC09/overhead-fig.tex
Adding the SC paper in its source form.
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@2 979fb63a-54b4-4514-8d67-9ab8ea1b59df

\begin{SCfigure}
    \centering
    \includegraphics[width=0.4\textwidth]{data/overhead/gpfs.eps}
    \mycaption{fig-overhead}{Overhead.}{
        Ideally, N-1
        patterns written through \plfs\ would be able to 
        achieve the bandwidth available to an N-N pattern
        written directly to the \upfs.  However, 
        this graph shows that various overheads make this
        difficult.  Even though
        there is overhead, the important
        point is that \plfs\ still allows an N-1 pattern to
        be written much more quickly than if it was written
        directly to the \upfs. 
        \vspace{.5cm}
    }
\end{SCfigure}

### 2009-07-17 SC09/patterns-fig.tex
Adding the SC paper in its source form.
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@2 979fb63a-54b4-4514-8d67-9ab8ea1b59df

\begin{figure}[tb]
  \centering
    \fbox{
        \subfloat[N-N]{\label{fig:nn}
            \includegraphics[width=0.25\textwidth]{figs/nn-arrows.eps}}
    }
    \fbox{
        \subfloat[N-1 segmented]{\label{fig:n1segmented}
           \includegraphics[width=0.25\textwidth]{figs/n1nonstrided-arrows.eps}}
    }
    \fbox{
        \subfloat[N-1 strided]{\label{fig:n1strided}
            \includegraphics[width=0.25\textwidth]{figs/n1strided-arrows.eps}}
    }
    \mycaption{fig-patterns}{Common Checkpointing Patterns.}{
This figure shows the three basic checkpoint patterns: from left to right, N-N,
N-1 segmented, and N-1 strided.  In each pattern, the parallel application is
the same, consisting of six processes spread across three compute nodes each of
which has three blocks of state to checkpoint.  The difference in the three
patterns is how the application state is logically organized on disk.  In the
N-N pattern, each process saves its state to a unique file.  N-1 segmented is
simply the concatenation of the multiple N-N files into a single file.
Finally, N-1 strided, which also uses a single file as does N-1 segmented, has
a region for each block instead of a region for each process.  From a parallel
file system perspective, N-1 strided is the most challenging pattern as it is
the most likely to cause small, interspersed, and unaligned writes.  Note that
previous work~\cite{hedges-lunatic} refers to N-N as \Term{file per process}
and N-1 as \Term{shared file} but shares our segmented and strided terminology.
}
\end{figure}

### 2009-07-17 SC09/plfs.tex
Adding the SC paper in its source form.
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@2 979fb63a-54b4-4514-8d67-9ab8ea1b59df

\documentclass[10pt]{article}
\pagenumbering{arabic}
\pagestyle{plain}
%\pagestyle{empty}
\usepackage{fullpage}
\usepackage{setspace}
\usepackage[top=2cm,bottom=2cm,left=2cm,right=2cm]{geometry}
\usepackage{wrapfig}
\usepackage{graphicx}
\usepackage{sidecap}
\usepackage{subfig}
\usepackage{eso-pic,type1cm,color}

\input{macros}
\input{watermark}

\begin{document}
%\doublespace
\input{title}
\input{abstract}
\input{intro}
\input{motivation}
\input{arch}
\input{eval}
\input{related}
\input{future}
\input{conclude}
\input{bib}
\end{document}

### 2009-07-17 SC09/read-graphs.tex
Adding the SC paper in its source form.
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@2 979fb63a-54b4-4514-8d67-9ab8ea1b59df

\begin{figure}[tb]
    \centering
        \readgraph{Uniform Restart}{nn}{data/reads/nn.eps}
        \readgraph{Non-uniform Restart}{nm}{data/reads/nm.eps}
        \readgraph{Archive Copy}{n1}{data/reads/archive.eps}
    \\
    \mycaption{fig-reading}{Read Bandwidth.}{
These three graphs show the results of our read measurements on the
\rrz. We created a set 20~\GB\ N-1 checkpoint files through
\plfs\ and another directly on PanFS.  Each file was produced by a different
number of writers; all of the \syscall{writes} were 47001 bytes in size.  For
each graph, the \yaxis\ shows the read bandwidth as a function of the number of
writers who created the file.  The graph on the left shows the read bandwidth
when the number of readers is the same as the number of writers, as is the case
in a typical uniform restart; in this case, the size of the reads is the same
as the size of the original writes.  The graph in the middle emulates a
non-uniform restart in which the application resumes on one fewer compute
nodes; in this case, the size of the \syscall{reads} is slightly larger than
the size of the original \syscall{writes}.  Finally, the graph on the right
shows the read bandwidth when there only four readers; we used LANL's archive
copy utility and modelled the common scenario of copying checkpoint data to an
archive system using a relatively small number of readers.  To enable
comparison across graphs, the axis ranges are consistent.
}

\end{figure}

### 2009-07-17 SC09/related-table.tex
Adding the SC paper in its source form.
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@2 979fb63a-54b4-4514-8d67-9ab8ea1b59df

\newcommand{\C}[1]{\centering{#1}}
%\newcommand{\C}[1]{\begin{center}#1\end{center}}
\hyphenation{Immediately}
\begin{table}
\begin{center}
\begin{tabular}{|r|p{19mm}|p{20mm}|p{23mm}|p{15mm}|p{19mm}| p{19mm}|}
\hline 
{\em } & 
\C{{\em Interposition Technique Used}} & 
\C{{\em No Extra Resources Used During}} & 
\C{{\em No Extra Resources Used After}} & 
\C{{\em Maintains Logical Format}} & 
\C{{\em Works with Unmodified Applications}} & 
\C{{\em Data Immediately Available}} 
\tabularnewline \hline
ADIOS & \C{Library} & \C{Yes} & \C{Yes} & \C{Yes} &  \C{No} & \C{Yes}
\tabularnewline \hline
{\tt stdchk} & \C{FUSE} & \C{No (LD, M)} & \C{No (LD, N, M)} & \C{Yes} & \C{Yes} & \C{Yes}
\tabularnewline \hline
%Neighbor &  \C{FUSE} &\C{No (M)} & \C{No (M, N)} & \C{Yes} & \C{Yes} & \C{Yes} 
%\tabularnewline \hline
Diskless &  \C{Library} &\C{No (M)} & \C{No (M)} & \C{No} & \C{No} & \C{Yes} 
\tabularnewline \hline
ZEST &  \C{FUSE} &\C{No (RD)} & \C{No (RD)} & \C{No} & \C{No} & \C{No} 
\tabularnewline \hline
PLFS & \C{FUSE} & \C{Yes} & \C{Yes} & \C{Yes} & \C{Yes} & \C{Yes} 
\tabularnewline \hline
\end{tabular}
\mycaption{tab-feature}{Techniques for improving N-1 Checkpointing}{
This table presents a comparison of the various techniques for reducing
N-1 checkpoint times.  For exposition, we have used various
shorthands: 
%Neighbor for Neighbor's Memory,
Diskless for Diskless 
Checkpointing, LD for local disk on the compute nodes, RD for remote disk 
on the storage system, M for memory, and N for network.  
}
\end{center}
\end{table}

\if 0
other possible columns:

            Speed
    ADIOS: Disk                                                                 
    stdck: Disk                                                                 
    Neighbor: Network                                                           
    Diskless: Network                                                           
    ZEST: Spindle                                                               
    PLFS: N-N

lines of code, user or kernel mode, interposition technique, maintainability,
portability

also if we add more columns we'll have to rotate it so columns become rows and
rows become columns.

\fi

### 2009-07-17 SC09/related.tex
Adding the SC paper in its source form.
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@2 979fb63a-54b4-4514-8d67-9ab8ea1b59df

\section{Related Work}
\label{related}

% Remzi says refer to Erez Zadok's work on stackable file systems
% also says to refer to Thain's interposition work.  But we already did.

%Sudharshan Vazhkudai has several papers on checkpointing, typically using
%memory or scavenged local disk as a buffer.  Descriptions available at his
%webpage: \url{http://www.csm.ornl.gov/~vazhkuda/}.

%Frank Shorter's Master Thesis at Clemson is a big discussion about parallel
%benchmarks: \url{ftp://ftp.parl.clemson.edu/pub/techreports/2003/PARL-2003-001.ps}.

Translating random access checkpoint writes into a sequential pattern is an
idea which extends naturally from work on log-structured file
systems~\cite{lfs} such as NetApp's WAFL~\cite{WAFL} and Panasas's Object
Storage~\cite{welch08scalableperformance}. While these ideas reorganize disk
layout for sequential writing, they do not decouple concurrency caused by
multiple processes writing to a single file.  Another approach to
log-structuring N-1 patterns addressed only physical layout and also did not
decouple concurrency~\cite{pvfs-log}.  We believe our contribution of
rearranging N-1 checkpoints into N-N is a major advance.

Checkpoint-restart is an old and well studied fault tolerance strategy. A broad
evaluation of rollback-recovery protocols and checkpointing and a discussion of
their future at the petascale can be found in \cite{survey1, survey2}.

Berkeley Lab Checkpoint/Restart~\cite{blcr} and Condor
Checkpointing~\cite{condor-ckpt} both allow unmodified applications to
checkpoint the state of a single node.  They leverage the operating system's
ability to swap a process to save the totality of the swap image persistently.
The great strength of this approach is that it can be used by applications that
have no internal checkpoint mechanism.  A disadvantage is that these
checkpoints are larger then when an application specifies only the exact data
that it needs saved.

%allows user applications to checkpoint their state, including system
%resources, without the application programmer having to explicitly code the
%process. In this regard, it has a similar philosophy to \plfs\ of making
%checkpointing easier for the application developer, but our focus is on
%enabling better checkpointing performance for existing applications rather than
%writing an entire checkpointing system.

%\scribble{Couldn't find the incremental checkpointing reference}

Conversely, incremental checkpointing and memory
exclusion~\cite{pcl:99:me,pxn:95:scd} reduce the size of checkpoint data by
only saving data that has changed since the previous checkpoint.  Buffering
with copy-on-write~\cite{li-298215} can also reduce checkpoint latency.

\input{related-table}

%A variant of N-N checkpoint uses the disks or memory of computer nodes to
%record a checkpoint writing it to storage as a background process during the
%next compute period.

{\tt stdchk}~\cite{stdchk} saves checkpoints into a cloud of free disk space
scavenged from a cluster of workstations.  A similar
approach~\cite{aggregate-memory} reserves compute node memory to temporarily
buffer checkpoints and then asynchronously saves them to persistent storage.
Diskless checkpointing~\cite{diskless} also  
saves checkpoints into compute node memory, but does not subsequently transfer
them to persistent storage.  Rather it achieves fault protection by using erasure 
coding on the distributed checkpoint image. 
Although these techniques work well for many applications, large HPC parallel
applications jealously utilize all memory and demand a high degree of
determinism in order to avoid jitter~\cite{DusseauPhd98} and are therefore seen
to be poorly served by techniques reducing available memory or techniques which
require background processes running during computation.

The Adaptable IO System (ADIOS)~\cite{adios} developed by the Georgia Institute
of Technology and Oak Ridge National Laboratory provides a high-level IO API
that can be used in place of HDF5 to do much more aggressive write-behind and
log-like reordering of data location within the checkpoint. While this technique
requires application modification, it enables interoperability with other
middleware libraries.

ZEST~\cite{zest}, developed at Pittsburgh Supercomputing Center, is a file
system infrastructure that is perhaps the most similar in philosophy to \plfs,
particularly in its borrowing of concepts from log-structured file systems.
Rather than each client process pushing their writes sequentially to storage,
in ZEST manager threads running on behalf of each disk pull data from a series
of distributed queues, in a technique borrowed from
River~\cite{Arpaci-Dusseau03-TOCS}.  The key aspect of ZEST is that no
individual write request is ever assigned to a specific disk; disks pull from
these queues whenever they are not busy, resulting in high spindle utilization
even in a system where some devices are performing more slowly than others.
Unlike \plfs, however, data is not immediately available to be read, requiring
a subsequent phase to first rewrite the file before it can be accessed.  Since
this phase happens in the relatively long time between checkpoints and since it
happens on the server nodes and not on the compute nodes, this subsequent
rewrite phase is not typically problematic for a dedicated checkpoint file
system.

### 2009-07-17 SC09/summary-fig.tex
Adding the SC paper in its source form.
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@2 979fb63a-54b4-4514-8d67-9ab8ea1b59df

\begin{SCfigure}
    \centering
    \includegraphics[width=0.6\textwidth]{data/summary/page1.eps}
    \mycaption{fig-page1}{Summary of our results.}{
        This graph summarizes our results which will be
        explained in detail in Section~\ref{eval}.  
        The key observation here is that our technique
        has improved checkpoint bandwidths for all seven
        studied benchmarks and applications by up to
        several orders of magnitude.  
        \vspace{1.0cm}
    }
\end{SCfigure}

### 2009-07-29 SC09/plfs.tex
messed w/ graph fonts some more
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@4 979fb63a-54b4-4514-8d67-9ab8ea1b59df
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@5 979fb63a-54b4-4514-8d67-9ab8ea1b59df
got the graphs looking good. got the abstract right. got the keywords, crap, authors, copyright.
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@7 979fb63a-54b4-4514-8d67-9ab8ea1b59df

@@
-\documentclass[10pt]{article}
+%\documentclass[10pt]{article}
+\documentclass{sig-alternate}
 \pagenumbering{arabic}
@@
 %\pagestyle{empty}
-\usepackage{fullpage}
-\usepackage{setspace}
-\usepackage[top=2cm,bottom=2cm,left=2cm,right=2cm]{geometry}
-\usepackage{wrapfig}
+%\usepackage{fullpage}
+%\usepackage{setspace}
+%\usepackage[top=2cm,bottom=2cm,left=2cm,right=2cm]{geometry}
+%\usepackage{wrapfig}
 \usepackage{graphicx}
@@
 \input{macros}
-\input{watermark}
+%\input{watermark}
 
 \begin{document}
+\conferenceinfo{SC09} {November 14--20, Portland, Oregon, USA.}
+\CopyrightYear{2009}
+\crdata{978-1-60558-744-8/09/11}
+\boilerplate{(c) 2009 Association for Computing Machinery. ACM acknowledges that
+this contribution was authored or co-authored by a contractor or
+affiliate of the U.S. Government. As such, the Government retains a
+nonexclusive, royalty-free right to publish or reproduce this article,
+or to allow others to do so, for Government purposes only.}
 %\doublespace
@@
 \input{abstract}
+\category{D.4.3}{Operating Systems}{File Systems Management}{File organization}
+\terms{Performance, Design}
+\keywords{High performance computing, parallel computing, checkpointing}
 \input{intro}

### 2009-07-29 SC09/abstract.tex
got the graphs looking good. got the abstract right. got the keywords, crap, authors, copyright.
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@7 979fb63a-54b4-4514-8d67-9ab8ea1b59df
attached is the most recent version of the paper.  I'll do another big push tomorrow and then I'll submit it since I'll be on vacation all next week.  I think I've addressed most of the reviewer's comments but here's a couple that maybe still need to be addressed (or ignored):
the addition of metadata benchmark such as mdtest would make paper
stronger

the paper does not mention about file locks.  f

@@
 \em{
+\if 0
 Parallel applications running across thousands of processors must protect
@@
 adoption of complex formatting libraries such as NetCDF and HDF5. 
+\fi
 
-\if 0
 % here's the 150 word limited version we submitted to SC
-Parallel applications running across thousands of processors must protect themselves from inevitable system failures.
-Many applications insulate themselves from failures by checkpointing.
-For many applications, checkpointing into a shared single file is most convenient.
-With such an approach, the size of writes are often small and not aligned with file system boundaries.
-Unfortunately for these applications, this preferred data layout results in pathologically poor performance from the underlying file system which is optimized for large, aligned writes to non-shared files.
-To address this fundamental mismatch, we have developed a virtual parallel log structured file system, PLFS.
-PLFS remaps an application’s preferred data layout into one which is optimized for the underlying file system.
-Through testing on PanFS, Lustre, and GPFS, we have seen that this layer of indirection and reorganization can reduce checkpoint time by an order of magnitude for several important benchmarks and real applications.
-\fi
+Parallel applications running across thousands of processors must protect
+themselves from inevitable system failures. Many applications insulate
+themselves from failures by checkpointing. For many applications, checkpointing
+into a shared single file is most convenient. With such an approach, the size
+of writes are often small and not aligned with file system boundaries.
+Unfortunately for these applications, this preferred data layout results in
+pathologically poor performance from the underlying file system which is
+optimized for large, aligned writes to non-shared files. To address this
+fundamental mismatch, we have developed a virtual parallel log structured file
+system, PLFS. PLFS remaps an application's preferred data layout into one which
+is optimized for the underlying file system. Through testing on PanFS, Lustre,
+and GPFS, we have seen that this layer of indirection and reorganization can
+reduce checkpoint time by an order of magnitude for several important
+benchmarks and real applications.  
 } 

### 2009-07-29 SC09/eval.tex
attached is the most recent version of the paper.  I'll do another big push tomorrow and then I'll submit it since I'll be on vacation all next week.  I think I've addressed most of the reviewer's comments but here's a couple that maybe still need to be addressed (or ignored):
the addition of metadata benchmark such as mdtest would make paper
stronger

the paper does not mention about file locks.  file system semantics is
not clear.

there should be a discussion about file system approach vs library
approach.  library can speed up N-1 w/out sacrificing performance for
other ops like metadata q

@@
 Test}~\cite{mpi-io-test}, on three different parallel file systems, PanFS, GPFS,
-and Lustre. For each of these graphs, the size of each \syscall{write} was
-47001 bytes (a small, unaligned number observed in actual applications to be
-particularly problematic for file systems). \syscall{Writes} were
-issued until two minutes had elapsed. Although this is atypical since 
-applications tend to write a fixed amount of data instead of writing for
-a fixed amount of time, we have observed that this allows representative 
-bandwidth measurements with a predictable runtime.
-
+and Lustre. 
 There are several things to notice in these graphs. The first is that these
@@
 Although \plfs\ does improve the bandwidth of N-1 patterns on GPFS and Lustre,
-the improvement is not as large as it is on PanFS. This is because the scale
-of the experiments on PanFS are 200 times larger than on the other two
-platforms. In the inset in Figure~\ref{eval:panfs} at the extreme low values
-for the number of processes, we see that \plfs\ does not scale N-1 bandwidth
-as fast as N-N scales for PanFS as well. This is due to overheads incurred by
-both FUSE and \plfs; these overheads limit the total bandwidth achieved by any
-single compute node relative to N-N. For HPC systems at extreme scale this
-limitation does not matter since aggregate bandwidth across multiple nodes is
-more relevant than the bandwidth from a single node. In this case the real
-limiting factor is the total bandwidth capacity of the storage fabric, which
-generally saturates when using only a small fraction of compute nodes.
-However, with poorly-behaved IO patterns (\ie\ N-1), even very large jobs may
-not reach these bandwidth limits because they will be more severely
-constrained by file system limitations as exemplified by
-Figure~\ref{fig-motivation}. \plfs\ is designed to remove these file system
-limitations, so that jobs can achieve higher bandwidth and reach the
-same limitation of the storage fabric which is constraining N-N bandwidth.
+the improvement is not as large as it is on PanFS. This is because the scale of
+the experiments on PanFS are 200 times larger than on the other two platforms.
+In the inset in Figure~\ref{eval:panfs} at the extreme low values for the
+number of processes, we see that \plfs\ does not scale N-1 bandwidth as fast as
+N-N scales for PanFS as well. This is due to overheads incurred by both FUSE
+and \plfs; these overheads limit the total bandwidth achieved by any single
+compute node relative to N-N. For HPC systems at extreme scale this limitation
+does not matter since aggregate bandwidth across multiple nodes is more
+relevant than the bandwidth from a single node. In this case the real limiting
+factor is the total bandwidth capacity of the storage fabric, which generally
+saturates when using only a small fraction of compute nodes.  However, with
+poorly-behaved IO patterns (\ie\ N-1), even very large jobs may not reach these
+bandwidth limits because they will be more severely constrained by file system
+limitations as exemplified by Figure~\ref{fig-motivation}. \plfs\ is designed
+to remove these file system limitations, so that jobs can achieve higher
+bandwidth and reach the same limitation of the storage fabric which is
+constraining N-N bandwidth.  We will examine the FUSE and PLFS overheads in
+greater detail in Section~\ref{overhead}.
 
-An unusual feture of the inset graph in Figure~\ref{eval:panfs} is that there
-are localized bottlenecks about 700 processes.  This is due to the
-architecture of Roadrunner and it's connection to PanFS.  Roadrunner is split
-into seventeen sub-clusters, \Term{CUs}, each of which can run 720 proceses.
-The CUs are fully connected via an Infiniband fat tree, however, in order to
-minimize network infrastructure costs, the storage system is partitioned into
-multiple sub-nets. Each CU currently has six specialized IO nodes, one for
-each sub-net; these six IO nodes are the storage bandwidth bottleneck.
-Therefore, as a job grows within a CU, it quickly saturates the IO node
-bandwidth; when it grows into another CU, its bandwidth increases sharply due
-to gaining six additional IO nodes: this explains the "stair-step" behavior of
-the N-N and \plfs\ lines in this graph.
+An unusual feature of the inset graph in Figure~\ref{eval:panfs} is that there
+are localized bottlenecks about 700 processes.  This is due to the architecture
+of Roadrunner and it's connection to its 1000 blade PanFS storage system.
+Roadrunner is split into seventeen sub-clusters, \Term{CUs}, each of which can
+run 720 proceses.  The CUs are fully connected via an Infiniband fat tree,
+however, in order to minimize network infrastructure costs, the storage system
+is partitioned into multiple sub-nets. Each CU currently has six specialized IO
+nodes, one for each sub-net; these six IO nodes are the storage bandwidth
+bottleneck.  Therefore, as a job grows within a CU, it quickly saturates the IO
+node bandwidth; when it grows into another CU, its bandwidth increases sharply
+due to gaining six additional IO nodes: this explains the ``stair-step"
+behavior of the N-N and \plfs\ lines in this graph.
 
@@
 saturate the available storage bandwidth and thus neither the N-N nor the
-\plfs\ lines in these graphs reach a storage bandwidth bottleneck. We will
-examine the FUSE and PLFS overheads in greater detail in
-Section~\ref{overhead}.
+\plfs\ lines in these graphs reach a storage bandwidth bottleneck.
 
@@
 
-The second row in Figure~\ref{fig-eval}, consisting of
-Figures~\ref{eval:lanl1},~\ref{eval:lanl2},~and~\ref{eval:lanl3}, shows the
-results of using \plfs\ to improve the bandwidth of three important LANL
-applications which use N-1 checkpoint patterns. The application
-shown in Figure~\ref{eval:lanl1} is the application whose developers augmented
-their checkpoint routine with a new checkpoint library called bulkio, which
-aggregates and buffers writes into larger, contiguous writes more friendly to
-the \upfs. Therefore we show four lines in this graph: one for the old
-checkpoint method which uses MPI-IO routines to directly write data to PanFS, a
-line for the bulkio method writing directly to PanFS, and a second line for
-each of these techniques writing indirectly to PanFS via \plfs. Notice that
-\plfs\ is not able to improve the bulkio technique which loses a significant
-amount of bandwidth due to aggregating data and reducing parallelism by
-reducing the number of writers. However, when the older MPI-IO technique is
-used with \plfs, \plfs\ is able to exploit the additional parallelism possible
-by using all of the processes as writers. This data was collected on a small,
-Opteron-only, \rrz. The next two graphs show similar results; using \plfs\ to
-rearrange an N-1 pattern yields significantly improved bandwidths.
-Figure~\ref{eval:lanl2} was also run on the \rrz, whereas
-Figure~\ref{eval:lanl3} was run on Roadrunner itself; this can be seen by the
-much larger scale (on both the \yaxis\ and the \xaxis) for
-Figure~\ref{eval:lanl3}. Note that hese are extremely important applications
-at LANL; it has been estimated that they account for more than half of the
-total computing cycles at LANL. Although these applications must remain
-anonymous, traces of their IO are available~\cite{plfs-maps}.
+The second row in Figure~\ref{fig-eval} shows the results of using \plfs\ to
+improve the bandwidth of three important LANL applications which use N-1
+checkpoint patterns. The application shown in Figure~\ref{eval:lanl1} is the
+application whose developers augmented their checkpoint routine with a new
+checkpoint library called bulkio, which aggregates and buffers writes into
+larger, contiguous writes more friendly to the \upfs. Therefore we compare the
+bulkio method writing directly to PanFS to the old MPI-IO method writing
+indirectly to PanFS via \plfs.  This data was collected on Roadrunner.  The
+next two graphs show similar results; using \plfs\ to rearrange an N-1 pattern
+yields significantly improved bandwidths.  Figure~\ref{eval:lanl2} was run on a
+small \rrz\ which consists of 128 8-core compute nodes with a 66 blade PanFS
+storage system, whereas Figure~\ref{eval:lanl3} was run on Roadrunner itself.
+Note that these are extremely important applications at LANL; it has been
+estimated that they account for more than half of the total computing cycles at
+LANL.  Although these applications must remain anonymous, traces of their IO
+are available~\cite{plfs-maps}.
 
@@
 data written by each process decreases. BTIO~\cite{btio-io} in
-Figure~\ref{eval:btio} also shows a large improvement but we currently only
-have one data point; we will have more in the final version.
+Figure~\ref{eval:btio} also shows a large improvement but only
+for one data point.
 FLASH~IO~\cite{flash-io} and Chombo~IO\cite{chombo-io}, shown respectively in
@@
 which was improved in the best case by two orders of magnitude.
-
-[Note to the reviewer: Only FLASH and LANL3 were run at large-scale using both
-\plfs\ and the \upfs. We will have fuller results for the final experiment and
-we expect at least two orders of magnitude improvement for all of them except
-QCD.]
 
@@
 processes resuming the application is the same as the number that wrote the
-last checkpoint. In this case, each rank within the parallel job reads the
-same data in the same order as the previous writer of that same rank. In
+last checkpoint. In this case, each process within the parallel job reads the
+same data in the same order as one of the previous writers. In
 contrast, a \Term{non-uniform restart} is one in which the number of

### 2009-07-29 SC09/future.tex
attached is the most recent version of the paper.  I'll do another big push tomorrow and then I'll submit it since I'll be on vacation all next week.  I think I've addressed most of the reviewer's comments but here's a couple that maybe still need to be addressed (or ignored):
the addition of metadata benchmark such as mdtest would make paper
stronger

the paper does not mention about file locks.  file system semantics is
not clear.

there should be a discussion about file system approach vs library
approach.  library can speed up N-1 w/out sacrificing performance for
other ops like metadata q

@@
 into a production file system. Consisting of about three thousand lines of
-code, \plfs\ is currently a permanent mount point on Roadrunner where several
-applications have begun evaluating whether it is appropriate for their write 
-patterns.
+code, \plfs\ is currently an optional mount point on Roadrunner where several
+applications have begun using it to improve their N-1 checkpoint performance.
 
@@
 Although \plfs\ works transparently with unmodified applications, in practice
-some applications may need to make minor changes.  Applications that open
-\plfs\ files in read-write mode will find that reading from these files can be
-very slow.  As we discussed in Section~\ref{arch-read}, reads in \plfs\ are
-handled by reading and aggregating the index files.  When files are opened in
-write-read mode, this process must be repeated for every \syscall{read} since
-intervening \syscall{writes} may have occurred.  Although the vast majority of
-HPC applications are able to read in read-only mode, we do plan to investigate
-removing this limitation in the future for those few applications who must occasionally
-read while writing.  One possibility is to introduce metadata
-servers which can synchronously maintain an aggregated index in memory.
+some applications may need to make minor changes.  Many users find it most
+convenient to stage their input data sets and their executable code in the same
+directory where they write their checkpoint files.  Since \plfs\ is
+specifically a checkpoint file system and is not a general purpose file system,
+this usage pattern may suffer a performance hit.  These users should instead
+stage their input data sets and their executable code in a general purpose file
+system and merely direct their N-1 checkpoint files to a \plfs\ directory.
+
+Also, applications that open \plfs\ files in read-write mode will find that
+reading from these files can be very slow.  As we discussed in
+Section~\ref{arch-read}, reads in \plfs\ are handled by reading and aggregating
+the index files.  When files are opened in write-read mode, this process must
+be repeated for every \syscall{read} since intervening \syscall{writes} may
+have occurred.  Although the vast majority of HPC applications are able to read
+in read-only mode, we do plan to investigate removing this limitation in the
+future for those few applications who must occasionally read while writing.
+One possibility is to introduce metadata servers which can synchronously
+maintain an aggregated index in memory.
 

### 2009-07-29 SC09/intro.tex
got the graphs looking good. got the abstract right. got the keywords, crap, authors, copyright.
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@7 979fb63a-54b4-4514-8d67-9ab8ea1b59df
attached is the most recent version of the paper.  I'll do another big push tomorrow and then I'll submit it since I'll be on vacation all next week.  I think I've addressed most of the reviewer's comments but here's a couple that maybe still need to be addressed (or ignored):
the addition of metadata benchmark such as mdtest would make paper
stronger

the paper does not mention about file locks.  f

@@
 \section{Introduction}
+
+\input{summary-fig}
 
@@
 number for a supercomputer of its class.  Unfortunately, even at this size,
-component failures are a frequent event.  These failures are particularly
-problematic as many applications at LANL, and other High Performance Computing
-(HPC) sites, have long run times in excess of days, weeks, and even months.
+component failures are a frequent event.  For large parallel simulations at
+LANL, and other High Performance Computing (HPC) sites, which can run for
+several months, restarting a calculation from the beginning is not viable,
+since the mean time to interrupt for today's supercomputers is measured in days
+or hours and not months~\cite{google-failure,schroeder-failure}.
+
+\input{patterns-fig}
 
@@
 
-\input{summary-fig}
-
 As we shall discuss, and is summarized in Figure~\ref{fig-page1}, PLFS is

### 2009-07-29 SC09/motivation.tex
attached is the most recent version of the paper.  I'll do another big push tomorrow and then I'll submit it since I'll be on vacation all next week.  I think I've addressed most of the reviewer's comments but here's a couple that maybe still need to be addressed (or ignored):
the addition of metadata benchmark such as mdtest would make paper
stronger

the paper does not mention about file locks.  file system semantics is
not clear.

there should be a discussion about file system approach vs library
approach.  library can speed up N-1 w/out sacrificing performance for
other ops like metadata q

@@
 % Why do we care about checkpointing
-For large parallel simulations which can run for several months, restarting a
-calculation from the beginning is not viable, since the mean time to interrupt
-for today's supercomputers is measured in days or hours and not
-months~\cite{google-failure,schroeder-failure}.  As a result, applications 
+\if 0
+As a result, applications 
 periodically save
@@
 checkpoint. 
+\fi
 
@@
 
-\input{patterns-fig}
 \scribble{Should probably explain or remove 'false sharing' in figure}
@@
 
-Figure~\ref{fig-motivation} presents experimental data validating this discrepancy: An N-1 checkpoint pattern receives
-only a small fraction of the bandwidth achieved by an N-N pattern on PanFS,
-GPFS, and Lustre and does not scale with increased numbers of nodes.  The 
-PanFS experiment, run on LANL's Roadrunner
-supercomputer, shows a maximum observed N-N bandwidth of 31~\GBs\ compared
-to a maximum observed N-1 bandwidth of less than 100~\MBs. 
-Although we show PanFS results
-using their default RAID-5 configuration, PanFS has also a RAID-10
-configuration which reduces the implicit sharing caused by N-1 patterns when
-two writers both need to update the same parity.  While this solution improves
-scaling and offers much higher N-1 write bandwidth without sacrificing
-reliability, it does so by writing every byte twice, a scheme that, at best,
-can achieve only approximately half of the write bandwidth of N-N on RAID-5.
-\plfs, however, as will be shown in Section~\ref{eval}, can get much closer.
+\input{motivation-graphs}
+
+Figure~\ref{fig-motivation} presents experimental data validating this
+discrepancy: An N-1 checkpoint pattern receives only a small fraction of the
+bandwidth achieved by an N-N pattern on PanFS, GPFS, and Lustre and does not
+scale with increased numbers of nodes.  The PanFS experiment, run on LANL's
+Roadrunner supercomputer using its 1000 blade PanFS storage system, shows a
+maximum observed N-N bandwidth of 31~\GBs\ compared to a maximum observed N-1
+bandwidth of less than 100~\MBs.  Although we show PanFS results using their
+default RAID-5 configuration, PanFS has also a RAID-10 configuration which
+reduces the implicit sharing caused by N-1 patterns when two writers both need
+to update the same parity.  While this solution improves scaling and offers
+much higher N-1 write bandwidth without sacrificing reliability, it does so by
+writing every byte twice, a scheme that, at best, can achieve only
+approximately half of the write bandwidth of N-N on RAID-5.  \plfs, however, as
+will be shown in Section~\ref{eval}, can get much closer.
 
@@
 file systems exhibit similar behavior; N-N bandwidths are consistently higher
-than N-1 by at least an order of magnitude.
+than N-1 by at least an order of magnitude.  Measurements were gathered using
+the LANL synthetic checkpoint tool, \Term{MPI-IO Test}~\cite{mpi-io-test}.
+For each of these graphs, the size of each \syscall{write} was
+47001 bytes (a small, unaligned number observed in actual applications to be
+particularly problematic for file systems). \syscall{Writes} were
+issued until two minutes had elapsed. Although this is atypical since 
+applications tend to write a fixed amount of data instead of writing for
+a fixed amount of time, we have observed that this allows representative 
+bandwidth measurements with a predictable runtime.
 
@@
 single file is more complicated.
-
-\input{motivation-graphs}
 

### 2009-07-29 SC09/related-table.tex
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@5 979fb63a-54b4-4514-8d67-9ab8ea1b59df
attached is the most recent version of the paper.  I'll do another big push tomorrow and then I'll submit it since I'll be on vacation all next week.  I think I've addressed most of the reviewer's comments but here's a couple that maybe still need to be addressed (or ignored):
the addition of metadata benchmark such as mdtest would make paper
stronger

the paper does not mention about file locks.  file system semantics is
not clear.

there should be a discussion about file system approach vs li

@@
 \hyphenation{Immediately}
-\begin{table}
+\begin{table*}
 \begin{center}
@@
 \tabularnewline \hline
+Sp Writ &  \C{Library} &\C{Yes} & \C{Yes} & \C{Yes} & \C{No} & \C{No} 
+\tabularnewline \hline
 ZEST &  \C{FUSE} &\C{No (RD)} & \C{No (RD)} & \C{No} & \C{No} & \C{No} 
@@
 Diskless for Diskless 
-Checkpointing, LD for local disk on the compute nodes, RD for remote disk 
+Checkpointing, Sp Writ for Lustre Split Writing, LD for local disk on the compute nodes, RD for remote disk 
 on the storage system, M for memory, and N for network.  
@@
 \end{center}
-\end{table}
+\end{table*}
 

### 2009-07-29 SC09/related.tex
attached is the most recent version of the paper.  I'll do another big push tomorrow and then I'll submit it since I'll be on vacation all next week.  I think I've addressed most of the reviewer's comments but here's a couple that maybe still need to be addressed (or ignored):
the addition of metadata benchmark such as mdtest would make paper
stronger

the paper does not mention about file locks.  file system semantics is
not clear.

there should be a discussion about file system approach vs library
approach.  library can speed up N-1 w/out sacrificing performance for
other ops like metadata q

@@
 requires application modification, it enables interoperability with other
-middleware libraries.
+middleware libraries.  Similarly, Lustre Split Writing~\cite{lsw}, uses a 
+library approach and leverages Lustre's file joining mechanism to decouple
+concurrent access at runtime as does \plfs.  However, Lustre Split Writing is
+wed to the Lustre file system, requires application modification, and prevents
+the data from being immediately available following a checkpoint.
 

### 2009-07-30 SC09/arch.tex
attached is the most recent version of the paper.  I'll do another big push tomorrow and then I'll submit it since I'll be on vacation all next week.  I think I've addressed most of the reviewer's comments but here's a couple that maybe still need to be addressed (or ignored):
the addition of metadata benchmark such as mdtest would make paper
stronger

the paper does not mention about file locks.  file system semantics is
not clear.

there should be a discussion about file system approach vs library
approach.  library can speed up N-1 w/out sacrificing performance for
other ops like metadata q

@@
 
-\plfs\ is a virtual file system situated between the parallel application and
-an \upfs\ responsible for the actual data storage. As \plfs\ is a virtual file
+\plfs\ is a virtual FUSE file system, mounted on the compute nodes, situated between the parallel application and
+an \upfs\ responsible for the actual data storage. 
+As \plfs\ is a virtual file
 system, it leverages many of the services provided by the \upfs\ such as
@@
 then construct a \Term{global index} by aggregating the multiple index files 
-into an offset lookup table.
+into an offset lookup table.  This global offset is constructed as needed to
+satisfy read operations and lookup operations when cached metadata is not
+available (as will be discussed in Section~\ref{arch-meta}).
 

### 2009-07-30 SC09/abstract.tex
final version prepared by John Bent.
Milo it's all yours!



git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@10 979fb63a-54b4-4514-8d67-9ab8ea1b59df

@@
 reduce checkpoint time by an order of magnitude for several important
-benchmarks and real applications.  
+benchmarks and real applications without any application modification.  
 } 

### 2009-07-30 SC09/arch.tex
final version prepared by John Bent.
Milo it's all yours!



git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@10 979fb63a-54b4-4514-8d67-9ab8ea1b59df

@@
 
-One difficultly in constructing this global index stems from concurrent
-processes that may write the same offset at the same time.  These processes
+One difficulty in constructing the global index stems from multiple
+processes that may write the same offset concurrently.  These processes
 cannot know which will be the ultimate writer, so they need to synchronize to
@@
 be sufficient. In practice checkpoints do not experience overlapping writes, so
-at this time \plfs\ has not implemented an overlap conflict resolution scheme.
+at this time \plfs\ has not implemented overlap resolution.
 
@@
 sixteen processes on a node; on the next LANL supercomputer, this could be up
-to sixty-four. We tried reducing the number of data files in the same manner.
+to sixty-four. We tried reducing the number of data files similarly.
 Write bandwidth was not affected, but reads were slowed for \Term{uniform
@@
 differences between the write and read phases, multiple processes in a uniform
-restart may not always read from a shared file in sequential order.
+restart may not always read from a shared file sequentially.
 
@@
 the same logical name as the \plfs\ file. This allows \plfs\ to pass a
-\syscall{readdir} system call directly to the \upfs\ and return its result to
-the application without any translation.  \plfs\ also handles \syscall{mkdir}
-system calls in the same way (\ie\ without any translation). The implication of
-leveraging the \upfs\ in this way is that \plfs\ requires some other mechanism
-by which to distinguish between regular directories and containers in order to
-implement the \syscall{stat} system call correctly. As the \textit{SUID} bit is
-rarely used on directories and yet is allowed to be set by the underlying file
-system, \plfs\ sets this bit on containers; this does mean, however, that
-\plfs\ must disallow setting this bit on a regular directory.
+\syscall{readdir} system call directly to the \upfs\ and return its result
+without any translation.  \plfs\ also handles \syscall{mkdir} system calls
+without any translation.  Leveraging the \upfs\ in this way requires that
+\plfs\ use some other mechanism to distinguish between regular directories and
+containers in order to implement the \syscall{stat} system call correctly. As
+the \textit{SUID} bit is rarely used on directories and yet is allowed to be
+set by the underlying file system, \plfs\ sets this bit on containers; the
+implication however is that \plfs\ must disallow setting this bit on a regular
+directory.
 
@@
 Computing these sums and maximums with every \syscall{stat} call on a \plfs\
-file is expensive. Our strategy for speeding this up is to cache recently
+file is expensive. Our strategy for quickening this is to cache recently
 computed values in the \Term{metadata subdirectory}. To make this cache as
-effective as possible we have each FUSE process cache into this metadata
-subdirectory any information it has in its in memory data structures when the
-last writer on that node closes the file. On this close, the
-FUSE process creates a file named H.L.B.T, where H is the node's hostname, L is
-the last offset recorded on that node, B is the sum of the capacity of all
-files that this node manages, and T is the maximum timestamp among these files.
+effective as possible we have each \plfs\ server cache into this metadata
+subdirectory any information it has in its memory data structures when the last
+writer on that node closes the file. On this close, the FUSE process creates a
+file named \Term{H.L.B.T}, where \Term{H} is the node's hostname, \Term{L} is
+the last offset recorded on that node, \Term{B} is the sum of the capacity of
+all files that this node manages, and \Term{T} is the maximum timestamp among
+these files.
 
 When no process has this container open for write, a stat call on the container
-is implemented by issuing a \syscall{readdir} on the metadata subdirectory,
-then reporting the maximum of the Ls for last byte offset, the maximum of the
-Ts for modification timestamp and the sum of the Bs for capacity.
+issues a \syscall{readdir} on the metadata subdirectory,
+then reports the maximum of the offsets for last byte offset, the maximum of
+the timestamps for modification timestamp and the sum of the blocks for
+capacity.
 

### 2009-07-30 SC09/eval.tex
final version prepared by John Bent.
Milo it's all yours!



git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@10 979fb63a-54b4-4514-8d67-9ab8ea1b59df

@@
 Eleven of these twelve graphs present one experiment each. The twelfth,
-Figure~\ref{eval:summary}, presents a summary. In the majority of these graphs,
-the write bandwidth is shown on the \yaxis\ in \MBs\ as a function of the
-number of processes. We will note it in the text for those few graphs for
-which we deviate from this general configuration. The write bandwidth that we
-report is whatever is reported by the particular benchmark; whenever possible,
-we report the most conservative value (\ie\ we include open and close times in
-our write times, and we either barrier after close or we use the time reported
-by the slowest writer). Finally we have attempted to run multiple iterations
-for each experiment; where applicable, the standard deviation is therefore
-included.
+Figure~\ref{eval:summary}, is a summary. In the majority of these graphs, the
+write bandwidth is shown on the \yaxis\ in \MBs\ as a function of the number of
+processes. We will note it in the text for those few graphs for which we
+deviate from this general configuration. The write bandwidth that we report is
+whatever is reported by the particular benchmark; whenever possible, we report
+the most conservative value (\ie\ we include open and close times in our write
+times, and we either barrier after close or we use the time reported by the
+slowest writer). Benchmarks were run on LANL's Roadrunner supercomputer using
+the default parameters unless indicated otherwise.  Finally we have attempted
+to run multiple iterations for each experiment; where applicable, the standard
+deviation is included.
 
@@
 only. Secondly, the structure of the \plfs\ container spreads data across
-multiple sub-directories within a top-level directory whereas a typical N-N
+multiple sub-directories whereas a typical N-N
 pattern confines all N files into only a single parent directory.
@@
 larger, contiguous writes more friendly to the \upfs. Therefore we compare the
-bulkio method writing directly to PanFS to the old MPI-IO method writing
-indirectly to PanFS via \plfs.  This data was collected on Roadrunner.  The
-next two graphs show similar results; using \plfs\ to rearrange an N-1 pattern
-yields significantly improved bandwidths.  Figure~\ref{eval:lanl2} was run on a
-small \rrz\ which consists of 128 8-core compute nodes with a 66 blade PanFS
-storage system, whereas Figure~\ref{eval:lanl3} was run on Roadrunner itself.
-Note that these are extremely important applications at LANL; it has been
-estimated that they account for more than half of the total computing cycles at
-LANL.  Although these applications must remain anonymous, traces of their IO
-are available~\cite{plfs-maps}.
+bulkio method writing directly to PanFS to the MPI-IO method writing indirectly
+to PanFS via \plfs.  This data was collected on Roadrunner.  The next two
+graphs show similar results; using \plfs\ to rearrange an N-1 pattern yields
+significantly improved bandwidths.  Figure~\ref{eval:lanl2} was run on a small
+\rrz\ which consists of 128 8-core compute nodes with a 66 blade PanFS storage
+system, whereas Figure~\ref{eval:lanl3} was run on Roadrunner itself.  Note
+that these are extremely important applications at LANL, consuming more than
+half of LANL's total cycles.  Although they must remain anonymous, traces of
+their IO are available~\cite{plfs-maps}.
 
@@
 \input{overhead-fig}
+\input{read-graphs.tex}
 
@@
 \subsubsection{Read Bandwidth}
-\input{read-graphs.tex}
 
@@
 
-We measured the time to read these files using three different read
-scenarios. One emulates a \Term{uniform restart} in which the number of
-processes resuming the application is the same as the number that wrote the
-last checkpoint. In this case, each process within the parallel job reads the
-same data in the same order as one of the previous writers. In
-contrast, a \Term{non-uniform restart} is one in which the number of
-readers is different from the number of writers and the read offsets are
-therefore not aligned with the previous writes. In this case, we emulate the
-scenario where an application resumes after a failure by simply running on the
-newly reduced number of nodes. The number of reads is not affected by the size
-of the job for typical N-1 checkpoints. Each read is extracting a region for a
-particular variable within the simulation. The size of this region depends on
-the number of processes within the job. Each gets $1/Nth$ of the region.
-Specifically, for a non-uniform restart in which the number of processes, $N$,
-has been reduced by $M$, the size of each read will be $N/(N-M)$ times the size
-of the original writes. The third read scenario we emulated is a typical
-scenario in which a relatively, small fixed number of processes reads a
-checkpoint in order to save it onto an archive system. For these measurements,
-we used LANL's archive copy utility with fours readers each running on their
-own node; each reader read just one contiguous 5~\GB\ region by issuing 
-sequential 1~\MB\ reads.
+We measured the time to read these files using three different read scenarios.
+One emulates a \Term{uniform restart} in which the number of processes resuming
+the application is the same as the number that wrote the last checkpoint. In
+this case, each process reads the same data in the same order as one of the
+previous writers. In contrast, a \Term{non-uniform restart} is one in which the
+number of readers is different from the number of writers and the read offsets
+are therefore not aligned with the previous writes. In this case, we emulate
+the scenario where an application resumes after a failure by simply running on
+the newly reduced number of nodes. The {\em{number}} of reads is not affected
+by the size of the job for typical N-1 checkpoints since each read is
+extracting a region for a particular variable within the simulation. However,
+the {\em{size}} of this region depends on the number of processes within the
+job since each gets $1/Nth$ of the region.  Specifically, for a non-uniform
+restart in which the number of processes, $N$, has been reduced by $M$, the
+size of each read will be $N/(N-M)$ times the size of the original writes. The
+third scenario emulated is where a relatively, small fixed number of processes
+reads a checkpoint in order to save it onto an archive system. For these
+measurements, we used LANL's archive copy utility with fours readers each
+running on their own node; each reader reads a single contiguous 5~\GB\ region
+by issuing sequential 1~\MB\ reads.
 

### 2009-07-30 SC09/future.tex
final version prepared by John Bent.
Milo it's all yours!



git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@10 979fb63a-54b4-4514-8d67-9ab8ea1b59df

@@
 applications have begun using it to improve their N-1 checkpoint performance. 
+\plfs\ is publically available at
+http://sourceforge.net/projects/plfs.
 
@@
 stage their input data sets and their executable code in a general purpose file
-system and merely direct their N-1 checkpoint files to a \plfs\ directory.
+system and merely direct their N-1 checkpoint files to \plfs.
 

### 2009-07-30 SC09/intro.tex
final version prepared by John Bent.
Milo it's all yours!



git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@10 979fb63a-54b4-4514-8d67-9ab8ea1b59df

@@
 number for a supercomputer of its class.  Unfortunately, even at this size,
-component failures are a frequent event.  For large parallel simulations at
+component failures are frequent.  For large parallel simulations at
 LANL, and other High Performance Computing (HPC) sites, which can run for
@@
 chosen by the application. 
-
-% our approach
 In this paper, we describe different checkpointing patterns, and show how some
 result in very poor storage performance on three of the major HPC
-parallel file systems: PanFS, GPFS, and Lustre.  We then posit that an
+parallel file systems: PanFS, GPFS, and Lustre.
+
+We then posit that an
 interposition layer inserted into the existing storage stack can rearrange this
@@
 \upfs.  To test this hypothesis, we have developed \plfs, a Parallel
-Log-structured File System, which is one such interposition layer.
-We present measurements using \plfs\ on several synthetic benchmarks and real applications at multiple HPC supercomputing centers.
-The
+Log-structured File System, one such interposition layer.  We present measurements using \plfs\ on several synthetic benchmarks and real applications at multiple HPC supercomputing centers.  The
 results confirm our hypothesis: writing to the \upfs\ through
@@
 
-As we shall discuss, and is summarized in Figure~\ref{fig-page1}, PLFS is
+As we shall discuss, and is summarized in Figure~\ref{fig-page1}, \plfs\ is
 already showing very large speed ups for widely used HPC benchmarks and
-important real HPC codes, and at extreme scale.
+important real HPC codes, and at extreme scale.  \plfs\ works on unmodified
+applications as an optional mount point on the compute nodes and is designed to
+augment, and not replace, an existing \upfs.  \plfs\ consists of three thousand
+lines of code and is publically available at
+http://sourceforge.net/projects/plfs.
 

### 2009-07-30 SC09/motivation.tex
final version prepared by John Bent.
Milo it's all yours!



git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@10 979fb63a-54b4-4514-8d67-9ab8ea1b59df

@@
 of compute nodes is easier to code as the checkpoint format is independent of
-the number of processes that captured the checkpoint; conversely, gathering the
+the number of processes that wrote the checkpoint; conversely, gathering the
 appropriate regions from multiple files or from multiple regions within a
@@
 LANL's N-1 applications is a twenty-year old Fortran checkpointing library.
-About a decade ago, in response to a growing clamor about the
-limitations of N-1 checkpointing bandwidth, developers for this application
-augmented their checkpointing library with fifteen thousand lines of code.
-However, instead of changing the application to write an N-N pattern, they
-added a prefer to the IO routiones in which interprocess communication is used to aggregate and buffer writes.
+About a decade ago, in response to a growing clamor about the limitations of
+N-1 checkpointing bandwidth, developers for this application augmented their
+checkpointing library with fifteen thousand lines of code.  However, instead of
+changing the application to write an N-N pattern, they added a new IO routine in
+which interprocess communication is used to aggregate and buffer writes.
 Although they did not improve the performance to match that of other
-applications using an N-N checkpoint pattern, 
-this effort was considered a success as they did improve the N-1
-performance by a factor of two to three. This new checkpointing library,
-called \Term{bulkio}, has been maintained over the past decade and ported to
-each new successive supercomputer at LANL~\cite{bent-personal-bulk}. 
-Furthermore, N-1 patterns continue to be developed anew in many new
-applications.  High level data formatting libraries such as Parallel
-NetCDF~\cite{pnetcdf} and HDF5~\cite{hdf5} offer convenience to application
-developers who simply describes the logical format of their data and need no
-longer consider how that data is physically organized in the file system.  Once
-again this convenience merely shifts the complexity to the \upfs\ 
+applications using an N-N checkpoint pattern, this effort was considered a
+success as they did improve the N-1 performance by a factor of two to three.
+This checkpointing library, called \Term{bulkio}, has been maintained over
+the past decade and ported to each new successive supercomputer at
+LANL~\cite{bent-personal-bulk}.  Furthermore, N-1 patterns continue to be
+developed anew in many new applications.  High level data formatting libraries
+such as Parallel NetCDF~\cite{pnetcdf} and HDF5~\cite{hdf5} offer convenience
+to application developers who simply describes the logical format of their data
+and need no longer consider how that data is physically organized in the file
+system.  Once again this convenience merely shifts the complexity to the \upfs\
 since these libraries use an N-1 pattern.

### 2010-01-28 fast10wip/FAST_WIP.tex
Adding draft of FAST 10 WIP
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@69 979fb63a-54b4-4514-8d67-9ab8ea1b59df

\documentclass[letterpaper,10pt]{article}
%\pagestyle{empty}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{authblk}
\usepackage[top=0.75in, bottom=0.75in, left=0.75in, right=0.75in]{geometry}

%\setlength\textheight{9.0in}

\title{Enabling Scientific Application I/O on Cloud FileSystems}

\author[1]{Milo Polte (student)}
\author[2]{Esteban Molina-Estolano (student)}
\author[3]{John Bent}
\author[2]{Scott Brandt}
\author[1]{Garth Gibson}
\author[4]{Maya Gokhale}
\author[2]{Carlos Maltzahn}
\author[3]{Meghan Wingate}
\affil[1]{Carnegie Mellon University}
\affil[2]{UC Santa Cruz}
\affil[3]{Los Alamos National Laboratory}
\affil[4]{Lawrence Livermore National Laboratory}

\date{}

\begin{document}

\bibliographystyle{plain}

\maketitle

    Today, the storage of many of the world's largest computing
    clusters is being supported not through traditional parallel
    filesystems, such as Lustre\cite{schwan:linux03}, the Parallel Virtual
    Filesystem (PVFS)\cite{carns:PVFS}, or PanFS\cite{nagle:sc04}, but
    through ``cloud'' filesystems such as GoogleFS\cite{ghemawat:gfs},
    the Hadoop Distributed Filesystem (HDFS)\cite{hdfs}, and the
    Amazon Simple Storage Service (Amazon S3).\cite{amazons3} Although
    these filesystems provide appropriate performance and resilience
    for web apps, their APIs are often unsuitable for the I/O patterns
    of a scientific application. Scientific applications typically use
    a POSIX interface, often with additional support for semantics
    such as concurrent writers or out of order writes. These
    applications cannot presently run at all on a cloud filesystem,
    since these filesystems lack the necessary semantics. We would
    like to remedy this deficiency: running scientific applications on
    cloud filesystems would allow sites with investments in cloud
    filesystems to leverage existing hardware to run scientific
    computing applications as well.

    Previously, the authors have designed a thin interposition layer,
    the Parallel Log-structured Filesystem (PLFS)\cite{plfs}, to
    improve the I/O speed of scientific applications running on a
    parallel filesystem. PLFS transparently decouples inefficient,
    concurrent writers accessing a shared file into efficient,
    log-style writers maintaining individual data files, while still
    providing the application with a view of a single, flat file.

    Although this previously work was motivated by performance rather
    than enabling functionality, we believe the same techniques can be
    adapted to support concurrent writers in a cloud filesystem. In
    this work, we will extend PLFS to run on the Hadoop filesystem and
    add shared write semantics to it. This will enable scientific
    applications, such as simulation checkpointing, to run unmodified
    on top of the Hadoop filesystem. We plan to compare the
    performance against the same scientific apps running on PVFS (a
    common parallel filesystem used at National Labs) on the same
    hardware and evaluate the trade-offs in performance.

\bibliography{references}

\end{document}

### 2010-04-29 devel_path/abstract.tex
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@154 979fb63a-54b4-4514-8d67-9ab8ea1b59df

\begin{abstract}
Experiment management requires a constant feedback loop cycling through
planning, execution, measurement, and analysis.  Experiment management in High
Performance Computing (HPC) follows this general pattern with three additional
characteristics. One, HPC applications must deal with frequent platform
failures which can interrupt, perturb, or terminate experiments. Two, these
applications typically use MPI.  Three, a scheduling system typically acts as a
gate-keeper.  This paper introduces \namefull\ (\name), an experimental
management framework for HPC which simplifies all four phases of experiment
management. Planning is simplified by allowing the user to merely describe
their experimental goals instead of constructing parameters for each individual
task. To simplify execution, \name\ dispatches the tasks itself thereby freeing
the user from remembering the often arcane methods for interacting with various
scheduling systems. \name\ provides transducers that automatically measure and
record important information; these can be extended to collect measurements
specific to each experiment.  Finally, analysis is simplified by providing 
\dbviz, a visualization tool enabling interactive data exploration. 
\end{abstract}

### 2010-04-29 devel_path/arch-fig.tex
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@154 979fb63a-54b4-4514-8d67-9ab8ea1b59df

\begin{figure*}[tb]
    \centering
    \subfloat[\label{typical}Typical Experiment Lifecycle]{
        \includegraphics[width=0.4\textwidth]{figs/planning.pdf}
    }
    \hspace{0.5in}
    \subfloat[\label{lem-plan}\name\ Augmented Experiment Lifecycle]{
        \includegraphics[width=0.4\textwidth]{figs/arch.pdf}
    }
    \mycaption{fig-plan}{Experiment Lifecycle.}{
The figure on the left depicts a typical experiment lifecycle progressing
through planning, execution, measurement, and analysis.  A user devises an
experiment to study the effect of some number of independent variables,
executes and measures each \sub, analyzes the results, and, depending on that
analysis, can then revise the experiment and repeat the cycle.  The figure on
the right shows the workflow within \name.  The user describes their parameter
space to the \cg\ and specifies a path to an optional application transducer to
capture application specific \kv\ pairs.  \name\ adds a path to a general
transducer to capture \kv\ pairs common to all experiments.  The generated
commands and their transducer(s) are then passed to the \dispatcher\ which
either submits the commands to a scheduling system or runs them synchronously
in the foreground.  As the commands run, \kv\ pairs are captured by the
transducer(s) and inserted into a SQL server from which they can then be
queried to visualize and analyze the results.  Depending on this analysis the
user can then modify their parameter space and repeat the cycle.
}
\end{figure*}

### 2010-04-29 devel_path/arch.tex
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@154 979fb63a-54b4-4514-8d67-9ab8ea1b59df

\section{\name\ Architecture}
\label{arch}

Simplifying the design, planning, execution, and analysis of experiments, \name\
is comprised of four main parts as shown in Figure~\ref{lem-plan}.  A command
generator reads a simple experiment description and produces a command to run
each \sub.  The \subs\ are then passed to the\ \dispatcher\ who is responsible
for dispatching each.  As each \sub\ is executed, its output is parsed by
\Term{transducers} who insert parsed \kv\ pairs into a SQL server.  The user
then uses DBViz, which queries the SQL server, to visualize and analyze their
results.  Depending on this analysis, the user can modify their experiment
description and repeat the cycle.

\subsection{Command Generation}

\name\ provides a \cg\ engine which reads a description of the experiment,
computes all \subs\ needed to complete the experiment, and
produces a command to execute each \sub.  
For example, for an experiment wishing to study the
effect of a set of independent variables, the user merely specifies an array of
every target value for each independent variable.  If multiple independent
variables are specified, the \cg\ produces every possible combination of
commands.  The \cg\ assumes that each \sub\ will be executed with 
\Term{mpirun}, although this can be over-ridden.  The generated commands are 
then passed to the \dispatcher\ who initially just lists the commands so that 
the user can verify that the commands have been correctly generated; once 
verified, the \dispatcher\ can either submit the commands to the scheduling 
system or run them synchronously in the foreground.

\input{cg-example}

The simplest approach is to write a text file with one mpirun command in each
line; in this case, the \cg\ will pass each command directly to the
\dispatcher\ without modification.  For more complicated parameter studies,
\name\ provides a template where multiple list of parameters can be specified.
Figure~\ref{mdtest-desc} shows an example of an experiment description file for
the \Term{mdtest} parallel metadata benchmark.  This example shows how the
optional \Term{mpi\_options}, \Term{program\_options}, \Term{transducers}, and
\Term{transducer\_arguments} are used.  The command generator then creates all
possible commands from the permutations of the lists of options and arguments
as is shown in Figure~\ref{mdtest-commands}.  The \Term{--desc} argument is
parsed by the transducer, is removed from the arguments passed to
\Term{mdtest}, and is used as a tag for the experiment to easily group all of
the \subs\ for subsequent analysis.

Although the \cg\ can read from a simple list of mpirun commands, for any
reasonable parameter sweep, a description file is a drastic simplification.
Although this description file must be written in \Term{Python}, the basic
structure is already provided to the user in a template description file.  
So far two different users lacking any \Term{Python} experience have used
\name\ and were able to edit the provided template and start executing their
experiments within an hour.

\subsection{\Dispatcher}

\name\ provides an \manager\ program \Term{run\_expr.py}.  To run an experiment,
the user executes the \manager\ by passing it either a description file or a
text list of commands.  After invoking the \cg\ to obtain a list of commands
from the file provided by the user, the \manager\ then passes these commands to
the \dispatcher.  The \manager\ program takes several command line options
which it passes to the \dispatcher such as the mode of dispatch, the number of
times to run each command, and whether to randomly shuffle the commands before
dispatching them.  The dispatch mode specifies how the \dispatcher\ should
dispatch each command; by default it merely lists the generated commands so
that the user can verify that the parameter study has been correctly configured
in the description file.  After verifying the list of commands, the user can
ask the \dispatcher\ (with command line options passed to the \manager) to run
the jobs synchronously in the foreground or to submit them to a scheduling
system.

Additional command line options can be used to pass additional information to
the scheduling system.  Currently, \name\ only supports the \Term{moab}
scheduler although it can easily be modified to support other schedulers as
well; an earlier version of \name\ had support for both \Term{moab} and
\Term{PBS}.

The \dispatcher\ also pulls default parameters from a configuration file which
is also written in Python.  This configuration file provides different
scheduling system parameters for different supercomputers.  For example, at
LANL we have populated our \name\ configuration file with information about the
number of cores on each node in our different supercomputers.  This information
is then passed to the scheduling system which needs to know how many nodes to
allocate for different numbers of requested processes (\eg\ to dispatch a 128
processor job on the LANL RRZ supercomputer, the \dispatcher\ reads the
configuration file to discover that RRZ has eight cores per node and therefore
requests sixteen nodes from the scheduling system).  We also include
information about the default filesystem for each supercomputer; although the
naming conventions on each supercomputer are mostly consistent, some
supercomputers do have different paths to their parallel storage systems.  By
specifying a default value for each supercomputer, we can use a single
experiment description file across multiple supercomputers without modifying it
specifically for each.

Some of the functionality of the \dispatcher\ might seem redundant since users
can access scheduling systems directly; however, we believe that this
redundancy is actually one contribution of \name\ in that it abstracts the
scheduling system away.  Users create an experiment description file, and a
configuration file, and can then dispatch commands on any scheduling system on
any supercomputer without having to modify any of their scripts
since \name\ transparently interacts with the scheduling system for them.

\subsection{Transducers}

As command generation and dispatch assist the planning and execution phases of
an experiment \lifecycle, so do the \name\ \Term{transducers} assist in the
measurement phase.  The term transducers is borrowed from Semantic File Systems
(SFS)~\cite{gifford-sfs}.  Just as transducers in SFS create searchable
semantic information about files, transducers in \name\ create searchable
semantic information about each \sub.  \name\ provides a general transducer
which collects information about every \sub; this transducer can be augmented
by the user to collect information specific to the application.  As is shown in
Figure~\ref{mdtest-commands}, the transducer parses the output of each \sub,
records the command line of each, collects additional information about the
execution environment, and then inserts all of this collected information as
\kv\ pairs into a SQL server.  Some of the \kv\ pairs collected from each \sub\
are as follows:

\begin{quote}
\begin{itemize}
\mydefinition{application}{Name of the application}
\mydefinition{commandline}{Command line used}
\mydefinition{description}{Optional experiment tag}
\mydefinition{epoch}{UNIX epoch at time of execution}
\mydefinition{hostlist}{A list of each compute node}
\mydefinition{numhosts}{Number of compute nodes used}
\mydefinition{retval}{Return value}
\mydefinition{system}{Name of the supercomputer}
\mydefinition{username}{Username of the submitting user}
\mydefinition{walltime}{Total runtime}
\end{itemize}
\end{quote}

The provided transducer can be easily extended by users comfortable with
regular expressions who can parse the application output to gather additional
\kv\ pairs specific to their applications.  For example, our transducer for
\Term{mdtest} requires only 200 lines to parse the output and collect
semantic information about the configuration and results for each \sub.

\subsection{\label{sec-analysis}Visualization and Analysis}

\begin{figure*}[tb]
\centering
\includegraphics[width=0.7\textwidth]{dbvizGraphsAndConfigs/screenShot.png}
\mycaption{dbviz-screen}{Interactive Data Exploration.}{This screenshot of
\dbviz\ shows how a graph is generated from a specified filter, \xaxis,
\yaxis\, and a set of columns to compare.}
\end{figure*}

To enable interactive data exploration, we have developed \dbviz, a data
analysis tool written in \Term{Java Swing}.  This tool pulls data from any
\Term{SQL} database and plots it according to the user's specifications.  The
resulting graph is interactive which allows the user to more closely analyze
the data.  After ascertaining the table to be queried, \dbviz\ fetches the
schema and uses it to populate pull-down menus which are used to specify which
fields to display on the axes and which fields to use to create comparisons
(\ie\ different lines in the resulting graph).  This eliminates the need for
the user to remember the exact schema.  Although the user can use the pull-down
menus as a shortcut, the user can also request that more complicated operations
be performed.  For example, an axis can be just a single field or it can be a
function of multiple fields and constants; this could be used for instance to
create a bandwidth by dividing a size field by a time field.  When the user
does create more complicated functions for the axes or the compare fields,
these are remembered persistently by \dbviz\ and are added to the pull-down
menus so that they need be created only once.

\input{dbviz-fig}

After specifying the columns that should be displayed on the \xaxis\ and the
\yaxis, a filter (\eg\ an \Term{SQL} \Term{WHERE} clause) can also be
specified.  The filter allows the user to be more selective about the data that
will be returned.  By specifying a set of columns to be compared, different
configurations can be compared.  Currently, the available formats of the plots
are linespoints, xy-scatter, and bar graphs as is illustrated in
Figure~\ref{fig-dbviz}.  \dbviz\ allows sessions to be saved so that 
interesting data explorations and interactive sessions can be interrupted
and resumed.  These saved sessions are stored in text files which allows them
to be easily shared between users.

There are two important features in \dbviz\ that make its interactivity so
valuable.  One is \Term{outlier analysis}; the other is \Term{hidden difference
search}.

\subsubsection{Outlier analysis}
Outlier analysis is available in xy-scatter plots; the user can click
on any point and \dbviz\ will transparently query the database and present
every \kv\ pair for that point thereby allowing the user to quickly discover
the particular data for any \sub; we have found this to be of particular value
for analyzing outliers.

\subsubsection{Hidden difference search}
Hidden difference search is another feature provided
by \dbviz\ to help the user refine their analysis.  Hidden differences are
caused by incompatible data being accidentally combined when the user specifies
an imprecise filter or set of comparison fields.  When the user suspects that
the visualization lacks sufficient precision, hidden difference search can
quickly identify suspect keys which need to be segregated or filtered to
increase analytical precision.

Both outlier analysis and
hidden difference search will be further illustrated by means of example in
Section~\ref{eval}.

\subsection{Monitoring and Replay}

By inserting executed commands into a \Term{SQL} server, \name\ allows
interactive data exploration as discussed above in Section~\ref{sec-analysis}.
In addition, \name\ provides two additional programs that utilize the
\Term{SQL} server.  \Term{replay\_expr.py} takes a \Term{SQL} filter (\ie\ a
WHERE statement), queries the database for all previously executed commands
matching that filter, and passes them to the \dispatcher.  This is frequently
utilized at our site to check differences in platform performance following a
configuration change such as an upgrade to the operating system or changing the
buffer size on a network switch.  In such cases, the platform administrators
will conduct a measurement, giving it a unique tag, before making the change;
after the change, they merely use \Term{replay\_expr.py} to rerun that same
experiment.  Following the replay, they can use \dbviz\ to check for changes in
performance.  Although we use \name\ at our site primarily for platform
measurement and observation, it is applicable to any HPC experiment.

A second tool that leverages the database is \Term{check\_expr.py} which takes
both a \Term{SQL} filter and an experiment description file.  By using the
command generator to generate the necessary \subs\ from the description file,
\Term{check\_expr.py} can then determine which have not yet been dispatched by
querying for each in the database.  The uncompleted \subs\ can then optionally
be passed to the \dispatcher\ to resubmit them.  Although this functionality
can also be achieved by querying the scheduling system, doing so within \name\
allows the user a single abstraction for interacting with multiple different
scheduling systems.  Additionally, we frequently encounter situations where the
scheduling systems' queues are purged by the administrators; in such cases,
querying the database is the simplest way to check progress of the experiment.
Finally, this approach drastically simplifies checking progress when the \subs\
have been spread across multiple computing sites, especially when each uses
different scheduling systems.

### 2010-04-29 devel_path/cg-example.tex
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@154 979fb63a-54b4-4514-8d67-9ab8ea1b59df

\begin{figure*}[tb]
\subfloat[Experiment Description]{\label{mdtest-desc}
\lstinputlisting{examples/mdtest.py}}
\hspace{20pt}
\subfloat[Generated Commands]{\label{mdtest-commands}
\lstinputlisting{examples/mdtest.out}}
\mycaption{mdtest}{Command Generation.}{
On the left is a simple experiment description for the parallel
metadata benchmark \Term{mdtest}.  Lists are specified 
the \Term{-l} and the \Term{-z} options whereas just a single value
is specified for the \Term{-d} and \Term{-b} options.  The options list for 
\Term{mpi}
specifies that each generated command should be run at different sizes 
(passed to \Term{mpirun} with the \Term{-np} option. 
This example also shows how transducers can be 
specified to the \cg.  The right side shows the commands generated
from this experiment description.  Note that the user can either create
an experiment description file as in the left column or can manually create
a list of commands like those on the right side.  The \cg\ will determine
whether it needs to automatically generate commands from a description file
or whether it merely reads the commands directly.  Once the \cg\ either
generates the commands or reads them in, it then passes them to the \dispatcher.
}
\end{figure*}

### 2010-04-29 devel_path/conclude.tex
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@154 979fb63a-54b4-4514-8d67-9ab8ea1b59df

\section{Conclusion}
\label{conclude}

Experiment management is a cumbersome process.  Users must plan, execute,
monitor, and analyze multiple \subs.  This process is often ad-hoc and
each new experiment requires the development of new experiment management
procedures.  In computational science, and High Performance Computing in
particular, developing a new experiment often requires that multiple new
scripts are written: some to generate multiple \subs, some to execute
them, and some to parse their outputs to enable analysis.

\name\ simplifies this process by providing a general and modular framework
by which users can describe their experiments and \name\ will automatically
generate \subs, execute them, and collect the resulting data.  Additionally,
\name\ provides a simple interface to augment the data collection and to
monitor and replay experiments.  A visualization component allows interactive
data exploration greatly simplifying analysis and experiment tuning.  
Production use at Los Alamos National Labs has demonstrated the value of \name\
in enabling end-to-end experiment management in HPC.

### 2010-04-29 devel_path/dbviz-fig.tex
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@154 979fb63a-54b4-4514-8d67-9ab8ea1b59df

\begin{figure*}[tb]
  \centering
  \subfloat[\label{nonrandom}Linespoints]{
    \includegraphics[width=0.3\textwidth,height=\figheight]{dbvizGraphsAndConfigs/allPoints.jpg}
  }
  \subfloat[\label{random}XY-scatter]{
    \includegraphics[width=0.3\textwidth,height=\figheight]{dbvizGraphsAndConfigs/allPointsAllScatter.jpg}
  }
  \subfloat[\label{random-full}Bars]{
    \includegraphics[width=0.3\textwidth,height=\figheight]{dbvizGraphsAndConfigs/allPointsBarGraph.jpg}
  }
  \mycaption{fig-dbviz}{Visualization Flexibility.}{
After \name\ populates a database with \kv\ pairs collected by transducers,
\dbviz\ allows interactive exploration of that data.  These figures 
show typical visualizations where a filter has been applied to the database,
and a visualization using the resulting rows is displayed.  One line (or bar) is
produced for each set of parameters that the user wishes to compare as a
function of specified values for both axes. The axes values can be a single
field or they can be arbitrarily complex combinations of multiple fields and
constants (\eg\ to show a bandwidth in GB/s, the user might specify that the
\yaxis\ should be a field dividing the number of bytes by a time and then 
dividing that again by a gigabyte).  The figure on the left shows an aggregation of
multiple points into mean values with standard deviations.  In the middle is
the same data in an xy-scatter plot; the scatter plot is useful because it allows
the user to easily identify outlying data.  The interactivity is vital here
as a user can click on an outlying data point and \dbviz\ will transparently
query the database and display a table of all \kv\ pairs for that outlier.
Finally, \dbviz\ can produce bar-graphs as well as in the figure on the right.
}
\end{figure*}

### 2010-04-29 devel_path/future.tex
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@154 979fb63a-54b4-4514-8d67-9ab8ea1b59df

\section{Current Status and Future Work}
\label{future}

\name\ is used primarily by system integrators at Los Alamos National Lab to
monitor several parallel storage systems on LANL supercomputers.  It has been a
valuable tool enabling much easier analysis of the impact of configuration
changes and software upgrades on these systems.  The ease by which experiments
can be replayed allows significantly improved consistency of experiments
enabling much cleaner \Term{apples-to-apples} comparisons from before and after
configuration changes.  \dbviz\ and the central SQL server simplify cross
system comparisons that were previously extremely manually intensive.
Additionally, \name\ is increasingly being adopted by other groups at LANL and
has been used by several vendors to benchmark new systems.

We are working with one code team in particular at LANL that has current
workflow inefficiencies in their regression testing that could be drastically
improved using \name.  They currently use an ad-hoc set of tools that has some
of \name's functionality but it is not well consolidated nor organized and
other functionality is entirely absent.  This team runs nightly, weekly, and
monthly tests where various metrics such as CPU utilization and I/O bandwidth
are measured.  The data is held in a database but the team lacks the capability
to visualize the data interactively.  When measured values fall outside of
predetermined thresholds, the team is alerted to this but lacks an easy method
to identify a root cause.  Interactive sessions with \dbviz\ would enable the
team to quickly isolate, identify, and analyze the particular \subs\ that fell
outside of the thresholds.

Their static visualization tools are very lacking in this regard; we have seen
instances in which they have missed tests of interest because their predefined
graphing utilities left some \subs\ literally off the graphs.  Also, they have
been plagued by an abundance of data which overwhelms their graphs with noise;
\dbviz\ can allow them to interactively filter and focus their analysis.
Finally, even when they have ultimately identified root cause for performance
anomalies, they lack utilities to rerun just those \subs.  The replay ability
of \name\ to apply a query to fetch a particular set of \subs\ to resubmit will
drastically improve their workflow and hasten their analysis of performance
anomalies.

\dbviz\ is publicly available through SourceForge; the other components of
\name\ will also be made publicly available pending the copyright process.

We are continuing to improve \name\ through the development of additional
features.  In the future, we aim to add a web-based interface to centralize
experiment management and more easily allow multiple users to interact with
collaborative experiments.  Additionally, we plan to modularize the database
integration to remove the SQL-only restriction.

### 2010-04-29 devel_path/intro.tex
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@154 979fb63a-54b4-4514-8d67-9ab8ea1b59df

\section{Introduction}
Experiment management in any domain is challenging.  There is a perpetual
feedback loop cycling through planning, execution, measurement, and analysis.
The lifetime of a particular experiment can be limited to a single cycle 
although many require myriad more cycles before definite results can be 
obtained.  Within each cycle, a large number of \Term{\subs} may be executed
in order to measure the effects of one or more independent variables.

Experiment management in high performance computing (HPC) follows this general
pattern but also has three unique characteristics.  One, computational science
applications running on large supercomputers must deal with frequent platform
failures which can interrupt, perturb, or terminate running experiments.  Two,
these applications typically integrate in parallel using MPI as their
communication medium; each \sub\ within HPC is typically executed with an
\Term{mpirun} command.  Three, there is typically a scheduling system (\eg\
Condor, Moab, SGE, \etc) acting as a gate-keeper for the HPC resources.

In this paper, we introduce \namefull\ (\name), an experimental management
framework simplifying all four phases of experiment management.  \name\
simplifies experiment planning by allowing the user to describe their
experimental goals without having to fully construct the individual parameters
for each task.  To simplify execution, \name\ dispatches the \subs\ itself
thereby freeing the user from remembering the often arcane methods for
interacting with the various scheduling systems.  By providing
\Term{transducers}, \name\ automatically measures and records important
information about each \sub; these transducers can easily be extended to
collect additional measurements specific to each experiment.  Transducers are
implemented as wrappers that execute each \sub\ and can be extended to parse
the output of each \sub;  the term is borrowed from a similar concept in
Semantic File Systems~\cite{gifford-sfs}.  In both \name\ and Semantic File
Systems, transducers automatically generate semantically meaningful information
in \kv\ pairs which can then be easily searched.  Finally, experiment analysis
is simplified by providing \dbviz\, a general database visualization framework
that allows users to quickly and easily interact with their measured data.
\dbviz\ provides two important advanced features for improving data analysis:
\Term{outlier analysis} and \Term{hidden difference search}.  A typical
experiment workflow is shown in Figure~\ref{typical}; Figure~\ref{lem-plan}
shows the same workflow augmented with the various components of \name.

Although our experiment management framework has been designed for use at Los
Alamos National Lab (LANL) and therefore generally for an HPC environment, we
believe it can be easily modified to be suitable for experiment management in a
Grid environment as well.  Throughout this paper, we discuss how \name\ reduces
the complexity of managing experiments that may span multiple supercomputers
and the different scheduling systems on each.  In terms of experiment
management, supercomputers in HPC are very similar to clusters in the Grid.
Although the Grid introduces additional complexity due to multiple
administrative domains, we believe that \name\ can be useful in this
environment as well but this is not a focus of this paper.

\input{arch-fig}

The remainder of the paper is organized as follows.  We describe our design in
Section~\ref{arch}.  In Section~\ref{eval}, we offer a demonstration of our
usage model.  We provide a comparison to related work in Section~\ref{related},
current status and future work in Section~\ref{future}, and a conclusion in
Section~\ref{conclude}.

### 2010-04-29 devel_path/plfs_devel.tex
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@154 979fb63a-54b4-4514-8d67-9ab8ea1b59df

%\documentclass[10pt,onecolumn]{IEEEtran}
\documentclass[10pt]{article}
\pagenumbering{arabic}
\pagestyle{plain}
%\pagestyle{empty}
\usepackage{fullpage}
\usepackage{setspace}
\usepackage[top=2cm,bottom=2cm,left=2cm,right=2cm]{geometry}
\usepackage{wrapfig}
\usepackage{graphicx}
\usepackage{sidecap}
\usepackage{subfig}
\usepackage{eso-pic,type1cm,color}
\usepackage{listings}
\lstset{numbers=left, numberstyle=\footnotesize, numbersep=2pt,language=Python,basicstyle=\footnotesize}

\input{macros}
\input{watermark}

\begin{document}

\title{\plfs\ Versioning Plan}
\if 0
\author{
    John Bent,
    Ryan Kroiss,
    Alfred Torrez,
    Meghan Wingate
    \\
    Los Alamos National Lab
}
\fi
\date{}
\maketitle
\thispagestyle{empty}
\pagestyle{empty}
\section{Preface}

This document attempts to lay out a design path for the planned development
versions of \plfs.  It also attempts to describe how development versions can
move into production and the various issues for transitioning between versions.
Some of the issues are technical and some are policy.  Some of the technical
issues are things like how do we implement directory management in the future
\mds.  Policy issues are questions like whether incompatible layouts are 
handled transparently by \plfs\ or whether they are handled by explicitly
copying data from an old \plfs\ into the new one.

Our goal is to get \plfs\ into production at LANL.  However, we want to make 
sure that the production releases are bug-free, give performance attractive
to users, and are as easy to use as possible.  Hopefully by spelling out the
issues involved the current version of \plfs\ and the planned versions, we
can identify the target version for our first production release.  There is
of course a tension between releasing something soon and releasing something
good.  We could release the current version today but it has several 
limitations that will be discussed below and we need to determine whether
we are comfortable with those limitations.

At the end of this document, we will also describe current open issues
and other miscellaneous development plans.

\section{Current Version - Version 0}

The current version has support for both \fuse\ and for \adio.  However,
the \fuse\ implementation transparently maps logical paths into physical
ones.  It knows this mapping via a single mount argument naming the
physical \store.  As is true for PanFS access, each user's \plfs\ \store\ 
is located on a different PanFS volume.  This is accomplished via symbolic
links.  For example, \Path{/plfs/ben} is a symbolic link to 
\Path{/mnt/plfs/vol5/ben/.plfs\_store}.  When \plfs\ is mounted, it is mounted 
on \Path{/mnt/plfs} and directed to use \Path{/panfs/scratch} as its \store.
So a typical name resolution would work as follows:

\begin{enumerate}
\item{user ben opens logical file \Path{/plfs/ben/foo}}
\item{the sym link resolves this to logical path 
        \Path{/mnt/plfs/vol5/ben/.plfs\_store}}
\item{FUSE strips the \Path{/mnt/plfs} and \plfs\ inserts the \store\ path and resolves to physical path \Path{/panfs/scratch/vol5/ben/.plfs\_store}}
\end{enumerate}

This works great for \fuse.  The \store\ is hidden from the user.  However, for
\adio, there is no current support for mapping so when a user wants to use
\adio, they need to know and use the path to their \store.  We have observed
that users are confused by the notion of the \store\ and therefore we believe
that the \adio\ in the current version is not ready for a production release.

We can have a production release with just \fuse, but Meghan has observed much
higher bandwidth with the \adio\ layer and believes that the bandwidth with 
just the \fuse\ layer might not be enough to convince users to switch.  
Recently, the developer of Bulk-IO has made some optimization by tuning the
IO to match the underlying stripe configuration of PanFS and although the
performance of \fuse\ is better than Bulk-IO it is not several times better
whereas the performance of \adio\ is several times better.

\section{Version 1}

Version 1 will move the mapping logic from \fuse\ into the library so it will
be shared with \adio.  However, we can no longer rely on the mount argument to
declare the \store\ as that is not seen by \adio.  We could use the mount
argument for \fuse\ and use hints for \adio\ but this exposes the \store\ to
users and we have already decided we don't want to do that.  Therefore, we have
decided that we will use an \Path{/etc/plfsrc} file on every node to declare
mappings between logical paths and {\store}s.  This file will be parsed by both
\fuse\ and \adio.

However, one problem is that we can no longer depend on symbolic links to
spread users across volumes.  We can either put all {\store}s on a single
volume or we can put a mapping entry into \plfsrc\ for every user
(\eg\ \Path{/plfs/ben} maps to \Path{/panfs/scratch/vol5/ben/.plfs\_store}).
This might be a maintenance problem however as every time a new user is
added, every \plfsrc\ file will need to be updated.  I don't know
the administrative cost of this.

\subsection{Compatibility}

Version 1 will be backwards compatible with Version 0.

\section{Version 2}

A better way to distribute users across volumes is for the \plfsrc\ to list
multiple {\store}s for each logical path.  Then \plfs\ can hash every file
and place it on a different \store.  This then becomes the \mds\ version.
We didn't intend possibly for the \mds\ version to be necessary so early
but the necessary mapping for the \adio\ layer has forced this issue earlier
than we expected.

In order to do the full \mds\ implementation, we need to decide on mechanisms
for file creation and for directory creation.  File creation is mostly
straight-forward: choose a \store\ and create the file there.  One open
questions is how to choose the \store\ for a particular file.  One approach
is to hash the filename, another is to hash the name of the node.

\section{Open issues and miscellaneous plans}

\subsection{Container identification}

Container identification is a pain in the butt.  SUID doesn't really work
and now \Path{container/accessfile} also seems not to be working.

\subsection{\adio\ optimizations}

\end{document}

### 2010-04-29 devel_path/random.tex
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@154 979fb63a-54b4-4514-8d67-9ab8ea1b59df

\begin{figure*}[tb]
  \centering
  \subfloat[\label{nonrandom}Ordered Dispatch]{
    \includegraphics[width=0.3\textwidth,height=\figheight]{dbvizGraphsAndConfigs/pointsLTE4.jpg}
  }
  \subfloat[\label{random}Random Dispatch]{
    \includegraphics[width=0.3\textwidth,height=\figheight]{dbvizGraphsAndConfigs/random4Pts.jpg}
  }
  \subfloat[\label{random-full}Final Results]{
    \includegraphics[width=0.3\textwidth,height=\figheight]{dbvizGraphsAndConfigs/allPoints.jpg}
  }
  \mycaption{fig-random}{Harnessing Randomness.}{
These three graphs illustrate the value of randomizing the generated commands
before dispatching them.  The graphs on the left and the middle were created
after only sixteen \subs\ were completed; the left-most is the graph produced when
the commands are dispatched in the order of their generation whereas the middle
graph was produced when the commands were randomly dispatched.  Finally, the
graph on the right shows the final results after all 492 \subs\ were completed.
By providing a much better approximation of the final graph, randomized 
commands allow computational steering much earlier in the experiment lifecycle.
}
\end{figure*}

### 2010-04-29 devel_path/related.tex
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@154 979fb63a-54b4-4514-8d67-9ab8ea1b59df

\section{Related Work}
\label{related}

\name\ introduces a new way to manage experiments in a HPC environment, greatly
simplifying the experimental process.  \name\ submits and executes jobs, parses
the results, inserts them into a database, and allows the user to graphically
and interactively view the results. Similar efforts can be found, but none
offer the holistic approach of \name\ nor its HPC focus.

GridBot~\cite{gridbot} is a system for executing experiments across multiple
grids and clusters. Its high-level goal is very similar to \name; both systems
are designed to ease the burden of computational scientists to manage very
large experiments with many \subs\ (\subs\ in GridBot are called \Term{Bags of
Tasks}). GridBot's focus however is on dispatch; its dispatch engine is
designed for the Grid however where the dispatcher in \name\ is designed for an
HPC environment.  Focused on dispatch, GridBot lacks any analogues for
transducers, \dbviz, or the ability to monitor and replay experiments.
Ultimately, the two systems are highly complementary: \name\ provides
end-to-end experiment management for HPC; combining it with the grid-aware
dispatching in GridBot would enable end-to-end experiment management for the
Grid.

VGrADS~\cite{vgrads} is a similar approach that provides a virtualized Grid
execution system; like GridBot, it is geared for the Grid whereas \name\ is
geared for HPC. VGrADS also attempts to provide reliable quality of service and
fault tolerance using automated task replication.  These features complement
\name's more holistic approach.

Zoo~\cite{livny-zoo} is another experiment management framework for desktop
computing.  Similar in concept to \name\, it acknowledges the experiment
feedback loop and provides a framework for making the experimental
process more manageable.  Zoo has modules, such as Opossum and Frog, for each step of the process, from designing and setting up experiments through data
collection and analysis.  Although novel for its time, Zoo, and its focus on
the desktop, provides insufficient management for the much more complex HPC
environment.  In addition, Zoo lacks the advanced features of \dbviz, such
as outlier analysis and hidden difference search.

GridDB~\cite{griddb} provides a data-centric overlay to process-centric grid
middleware.  The key differences between \name\ and GridDB exist in how they
handle data processing and analysis.  GridDB uses a declarative interface and
type checking to deal with data processing and interactive query processing to
support analysis.  These techniques offer flexibility, but they also make the
data more abstract.  \name\ uses transducers to convert experimental results
into usable data.  Transducers are slightly less flexible, but they help to
maintain the raw form of the data.  \dbviz\ offers easier data analysis and
\name\ seeks to keep data processing and analysis more decoupled by using
transducers and \dbviz.  \name\ provides a larger end-to-end focus whereas
GridDB offers a greater emphasis on automated computational steering.

### 2010-04-29 devel_path/table.tex
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@154 979fb63a-54b4-4514-8d67-9ab8ea1b59df

\begin{table*}[tb]
\centering
\begin{tabular}{|lrrr|}
\hline
TEST & Description Lines & Keys Collected & Transducer Lines \\
\hline
mdtest & 48 & 88 & 200 \\
ior & 45 & 101 & 143 \\
LANL fs\_test & 36 & 149 & 2 \\
\hline
\end{tabular}
\mycaption{eval-table}{Defining Experiments.}{
This table shows how many lines are necessary to describe typical parameter
sweeps for three different applications which are currently using \name\ at
LANL.  It also shows the number of \kv\ pairs collected from each \sub, and the
lines needed for the application-specific transducers which parse the output of
each \sub\ and produce these \kv\ pairs.  The LANL fs\_test transducer is so
small because LANL fs\_test already produces output in \kv\ format so that
transducer simply echoes its input to its output.  The other transducers are
larger because those applications do not have consistent output formatting so
every \kv\ must be parsed differently. 
}
\end{table*}

### 2010-04-29 devel_path/examples/mdtest.py

@@ -1,0 +1,24 @@
+import sys
+sys.path += '../lib'
+import expr_mgmt
+
+mpi_options = { "np"   : [ 4, 8 ], }
+mpi_program = ( "/users/user1/mdtest/mdtest" )
+program_options = {
+  "I" : [ 1, 10 ],
+  "d" : [ "/pfs/user1/mdtest.out" ], 
+  "z" : [ 1, 2 ],
+  "b" : [ 1 ],
+}
+transducer = "/users/user1/mdtest/trans.py"
+transducer_arguments = [
+  [ "--desc expr1" ]
+]
+
+def get_commands():
+  return expr_mgmt.get_commands(
+    mpi_options=mpi_options,
+    mpi_program=mpi_program,
+    transducer=transducer,
+    transducer_arguments=transducer_arguments,
+    program_options=program_options )

### 2010-04-29 devel_path/plfs_roadmap.tex
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@156 979fb63a-54b4-4514-8d67-9ab8ea1b59df

\documentclass[10pt]{article}
\usepackage{fullpage}

\usepackage{enumitem}
\setlist{noitemsep,topsep=0pt}
\setitemize{noitemsep,topsep=0pt}
\setenumerate{noitemsep,topsep=0pt}

\input{macros}
\input{watermark}

\begin{document}

\title{\TITLE}
\if 0
\author{
    John Bent,
    Ryan Kroiss,
    Alfred Torrez,
    Meghan Wingate
    \\
    Los Alamos National Lab
}
\fi
\date{}
\maketitle
\thispagestyle{empty}
\pagestyle{empty}
\section{Preface}

This document attempts to lay out a design path for the planned development
versions of \plfs.  It also attempts to describe how development versions can
move into production and the various issues for transitioning between versions.
Some of the issues are technical and some are policy.  Some of the technical
issues are things like how do we implement directory management in the future
\mds.  Policy issues are questions like whether incompatible layouts are 
handled transparently by \plfs\ or whether they are handled by explicitly
copying data from an old \plfs\ into the new one.

Our goal is to get \plfs\ into production at LANL.  However, we want to make 
sure that the production releases are bug-free, give performance attractive
to users, and are as easy to use as possible.  Hopefully by spelling out the
issues involved the current version of \plfs\ and the planned versions, we
can identify the target version for our first production release.  There is
of course a tension between releasing something soon and releasing something
good.  We could release the current version today but it has several 
limitations that will be discussed below and we need to determine whether
we are comfortable with those limitations.

At the end of this document, we will also describe current open issues
and other miscellaneous development plans.

\section{Current Version - Version 0}

The current version has support for both \fuse\ and for \adio.  However, the
\fuse\ implementation transparently maps logical paths into physical ones.  It
knows this mapping via a single mount argument naming the physical \store.  To
balance metadata workloads across PanFS director blades, each user's \plfs\
\store\ is located on a different PanFS volume.  This is accomplished via
symbolic links.  For example, \Path{/plfs/ben} is a symbolic link to
\Path{/mnt/plfs/vol5/ben/.plfs\_store}.  When \plfs\ is mounted, it is mounted
on \Path{/mnt/plfs} and directed to use \Path{/panfs/scratch} as its \store.
So a typical name resolution would work as follows:

\begin{enumerate}
\item{user ben opens logical file \Path{/plfs/ben/foo}}
\item{the sym link resolves this to logical path 
        \Path{/mnt/plfs/vol5/ben/.plfs\_store}}
\item{FUSE strips the \Path{/mnt/plfs} and \plfs\ inserts the \store\ path 
        and resolves to physical path 
        \Path{/panfs/scratch/vol5/ben/.plfs\_store}}
\end{enumerate}

This works great for \fuse: the \store\ is hidden from the user.  However, for
\adio, there is no current support for mapping so when a user wants to use
\adio, they need to know and use the path to their \store.  We have observed
that users are confused by the notion of the \store\ and therefore we believe
that the \adio\ in the current version is not ready for a production release.

\policy{
We need to hide the \store\ from users.  Therefore an \adio\ layer
without a transparent mapping cannot be included in a production release.
}

We can have a production release with just \fuse, but Meghan has observed much
higher bandwidth with the \adio\ layer and believes that the bandwidth with 
just the \fuse\ layer might not be enough to convince users to switch.

\policyq{Can we release a production version with just \fuse\ and not \adio?}

Recently, the developer of Bulk-IO has improved its bandwidth by tuning its 
IO to match the underlying stripe configuration of PanFS.  Although the
performance of \fuse\ is better than Bulk-IO it is not several times better
whereas the performance of \adio\ is.

\version{0}{Version 0 has only a \fuse\ layer (\ie\ no \adio\ layer) with a
simple mapping to the \store\ and balances metadata load via symbolic links
external to \plfs\ placing each user on a different volume.}

\section{Shared Mappings - Version 1}

Version 1 will move the mapping logic from \fuse\ into the library so it will
be shared with \adio.  However, we can no longer rely on the mount argument to
declare the \store\ as that is not seen by \adio.  We could use the mount
argument for \fuse\ and use hints for \adio\ but this exposes the \store\ to
users and we have already decided we don't want to do that.  Therefore, we have
decided that we will use an \Path{/etc/plfsrc} file on every node to declare
mappings between logical paths and {\store}s.  This file will be parsed by both
\fuse\ and \adio.

However, one problem is that we can no longer depend on symbolic links to
spread users across volumes.  We can either put all {\store}s on a single
volume or we can put a mapping entry into \plfsrc\ for every user
(\eg\ \Path{/plfs/ben} maps to \Path{/panfs/scratch/vol5/ben/.plfs\_store}).
This might be a maintenance problem however as every time a new user is
added, every \plfsrc\ file will need to be updated.  I don't know
the administrative cost of this.

Once we have both an \adio\ and a \fuse\ layer, we quickly can identify 
many optimizations possible in the \adio\ layer that aren't possible in the
current \fuse\ layer.  [Of course, we could consider extending the \fuse\
layer so that the \plfs\ daemons could communicate with each other but we're
not going to do that.]  These optimizations will be discussed in more detail
in Section~\ref{sec-adio-opt}.  This difference between the \adio\ and \fuse\
layers raises an important policy consideration:

\policyq{Many complicated issues can be easily resolved in the \adio\ layer.
Should we concentrate our efforts there and only support \fuse\ for
functionality and access from the front-ends?  Should we even bother putting
\fuse\ on compute nodes?  Can \fuse\ even work on Cielo?  If it can't work
on Cielo, then we just make inconsistent user environments by putting it on
Roadrunner.}

PVFS has decided to concentrate on the ADIO layer and not attempt to provide a
high performance POSIX file system interface.  Perhaps we should follow their
lead.

\version{1}{Version 1 will have both the \fuse\ and \adio\ layers and a mapping
within the library layer will hide {\store}s from the users.  With a simple
map file, all \plfs\ metadata for all users will go to a single PanFS director.
With a more complicated map, each user's metadata workload will go to different
directors.}

\policyq{Use a simple map and no metadata distribution or a complicated map
which spreads metadata workloads across PanFS directors?  Or skip Version 1
in favor of Version 2?}

\compatibility{Version 1 will be backwards compatible with Version 0.}

\section{Distributed Metadata - Version 2}

A better way to distribute users across volumes is for the \plfsrc\ to list
multiple {\store}s (where each is on a different volume) for each logical path.
Then, instead of relying on symbolic links as in Version~0, or complicated
mapping files as in Version~1, \plfs\ itself will load balance metadata
workloads by hashing every file and placing it on a different \store.  In fact,
this is even better than the crude balancing in Versions~0 and 1 because {\em
every file} is placed on a different volume as opposed to every user.

In order to do the full \mds\ implementation, we need to decide on mechanisms
for file creation and for directory creation.

\subsection{File Creation}

File creation seems straight-forward: choose a \store\ and create the file
there.  One open question is how to choose the \store\ on which to create a
particular file.  One approach is to hash the filename, another is to hash the
name of the node.  Although we discuss directory creation in detail below in
Section~\ref{sec-dir}, I want to introduce two basic approaches since the
selection of the directory creation approach influences some of the trade-offs
in the file creation.  One directory creation approach is to eagerly create
each directory on every \store.  Another is to just create each directory on a
single \store\ and then replicate lazily whenever needed.

\implementationq{Place files by hashing file name or node name?}

\begin{itemize}
\item{Place files by hashing on the file name}
\begin{itemize}
\item{Will distribute metadata for
a single client workload.  But this isn't really our target and I doubt the
metadata load from a single client benefits much from distribution anyway.
}
\item{File lookups are easy because we know where to look for the file.}
\item{Intuitively this seems bad for N-1 workloads because all access to the
file will be directed to a single \store.  However, remember that in \plfs\
there are no N-1 workloads (\ie\ each container dropping will be hashed to
a different \store).}
\item{Although every container dropping is hashed to a different \store, the
container itself is hashed to just a single \store\
(if we use lazy directory creation).  Therefore in a large N-1 workload,
every node will try to create the same container in the same location.}
\item{There might be more network contention since every client potentially
interacts with every director}
\end{itemize}
\item{Place files by hashing on the node name}
\begin{itemize}
\item{File lookups must search every \store.} 
\item{Buffering might work better and there might be less network contention
since every client interacts with just a single director} 
\end{itemize}
\end{itemize}

\subsection{Directory Creation}
\label{sec-dir}

First remember that there are two types of directories within \plfs.  There are
the logical directories created by users and then there are the directories
created within \plfs\ containers.

\implementationq{Should we treat logical directories and directories within
containers in the same way?}

It's tempting to place the directory logic and the file logic and all the
mapping logic in a layer below \plfs.  This means that logical directories
and container directories are handled in the same way.  However, handling
these different types of directories in the same way will probably cause
performance degradations.  If we violate the layering and handle these
differently then the below question of whether to do eager or lazy creation
is easy.  We do eager creation for logical directories and lazy creation
for container directories.

If we treat logical directories and container directories in the same way, then
we have to decide whether to eagerly create an instance of each directory on
every \store\ or whether to lazily create each instance whenever it is needed
to store a child entry.

\implementationq{Eagerly replicate directories across every \store\ or
lazily whenever needed to store a child entry?}

\begin{itemize}
\item{lazy directory creation}
\begin{itemize}
\item{Creating a child entry might require that the parent directory be
created first.  In order to do so, the other {\store}s must be searched
in order to ensure that the parent directory is valid.  This can recurse.  
In fact when Ben and I did a whiteboard
session we saw that this would happen frequently in container creation 
when the files were placed by hashing on the file name.}
\end{itemize}
\item{eager directory creation}
\begin{itemize}
\item{This could be really problematic in an N-1 workload on \fuse\ because 
every client would be trying to create the container on every \store. Most
attempts would fail and I believe a failed \syscall{mkdir} should be faster
than a successful one.}
\item{This could be even more problematic for N-N workloads in both \fuse\
and \adio\ because every client would successfully create a container on
every \store.}
\item{Failures during a \syscall{mkdir} operation could leave the file system
in an inconsistent state.  We probably still require the mechanism that
we create missing parent directories as needed.}
\end{itemize}
\end{itemize}

The more I think about this, the more I believe that:

\implementation{Logical directories should be replicated eagerly and 
container directories should be replicated lazily.}

\version{2}

Version 2 has built-in metadata distribution which works by spreading
files and directories across multiple {\store}s.  Since each store can
be a separate PanFS volume and since each PanFS volume is serviced by 
a different director, this approach distributes the metadata workload 
across multiple metadata servers: an approach that has been heretofore
difficult to achieve in PanFS.

Version 2 will probably replicate logical directories eagerly and
container directories lazily.  Given this, we should re-evaluate the
decision about how to select the \store\ on which to place newly created
files and container directories.  Probably if it wasn't late and I wasn't
tired, I might consider re-ordering these subsections and discussing 
directory creation first.  I feel that treating the logical directories and
container directories differently is the right appraoch although I was
initially reluctant to do it due to increased implementation complexity.

Now that I feel comfortable with that decision, the file placement decision
should be easier.

\begin{itemize}
\item{Place by hashing on file or container directory path}
\begin{itemize}
\item{Parent dirs might need to be created}
\item{Lookup can go directly to the correct \store}
\item{The set of {\store}s in a map cannot ever change}
\end{itemize}
\item{Place by hashing on nodename}
\begin{itemize}
\item{Parent dirs should always exist}
\item{Lookup must search every \store}
\item{The set of {\store}s in a map can change}
\end{itemize}
\end{itemize}

\compatibility{Version~2 may not be backward compatible.  Older versions
will definitely not be able to read Version~2 containers.  Version~2 may be
able to read containers created by older versions.  If we choose placement
by hashing on nodename, then Version~2 will look for files in all {\store}s.
This will be somewhat unnecessary since all the files will be located in
just a single \store\ but it will work.}

\section{Open issues and miscellaneous plans}

\subsection{Container identification}

Container identification is a pain in the butt but it's very important because 
without it \plfs\ has no way to distinguish between a container and a logical
user-created directory.  We have currently 
identified three possible container identification mechanisms:

{\bf 1) SUID.}  We can set the SUID bit on the container.  However, PVFS, HDFS,
and possibly other global file systems do not support the SUID bit on 
directories.  We have also seen that some users like to set the SUID bit on
their directories.  If they did this on a directory stored in \plfs, then the
next time they tried to access that directory, \plfs\ would mistakenly export
it as a container.  Also, there is no way to atomically create a directory 
with the SUID bit set.  Therefore in order to use this mechanism, we must
create a temporary directory, set the SUID bit, and then rename it.

{\bf 2) Use a specially named file within the container.}  We can create and
check for a specially named file within the container.  This also runs the risk
that a user creates a directory with an entry that happens to match our
specially named file which would cause \plfs\ to mistakenly export that
directory as a container.  Using a specially named file is the current
mechanism in \plfs~0.1.6 and it is set to be \Path{.plfsaccess113918400} so the
possibility of this accident happening seems unlikely.  This mechanism is also
not atomic and therefore requires a temporary directory which must be renamed.

However, we have recently seen a bug where \plfs\ is failing to create a file;
when it tries to create a dropping in the container, it incorrectly identifies
the container as a logical directory (\ie\ the container identification is
failing).  This might be a \plfs\ bug but right now it appears more likely that
it is a cache coherency problem in PanFS.  One node has created a temporary
directory, created the specially named file within, and has renamed the
temporary directory to the correct container path.  Another node checks that
correct container path and finds a directory {\em without} the specially named
file.

{\bf 3) Container name mangling.}  A third mechanism would be to mangle the
names of containers (\eg\ when a user creates logical file \Path{foo}, \plfs\
creates a directory named \Path{foo.plfscontainer113918400}).  This has the
advantage that it is atomic and therefore doesn't require temporary
directories.  But the disadvantage is that we need to unmangle these names on
\syscall{readdir}, we need to check for the mangled name on \syscall{mkdir},
and we need to check for both the mangled and unmangled names on
\syscall{stat}.  This can also add weird consistency issues if a logical
directory and a container of the same name are simultaneously created on
different nodes.  Hmmm.  Actually that's a problem with the other mechanisms as
well.  Stupid users.

\subsection{\adio\ optimizations}
\label{sec-adio-opt}

All sorts of stuff about when and whether to flatten and replicate indices.
And questions about how to read indices.  Indices could be globablly
constructed by subsets of the rank reading subsets of the index files and then
broadcasting to create the global index.

\subsubsection{Gary's ADIO ideas}
Gary's written a bunch about this.
Here is his email (sorry I haven't formatted it nicely yet):

\input{gary-squash}

\section{Final thoughts}

I feel like I can identify a path forward.  Deciding to eagerly replicate
logical directories and lazily replicate container directories hugely
simplifies things.  However, we still need to decide

\begin{itemize}
\item{whether to place files and 
container directories by hashing their paths or by hashing the node names.  
}
\item{whether to attempt to optimize both the \adio\ and \fuse\ layers or
just \adio}
\item{whether to run \fuse\ on compute nodes or just front-ends}
\item{which version to push for production}
\end{itemize}

\end{document}

### 2010-04-29 devel_path/examples/mdtest.py

@@ -1,24 +1,0 @@
-import sys
-sys.path += '../lib'
-import expr_mgmt
-
-mpi_options = { "np"   : [ 4, 8 ], }
-mpi_program = ( "/users/user1/mdtest/mdtest" )
-program_options = {
-  "I" : [ 1, 10 ],
-  "d" : [ "/pfs/user1/mdtest.out" ], 
-  "z" : [ 1, 2 ],
-  "b" : [ 1 ],
-}
-transducer = "/users/user1/mdtest/trans.py"
-transducer_arguments = [
-  [ "--desc expr1" ]
-]
-
-def get_commands():
-  return expr_mgmt.get_commands(
-    mpi_options=mpi_options,
-    mpi_program=mpi_program,
-    transducer=transducer,
-    transducer_arguments=transducer_arguments,
-    program_options=program_options )

### 2010-06-08 devel_path/plfs_roadmap.tex
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@164 979fb63a-54b4-4514-8d67-9ab8ea1b59df

@@
 
+OK.  Here the current plan:
+
+\begin{itemize}
+\item{global mkdir for logical directories}
+\item{local mkdir for container directories hashed by node name}
+\item{local create for files hashed by node name}
+\item{global parallel metadata ops (including things like chmod)}
+\end{itemize}
+
+The global parallel metadata ops do leave us vulnerable to crashes.  We
+need a new term to explain this.  With striping across volumes, on an
+N-1 write workload, we'll have multiple nodes each create
+"subcontainers" on different volumes.  Each subcontainer will be created
+in exactly the same way that containers are created now.  On a read,
+PLFS will aggregate the indices from all the subcontainers and construct
+a global view of the container.  So the consistency problem is that we
+crash halfway through a chmod.  Each subcontainer will have an
+accessfile that holds its permissions.  When we go to aggregate the
+subcontainers' permissions into the permissions for the container, we'll
+discover that they are not consistent.  I guess in this specific case,
+we'll just use the strictest.  If the user is surprised, they'll just do
+a chmod again to fix it.  I'll have to think whether worse problems
+occur for other possible inconsistencies.
+
+\definition{Subcontainer}{If multiple nodes each create containers for
+the same logical file on different volumes, each of these is actually
+a \Term{subcontainer} and the logical aggregated view is a \Term{container}.
+To aggregate a view of a container, \plfs\ will look across all {\store}s
+for subcontainers for that container.}
+
+By the way, we used to use the link idea back before I stripped the
+volume striping out.  This work by hashing the container itself by name
+and the hostdirs within were hashed by node and then lazily linked in to
+the container.  You could do the link at create time but then there is a
+parallel open storm.  This avoided the consistency problem and made
+lookups easier.  It also meant that the read worked for free.  In the
+above approach, we might have to modify the global index build since the
+data droppings are striped across multiple directories.  But I didn't
+like that approach since lazily linking in the hostdirs was complicated.
+
 \end{document}

### 2010-07-20 devel_path/plfs_roadmap.tex
trying to speed up the svn co
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@204 979fb63a-54b4-4514-8d67-9ab8ea1b59df

@@
 \usepackage{fullpage}
-
+\usepackage[pdftex]{graphics}
 \usepackage{enumitem}
+\usepackage{subfig}
+\usepackage{amsmath}
+\usepackage{alltt}
 \setlist{noitemsep,topsep=0pt}
@@
 
-However, one problem is that we can no longer depend on symbolic links to
-spread users across volumes.  We can either put all {\store}s on a single
-volume or we can put a mapping entry into \plfsrc\ for every user
-(\eg\ \Path{/plfs/ben} maps to \Path{/panfs/scratch/vol5/ben/.plfs\_store}).
-This might be a maintenance problem however as every time a new user is
-added, every \plfsrc\ file will need to be updated.  I don't know
-the administrative cost of this.
+Figures~\ref{fig-conf}~and~\ref{fig-mapping} show how the mapping function
+can be defined such that the same logical path will be resolved to the same
+physical path on the \upfs.
 
@@
 layers raises an important policy consideration:
+
+\begin{figure}[t]
+\centering
+\begin{tabular}{ p{3in} p{3in}}
+\input{config.tex}
+&
+\input{admin.tex}
+\\
+\centering a) Configuration File & b) Administrator Script 
+\end{tabular}
+\mycaption{fig-conf}{Shared Mappings.}{On the left is an example \plfsrc\
+file listing various parameters used to control the behavior within
+the PLFS library.  Of interest here is how the mapping works to find a physical
+location from a logical path.  The hashing function itself works by turning
+the username into the uid and then taking the modulus of the uid and the
+number of backends and then using that result as an index into the backend
+array.  Note that the complexity of the hashing algorithm will really just
+exist in the PLFS library but is shown here for the sake of expediency.  
+Notice also that this method distributes each user's metadata workload 
+to a different metadata server.  Of course, the entire metadata workload
+for each individual user is isolated on a single metadata server.  We will
+fix this in Version~2.
+The right shows the script that an administrator
+can use to setup the necessary directories.  The \$user\_list is defined to
+be all users during initial creation and then set to be just a single user
+when a new user is added.}
+\end{figure}
+
+\begin{figure}[t]
+\centering
+\includegraphics{mapping.png}
+\mycaption{fig-mapping}{Path Resolution in Shared Mappings.}{This figure
+shows how the same logical path resolves in both \fuse\ and \adio\ to the
+same path on the \upfs.  We assume that the \plfsrc\ used is that shown
+in Figure~\ref{fig-conf}.}
+\end{figure}
 
@@
 \version{1}{Version 1 will have both the \fuse\ and \adio\ layers and a mapping
-within the library layer will hide {\store}s from the users.  With a simple
-map file, all \plfs\ metadata for all users will go to a single PanFS director.
-With a more complicated map, each user's metadata workload will go to different
-directors.}
-
-\policyq{Use a simple map and no metadata distribution or a complicated map
-which spreads metadata workloads across PanFS directors?  Or skip Version 1
-in favor of Version 2?}
+within the library layer will hide {\store}s from the users.  
+Path resolution in both \fuse\ and \adio\ will share the same mapping such
+that the same logical path presented to both will resolve to the same
+physical path on the \upfs.
+Each user's metadata workload will be isolated on one of a set of metadata
+servers such that simultaneous workloads from multiple users will be spread
+across multiple metadata servers.
+}
 

### 2010-10-05 design_choices/other_writes/plfs_other_write.tex
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@346 979fb63a-54b4-4514-8d67-9ab8ea1b59df

\documentclass[10pt]{article}
\usepackage{fullpage}
\usepackage[pdftex]{graphics}
\usepackage{enumitem}
\usepackage{subfig}
\usepackage{amsmath}
\usepackage{alltt}
\setlist{noitemsep,topsep=0pt}
\setitemize{noitemsep,topsep=0pt}
\setenumerate{noitemsep,topsep=0pt}

\addtolength{\textwidth}{.5in}
\addtolength{\hoffset}{-.25in}
\addtolength{\textheight}{.5in}
\addtolength{\voffset}{-.5in}

\input{macros}
%\input{watermark}

\begin{document}

\vspace{-1in}
\title{PLFS Policy Regarding Writes by Others}
\if 0
\author{
    John Bent,
    Ryan Kroiss,
    Alfred Torrez,
    Meghan Wingate
    \\
    Los Alamos National Lab
}
\fi
\date{}
\maketitle
\thispagestyle{empty}
\pagestyle{empty}
\vspace{-1.5cm}

PLFS files are physically stored in a directory structure.  This creates some
difficulty since the UNIX permissions work differently on directories than they
do on files.  We would like to access the security protections of the 
underlying file system (\store) as much as possible but we might occasionally
have to attempt some enforcement ourselves.  Of course, this opens a security 
hole since the physical files on the \store\ cannot be hidden.

One particular thorny problem is allowing users to grant write permissions to
other users on their files (writes by others).  We can disallow this but if
we'd like to export a file system with capabilities that users are accustomed
to, than we might try to support this.  A state diagram depicting possible
policy decisions is shown in Figure~\ref{fig-state}.

Our first option is to just disallow writes by others.  This is easy
to implement and will not allow any unauthorized access on the \store.  To
implement this, we merely remove write permissions for the group and
for others ({\tt \ie\ mode \& (!(S\_IWGRP|S\_IWOTH))})

\begin{figure}[h]
\centering
\includegraphics{state.png}
\mycaption{fig-state}{Possible Policies in Regards to Writes by Others.}{This 
state diagram depicts the different policies and their implications in regards
to whether and how PLFS can support owners of files granting write permissions
on those files to other users or groups.}
\end{figure}

If we want to allow writes by others, there are two ways in which we can 
implement this.  The first way is that when \droppings\ are created inside
\plfs\ {\container}s, they are created with permissions as defined by the 
{\tt mode\_t} passed to the {\tt open()} call.  The problem with this 
approach is that when user B writes to A's file, \droppings\ will be created
in A's container with permissions that may not allow A access to them.
This could prevent A from unlinking the file or potentially even reading the
file if B's {\tt mode\_t} is very restrictive.  To somewhat ameliorate this
problem, when A later does a {\tt chmod()} or a {\tt chown()} on the file 
through \fuse\ when \fuse\ is run as root, \fuse\ will recurse through the
file and restore ownership of all \droppings\ to A.  So A can eventually
have full access to the file after doing a {\tt chmod()} or {\tt chown()}.

The third approach, and the approach taken by Version 1.0.0, is for all
\droppings\ to always be created with full permissions (\ie\ ({\\tt
 (S\_IRWXU|S\_IRWXG|S\_IRWXO)}) which allows A to easily remove any 
\droppings\ created by B.  We then attempt to limit permissions by
setting the permissions of all directories within the \container\ 
appropriately.  The drawback of this approach is that if A gives B read
permissions on the file, then the \droppings\ will be globally writable
and the directories will allow B read access.  Therefore B cannot 
create new droppings but can overwrite and truncate existing droppings.
We explicitly check for this and prevent this for access through \fuse,
but this does allow B unauthorized access through the \store.

Given these constraints and our desire to make a filesystem conforming
the most closely to user expectations and given the lack of obvious
visibility to the \store, we have currently chosen to use the third
approach in Version 1.0.0.  This allows write by others and allows
the original owner full access to their files after write by others
without any intermediate steps such as {\tt chmod()} or {\tt chown()}.
This does allow unauthorized overwriting on the \store\ when an other
has only read access.

\end{document}

### 2010-10-05 roadmap/eval.tex
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@346 979fb63a-54b4-4514-8d67-9ab8ea1b59df

\section{Usage Model}
\label{eval}

\input{table}

To use \name, a user first creates an experiment description file.  Because
\name\ handles both command generation and command dispatch automatically,
these description files need merely identify parameters and values of interest.
An example of such a file is shown in Figure~\ref{mdtest-desc}.  If the user
desires more than the default set of \kv\ pairs captured by \name, they can
additionally create an application-specific transducer.  Table~\ref{eval-table} 
lists the lines needed to create description files and transducers for
several applications currently using our framework.

After defining descriptions and transducers, users can then begin to run their
experiments with the \dispatcher.  The user can optionally ask the \dispatcher\
to shuffle the \subs\ randomly before dispatching.  The value of this
randomness is shown in Figure~\ref{fig-random}.  By utilizing randomness, users
can more quickly see trends within their experiments than if the \subs\ are run
ordered by the parameter sweeps.  Note that utilizing randomness in this way is
not a new result nor is it our contribution.  Rather our contribution here is
that the \sub\ tracking and monitoring provided by \name\ allow randomness to
be exploited in this manner.  Without such simple tracking and monitoring, it
is much more difficult for the user to run \subs\ randomly; in the case of
interrupted experiments, naive resubmission will rerun previously completed
\subs\ and slow total time to completion for the experiment.

As the \subs\ begin to complete, the user can track their progress with the
monitoring within \name, and can visualize the results with \dbviz.  The
following describes an example interactive session with \dbviz.

\begin{enumerate}
\item{The \dispatcher\ has already executed a set of \subs;
each \sub\ was executed ten times to provide
statistically meaningful results and variability measurements.  }
\item{The entire experiment was tagged with a \Term{description} of \Term{expr1}.}
\item{The transducer has collected general \kv\ pairs and has parsed specific
ones and has inserted them into a SQL server.}
\item{The user specifies "WHERE description like 'expr1'".}
\item{The user specifies the column \Term{numhosts} as the \xaxis\ 
and the column \Term{walltime} as the \yaxis.}
\item{The user asks to compare columns \Term{depth} and \Term{items} which
were the parameters swept and which the transducer parsed from the command
line.}
\item{\label{interactive-graph}A graph is created similar to that shown in Figure~\ref{dbviz-screen}.  The user 
notices unusually high variability and changes the display from lines to a 
scatter plot and discovers a few outliers. }
\item{Using outlier analysis,
the user pulls the complete row from the table for each outlier and
notices that some of the outliers were from \subs\ that did not complete
successfully.  To remove them from the analysis, the user adds a
successful \Term{retval} condition to the WHERE clause.}
\item{The user returns to linespoints mode but the variability still looks higher than expected.  The user queries the set of hidden differences and discovers
that data from multiple supercomputers with drastically different performance
characters has been combined.  The user then either adds \Term{system} to
the comparison fields or refines the filter to examine \subs\ from only a
single \Term{system}.}
\item{The user discerns that performance has degraded from previous 
experiments.  The user then combines multiple experiments and queries 
hidden differences again which reveals that \Term{osversion} has changed.
The user adds \Term{osversion} to the comparison fields and identifies that
as a reason for the performance degradation.
}
\end{enumerate}

A screenshot of \dbviz\ corresponding to step~\ref{interactive-graph} above is
shown in Figure~\ref{dbviz-screen}.  Notice that the key for the graph is 
constructed using each unique combination of the columns as specified in the
\Term{compare} field.  By default, these keys would read as \Term{k=v,k=v,...}
but using regular expressions in the Preferences panel (not shown) allows the
user to create semantically meaningful substitutions for \kv\ pairs.  For example, in this screenshot, the user has created a substitution to replace \Term{target=.*plfs.*} with \Term{PLFS-MNT}.

\input{random}

### 2010-10-05 roadmap/gary-squash.tex
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@346 979fb63a-54b4-4514-8d67-9ab8ea1b59df

The \plfs\ library needs to support something like this functionality:

\begin{enumerate}
\item{ I have this open fd, please give me the local map for
this fd (this is not for the entire map, it is for my part
only
parms: open fd, pointer that the lib would put an address of some
space the lib mallocs where it puts the map part, and a length of
the map part}
\item{ please store this squashed entire map in this
specified subcontainer
parms: fd, address of map, length of map, container to store}
\item{ mark the metadata for this fd as successfully squashed and distributed
this tells the lib to put a metadata dropping in
parms: fd}

\item{ using this consolidated set of map blobs , make me a squashed map
parms: addr of consolidated map blobs, length of consolidatd map blobs,
addr to put squashed map, length of resultant squashed map}

\item{ is there a metadata dropping that this has been squashed and
distributed for this file and where are the locations one might
look for a replica - this gives a valid list of places where
indexes live and whether or not there is a valid squashed index
(subcontainers),
may need an fd and file path version of this or a flag for lookup type
parms: file path, fd, addr for lib to put location list, length of list}

\item{ open for read with squashed index replica location or a squashed map
provided in memory
parms: open stuff plus replica location or map squashed provided}
\end{enumerate}

So an mpi-io adio app on would do "something like" this:
\\
open for write
\\
write stuff
\\
on close for write "this close for write" is a bit of a complication"
(I include frees required by the lib - other malloc/free would
be required by the adio layer itself)
- all would ask for their fd (local) index blob  (new func 1)
 (I suppose we didnt write anything is a possible outcome)
\\
- all reduce these local index blobs to proc 0
\\
- proc 0 calls using these parts make me a squashed map (new func 4)
\\
- proc 0 calls is there squashed metadata for this file
  which gives proc 0 a list of subcontainers (new func 5)
\\
- proc 0 assigns work (squashed map and subcontainer) to get
  all the squashed maps distributed
\\
- all workers call please store this squashed map in this subcontainer
(new func 2)
\\
- all free the blobs local blobs (allocated by the lib)
\\
- proc 0 frees squashed map (allocated by the lib)
\\
- all but proc 0 barrier
\\
- proc 0 mark metadata for this fd as successfully squashed and
replicated  (new func 3)
\\
- proc 0 barrier
\\
- all call close

So you can see that the lib doesnt do any mpi magic and the adio (or
app) doesnt know what is in the map or a subcontainer, they are just
blobs of data used to distribute, because it has the distribution
mechanism.

You notice that the distribution of the squashed indexes to
the brethren does not just blindly do a broadcast and write
the index, because you might want to write fewer indexes
than procs.

So this gets the index squashed on close for write by adio.
We might want a hint on whether you want squashed indexes
on this mpi run so you could avoid this if you dont want/need
it.  Probably should plan on that up front even if we dont
want to use it.

Now the adio on open for read would
proc 0 in the adio layer would check to see if there
is there a valid squashed index and get the location list  (new func 5)
\\
--if valid squashed index
\\
----  assign replicas to the brethren
\\
---- everyone opens using that squashed index  (few func 6 using
subcontainer)
\\
--if there is no valid squashed index
\\
---- proc 0 has a list of places where indexes live - from the query for
squash above - even though there is no squashed index,
\\
---- proc zero assigns work to the brethren (may not be equal number of
workers and number of indexes to read),
\\
---- each proc does open read with their index part assigned (new func 6
using subcontainer)
\\
---- get local map (new func 1)
\\
---- send it back to proc 0,
\\
---- when proc 0 gets them all it calls build me a squashed index from
these parts into the lib (new func 4)
\\
---- broadcast resultant squashed map to all,
\\
---- everyone opens with squashed map provided (new func 6 - using mem
map) (I suppose you could write the squashed maps to disk here but that
seems bad on a read op)

adio close on read has nothing much to do maybe frees or something,
again you could write the squashed indexes here but on a read
that is probably bad

A stand alone mpi utility to squash would then:
\\
- just be an mpi program that open rw (because it is open for read it
would efficiently get the squashed map with as many procs as you have
available - doesnt have to match the number of containers) - everyone
would end up with a copy of the fully squashed map
\\
- close for rw  (Because this is a close for write the squashed indexes
are written)

A stand alone non mpi utility to squash would be essentially the same
thing as what the mpi open for rw and close for write is, but
the farm out the work would be a loop to do the work and not to assign
the work, or maybe do it with threads where there was procs, which
would make the stand alone utility to be basically the same as the mpi
utility.

Again, this close for read/close for write seems to be a slight
complication, but maybe an important capability.

Of course this is all notional, but

\begin{enumerate}
\item{this seems doable}
\item{works where num subcontainers doesnt equal num procs}
\item{leverages mpi flie open and close}
\item{gets us easy stand alone mpi and non mpi squasher}
\item{nearly optimal for mpi and helps posix hugely}
\end{enumerate}

Now the posix open for read does something similar to
what proc zero and the other procs do, it would need
to call the same sorts of routines.  It would probably
just need to randomly pick a subcontainer to
do the open with container for squashed index.
The posix close probably wouldnt do a squash, as it
wont know when people are done and how to do it and
who else is doing it etc.

### 2010-10-05 roadmap/plfs_roadmap.tex
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@346 979fb63a-54b4-4514-8d67-9ab8ea1b59df

--- 2010-06-08 devel_path/plfs_roadmap.tex
@@
 \end{itemize}
+
+\subsection{Archive Workloads}
+
+Gary Grider did a bunch of measurements on an HB machine and sent email on
+August 3, 2010 about this.  The archive workload consists of a small number of
+processes running on a small number of nodes reading a file in 500 MB strides.
+When the file was created with a large number of writers, the time to open it
+for reads (\ie\ the time to aggregate the index chunks into a global offset
+lookup table) becomes large.  We added multi-threading to deal with this and
+it helps but we anticipate that for Cielo-scale machines, these thread pools
+will be insufficient and we'll need to solve this with index squashing.  Also,
+the read bandwidth itself is OK for files created with 500K writes but runs
+slowly for files created with 50K writes.  Specifically, it reads at about
+a third of the bandwidth that files can be read directly from Panasas.  We
+don't know what to do about this at the current time but this doesn't need
+to be solved for Roadrunner and may not prove to be a problem even for Cielo.
 

### 2010-10-05 design_choices/other_writes/plfs_other_write.tex
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@347 979fb63a-54b4-4514-8d67-9ab8ea1b59df

@@
 We explicitly check for this and prevent this for access through \fuse,
-but this does allow B unauthorized access through the \store.
+but this does allow B unauthorized access through the \store.  However,
+given how we implement {\tt chmod()} and {\tt chown()} as described
+above, user A can {\tt chmod()} and {\tt chown()} the file and this
+will close the hole on the \store.
 
-Given these constraints and our desire to make a filesystem conforming
-the most closely to user expectations and given the lack of obvious
-visibility to the \store, we have currently chosen to use the third
-approach in Version 1.0.0.  This allows write by others and allows
-the original owner full access to their files after write by others
-without any intermediate steps such as {\tt chmod()} or {\tt chown()}.
-This does allow unauthorized overwriting on the \store\ when an other
-has only read access.
+Given these constraints and our desire to make a filesystem conforming the most
+closely to user expectations and given the lack of obvious visibility to the
+\store, we have currently chosen to use the third approach in Version 1.0.0.
+This allows write by others and allows the original owner full access to their
+files after write by others without any intermediate steps such as {\tt
+chmod()} or {\tt chown()}.  This does allow unauthorized overwriting on the
+\store\ when an other has only read access but only until the owner does a {\tt
+chmod()} or {\tt chown()}.
 

### 2012-09-17 hdf5/discs_2012/abstract.tex
Adding existing files

HDF5 is a data model, library and file format for storing and managing data. It is designed for flexible and efficient I/O for high volume and complex data. Natively, it uses a single-file format where multiple HDF5 objects are stored in a single file. In a parallel HDF5 application, multiple processes access a single file, thereby resulting in a performance bottleneck in I/O. Additionally, a single-file format does not allow semantic post processing on individual objects outside the scope of the HDF5 application. The HDF group has recently introduced a new layer called the Virtual Object Layer (VOL) that allows writing plugins for HDF5 to store data in different ways. We have written a new plugin for HDF5 using PLFS that provides serves two purposes: 1) it uses PLFS to convert the single-file layout into a data layout that is optimized for the underlying file system, and 2) it stores data in a unique way that enables semantic post-processing on the data, independently of HDF5. In this paper, we measure the performance improvements available through the PLFS VOL and discuss future work leveraging the new semantic post-processing functionality enabled.  We further discuss the applicability of this approach for exascale burst buffer storage systems.

### 2012-09-17 hdf5/discs_2012/background.tex
Adding existing files

\subsection{HDF5}
HDF5 is a data model , library and file format for storing and managing data. It defines an information set, which is a container of array variables, groups, and types. They are known as datasets, groups and datatype objects in HDF5 respectively. The data model defines mechanisms for creating associations between various information items. The main components of HDF5 are briefly described below.

File:
In the HDF5 data model the container of an HDF5 infoset
is represented by an HDF5 file. It is a collection of HDF5 objects and represents the relationship between them.
Every file begins with a root group "/", which serves as the "starting-point" in the object hierarchy. An empty file has atleast the root group in it.

Datasets:
HDF5 datasets are objects that represent actual data or content. Datasets are arrays which can have multiple dimensions. A dataset is characterized by a dataspace and a datatype. An HDF5 dataspace captures the rank (number of dimensions) , and the current and maximum extent in the respective dimensions. An HDF5 datatype describes the type of its data elements. 
HDF5 supports ten classes of datatypes including scalar types like integer, floating-point, string, etc. and compound datatypes.

Groups:
HDF5 groups are akin to directories in a file system (reference here). A group is an explicit association between HDF5 objects. A group could contain multiple other groups, datasets or datatypes /textit{within} it. 
Traversing a root group generates the /information set graph/.

Attributes:
Aattributes are used for annotating datasets, groups, and datatype objects. They are datasets themselves, and are /textit{attached} to the existing objects they annotate.

The above definitions are explained in the example shown in figure ~\ref{hdf5_example}. 
The file "sample.h5" contains a root group which itself contains a group G1 and two datasets, D1 and D2. Group G1 contains a dataset D3. Attribute A1 is linked to dataset D1. The datatype in all datasets in an integer.

\begin{figure}[!t]
\centering
\includegraphics[width=2.5in]{hdf5_example}
\caption{A sample HDF5 file}
\label{hdf5_example}
\end{figure}

HDF5 file format:
An HDF5 file is a self-describing format which combines data and metadata. High level objects are stored alongwith metadata that is used to describe various characteristics about the object, alongwith the relationship between various objects.

Parallel HDF5 (PHDF5):
As described previously, HDF5 supports parallelism using MPI. Multiple processes can be used to create and store objects in a file. PHDF5 exports a parallel I/O interface, where internally makes use of MPI-IO to perform I/O in parallel. 
Applications typically store multiple objects in a single HDF5 file. However, many popularly used parallel file systems like Lustre, Panasas etc. are known to perform poorly for workloads where multiple processes access a shared file (commonly referred to as the N-->1 access pattern).

\subsection{PLFS}
PLFS is a middleware virtual file system that converts writes to a shared logical file into writes to multiple physical files. 
It is situated between the application and the parallel file system responsible for the actual data storage. 
It transforms an application's N-->1 access pattern into an N-->N access pattern, where every process participating in I/O writes data to its own , separate file. 
Thus, it interposes on application I/O and converts its I/O pattern into one that is more suitable for the underlying parallel file system.
The basic operation of PLFs is as follows. For every logical file, PLFS creates a /texit{container} structure on the underlying parallel file system. 
This container is a hierarchical directory tree consisting of multiple sub-directories. Multiple processes opening the same logical file for writing get a unique data file within the container. When the processes write to their file, a record identifying the write is appended to an index file shared between all processes. Thus, PLFS maintains sufficient metadata to recreate the shared logical file. A FUSE daemon exports this container structure as a single, logical file to the end-users, thereby keeping the PLFS operations transparent from them.

### 2012-09-17 hdf5/discs_2012/introduction.tex
Adding existing files

Hierarchical Data Format (HDF5) is a data model , library and file format for storing and managing data. It is designed for flexible and efficient I/O. HDF5 defines an information set, which is a container of array variables, groups, and types. The data model defines mechanisms for creating associations between various information items. HDF5 is widely used in the industry and scientic domain, in understanding global climate change, special effects in film production, DNA analysis, weather prediction, financial data management etc. (citation here)
Parallel HDF5 (PHDF5) enables developing high performance, parallel applications using standard technologies like MPI in conjunction with HDF5. An HDF5 file created using PHDF5 is compatible with serial HDF5 files and is shareable between different platforms. PHDF5 exports a standard parallel I/O interface which then uses MPI's parallel I/O functionality.

The central component of HDF5, the file, is a self-describing format which combines data and metadata. Users typically store multiple HDF5 objects in a single file, and the library stores metadata alongwith data that describes the relationships between various objects amongst other metadata. 
However, this native single-file format has its disadvantages. A PHDF5 application has multiple processes accessing a single HDF5 file. Many popular parallel file systems are known to behave poorly under these circumstances. Secondly, since many HDF5 objects are stored in a single file, this eliminates any possibility of performing useful semantic analysis on objects beyond the scope of an HDF5 application. In this paper, we address these shortcomings of HDF5, viz. performance issues due to multiple processes accessing a single shared file, and lack of a way to perform useful post-processing on HDF5 objects inherent due to the native file format.

PLFS is a middleware virtual file system developed at Los Alamos National Lab (LANL). It converts writes to a shared logical file into writes to multiple physical files.  
Thus, it interposes on application I/O and converts its I/O pattern into one that is more suitable for the underlying parallel file system.
PLFS is popularly used as a checkpoint file system where applications demonstrating heavy checkpoint workloads are known to benefit significantly from PLFS.

In this work, we have developed a new plugin for HDF5 using its recently introduced Virtual Object Layer (VOL). The VOL exports an interface that allows writing plugins for HDF5, where it primarily passes references to HDF5 objects, thereby enabling developers to store objects in a format different from the default HDF5 file format. 
We have written a plugin with two main objectives. The plugin stores data in a unique way that enables semantic post-processing on HDF5 objects outside the scope of the HDF5 application. Secondly, PLFS converts N-->1 accesses into N-->N accesses, thereby showing a significant improvement in performance as well. Preliminary results using HDF5's h5perf performance tool show a Xx performance improvement in I/O overall traditional PHDF5.

The rest of the paper is organized as follows. We represent details about HDF5 and PLFS in section 2, describe our plugin design and implementation in section 3, and our evaluation in section 4. We present related work in section 5, the current status of the plugin alongwith future work in section 6, and finally we conclude in section 7.

### 2012-09-17 hdf5/discs_2012/plugin.tex
Adding existing files

The plugin stores data in a format which enables performing semantic analysis on the data. 
It stores every HDF5 object in a separate file/directory, so that data from different objects is not contained in the same file. 
The central object store in HDF5, the HDF5 file, is represented by a directory. Groups are stored as directories as well, whereas datasets, which contain raw data, are stored in files. 
The relationship between objects is explained by the relative paths between the objects at the file system level. As an example, the HDF5 object D1 which in HDF5's B-tree format can be represented as /G1/G2/D1 is stored on the file system as /path-to-hdf5-file/G1/G2/D1. This way , the plugin eliminates the need to store the metadata describing the relationship between objects. Metadata about datasets, such as the datatype, extent, dimensions etc., are stored as PLFS Xattributes (xattrs) as explained in the previous section.

Consider the example shown figure 1. Using the plugin, file test.h5 is stored as a directory. Groups G1, G2, G3 are represented as directories at the paths /tmp/test.h5/G1, /tmp/test.h5/G2 , /tmp/test.h5/G1/G3/ respectively. Datasets D1 and D2 are files stored under their respective groups.

This approach gives us the ability to distinguish HDF5 objects at the file system level, without requiring an HDF5 application to provide information about objects and the relationship between them. 
We can now perform post-processing and semantic analysis on the data, outside the scope of the HDF5 application.

%The following table lists the VOL functions currently provided by the PLFS plugin:
%\begin{itemize}
%\item Attribute yes
%\item Dataset yes
%\item File yes
%\item Group yes
%\item Object No
%\item Link No
%\end{itemize}

### 2012-09-19 hdf5/discs_2012/background.tex
Started with evaluation, writing bibliography now, included bib files

@@
 
+Virtual Object Layer (VOL):
+(I have simply copied this part; need to reframe it) The Virtual Object Layer (VOL) is a new abstraction layer internal to the HDF5 library (reference here).
+It is implemented just below the public API.
+The VOL intercepts all HDF5 API calls that could potentially access objects in the file and forwards those calls to an “object driver” plugin.
+The plugins could actually store the objects in variety of ways.
+A plugin could, for example, have objects be distributed remotely over different platforms, provide a raw mapping of the model to the file system, or even store the data in other file formats (like native netCDF or HDF4 format).
+The user still gets the same data model where access is done to a single HDF5 “container”; however the plugin object driver translates from what the user sees to how the data is actually stored.
+
 \subsection{PLFS}
@@
 A FUSE daemon exports this container structure as a single, logical file to the end-users, thereby keeping the PLFS operations transparent from them.
+We added a new feature to PLFS called XAttributes (xattrs). Xattrs serve as short metadata stored as key-value pairs.
+All xattrs are stored in a new directory inside the PLFS container structure.
+They can be used to store user-defined information about data for easy and fast retrieval, since they are not stored alongwith raw data.
+We use xattrs to store metadata about HDF5 datasets. The datatype, number of dimensions and their extent etc. are stored as xattrs.

### 2012-09-19 hdf5/discs_2012/evaluation.tex
Started with evaluation, writing bibliography now, included bib files

For evaluation purposes, we have used HDF5's \textit{h5perf} performance tool (reference here). h5perf is a tool for testing the performance of PHDF5. It allows configuring various parameters, some of which are the number of processes to run the test with, number of datasets to be created in the file, amount of data contributed by each process per dataset, amount of data read/written by a process in a single I/O call (transfer size) etc. For our measurements, we create 10 datasets in the file, and the total file size was 64 GB or more.

All tests were performed on the Lustre parallel file system on the Atlas cluster at University of Dresden. The file system has 12 OSTs configured with a stripe size of 1 MB. The file system is connected to the compute nodes using an SDR Infiniband link. 
The cluster has 92 nodes with 64 cores each and 64 to 512 GB memory. All nodes are AMD Opteron 6274 running at 2.2 GHz.

Our tests were configured such that we spawn a maximum of 4 processes per node. Tests run with less than 4 processes utilized a single node.

%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{4pes_interleaved_1m10_coll_writes}
%\caption{A sample HDF5 file}
%\label{writes}
%\end{figure}
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{4pes_interleaved_1m10_coll_reads}
%\caption{A sample HDF5 file}
%\label{reads}
%\end{figure}

### 2012-09-19 hdf5/discs_2012/abstract.tex
Added figures, references. Some corrections.

@@
 Additionally, a single-file format does not allow semantic post processing on individual objects outside the scope of the HDF5 application.
-The HDF group has recently introduced a new layer called the Virtual Object Layer (VOL) that allows writing plugins for HDF5 to store data in different ways.
-We have written a new plugin for HDF5 using PLFS that provides serves two purposes: 1) it uses PLFS to convert the single-file layout into a data layout that is optimized for the underlying file system, and 2) it stores data in a unique way that enables semantic post-processing on the data, independently of HDF5.
-In this paper, we measure the performance improvements available through the PLFS VOL and discuss future work leveraging the new semantic post-processing functionality enabled.
+We have written a new plugin for HDF5 using its Virtual Object Layer (VOL) that provides serves two purposes: 1) it uses PLFS to convert the single-file layout into a data layout that is optimized for the underlying file system, and 2) it stores data in a unique way that enables semantic post-processing on the data, independently of HDF5.
+In this paper, we measure the performance improvements achieved using the plugin and discuss future work leveraging the new semantic post-processing functionality enabled.
 We further discuss the applicability of this approach for exascale burst buffer storage systems.

### 2012-09-19 hdf5/discs_2012/background.tex
Added figures, references. Some corrections.

@@
 \subsection{HDF5}
-HDF5 is a data model , library and file format for storing and managing data.
-It defines an information set, which is a container of array variables, groups, and types.
-They are known as datasets, groups and datatype objects in HDF5 respectively.
-The data model defines mechanisms for creating associations between various information items.
-The main components of HDF5 are briefly described below.
+HDF5 is a versatile data model containing complex data objects and metadata, with no limit on the number or size of data objects. 
+Its information set is a collection of datasets, groups, datatype and metadata objects. The main components of HDF5 are briefly described below.
 
@@
 In the HDF5 data model the container of an HDF5 infoset
-is represented by an HDF5 file. It is a collection of HDF5 objects and represents the relationship between them.
+is represented by a file. It is a collection of objects and also explains the relationship between them.
 Every file begins with a root group "/", which serves as the "starting-point" in the object hierarchy. An empty file has atleast the root group in it.
@@
 A dataset is characterized by a dataspace and a datatype.
-An HDF5 dataspace captures the rank (number of dimensions) , and the current and maximum extent in the respective dimensions.
-An HDF5 datatype describes the type of its data elements.
+The dataspace captures the rank (number of dimensions) , and the current and maximum extent in the respective dimensions.
+The datatype describes the type of its data elements.
 HDF5 supports ten classes of datatypes including scalar types like integer, floating-point, string, etc. and compound datatypes.
@@
 Groups:
-HDF5 groups are akin to directories in a file system (reference here).
-A group is an explicit association between HDF5 objects.
-A group could contain multiple other groups, datasets or datatypes /textit{within} it.
-Traversing a root group generates the /information set graph/.
+A group is an explicit association between HDF5 objects. It is synonymous with directories in a file system. A group could contain multiple other groups, datasets or datatypes /textit{within} it. 
+Traversing a root group generates a B-tree representation of all objects in the file.
 
@@
 Attribute A1 is linked to dataset D1.
-The datatype in all datasets in an integer.
+The objects and the relationships between them are represented in a B-tree, which is used internally by HDF5 to index its objects.
 
@@
 HDF5 file format:
-An HDF5 file is a self-describing format which combines data and metadata.
+As discussed previously, an HDF5 file is a self-describing format which combines data and metadata~\cite{hdf5_fileformat}.
 High level objects are stored alongwith metadata that is used to describe various characteristics about the object, alongwith the relationship between various objects.
@@
 Parallel HDF5 (PHDF5):
-As described previously, HDF5 supports parallelism using MPI.
+HDF5 supports parallelism using MPI.
 Multiple processes can be used to create and store objects in a file.
-PHDF5 exports a parallel I/O interface, where internally makes use of MPI-IO to perform I/O in parallel.
+PHDF5 exports a parallel I/O interface, internally making use of MPI-IO interfaces to perform I/O.
 Applications typically store multiple objects in a single HDF5 file.
 However, many popularly used parallel file systems like Lustre, Panasas etc.
-are known to perform poorly for workloads where multiple processes access a shared file (commonly referred to as the N-->1 access pattern).
+are known to perform poorly for workloads where multiple processes access a shared file.
 
 Virtual Object Layer (VOL):
-(I have simply copied this part; need to reframe it) The Virtual Object Layer (VOL) is a new abstraction layer internal to the HDF5 library (reference here).
+The Virtual Object Layer (VOL) is a new abstraction layer internal to the HDF5 library.
 It is implemented just below the public API.
@@
 It is situated between the application and the parallel file system responsible for the actual data storage. 
-It transforms an application's N-->1 access pattern into an N-->N access pattern, where every process participating in I/O writes data to its own , separate file. 
+It transforms an application's N-1 access pattern into an N-N access pattern, where every process participating in I/O writes data to its own, separate file. 
 Thus, it interposes on application I/O and converts its I/O pattern into one that is more suitable for the underlying parallel file system.
@@
 We added a new feature to PLFS called XAttributes (xattrs). Xattrs serve as short metadata stored as key-value pairs.
-All xattrs are stored in a new directory inside the PLFS container structure.
+All xattrs are stored in a separate directory inside the PLFS container structure.
 They can be used to store user-defined information about data for easy and fast retrieval, since they are not stored alongwith raw data.
-We use xattrs to store metadata about HDF5 datasets. The datatype, number of dimensions and their extent etc. are stored as xattrs.
+We use xattrs to store metadata about HDF5 datasets. For example, the datatype, number of dimensions and their extent are stored as xattrs.

### 2012-09-19 hdf5/discs_2012/evaluation.tex
Added figures, references. Some corrections.

@@
-For evaluation purposes, we have used HDF5's \textit{h5perf} performance tool (reference here).
+For evaluation purposes, we have used HDF5's \textit{h5perf} performance tool~\cite{h5perf}.
 h5perf is a tool for testing the performance of PHDF5.
-It allows configuring various parameters, some of which are the number of processes to run the test with, number of datasets to be created in the file, amount of data contributed by each process per dataset, amount of data read/written by a process in a single I/O call (transfer size) etc.
-For our measurements, we create 10 datasets in the file, and the total file size was 64 GB or more.
+It allows configuring various parameters, such as the number of processes to run the test with, number of datasets to be created in the file, amount of data contributed by each process per dataset, amount of data read/written by a process in a single I/O call (transfer size) etc.
+For our measurements, we created 10 datasets in the file, and the total file size was 64 GB or more.
 
-All tests were performed on the Lustre parallel file system on the Atlas cluster at University of Dresden.
-The file system has 12 OSTs configured with a stripe size of 1 MB.
+All tests were performed on the Lustre parallel file system~\cite{lustre} on the Atlas cluster at University of Dresden.
+The file system has 12 OSTs with a stripe size of 1 MB.
 The file system is connected to the compute nodes using an SDR Infiniband link.

### 2012-09-19 hdf5/discs_2012/introduction.tex
Started with evaluation, writing bibliography now, included bib files
Added figures, references. Some corrections.

@@
-Hierarchical Data Format (HDF5) is a data model , library and file format for storing and managing data.
+Hierarchical Data Format (HDF5) is a data model, library and file format for storing and managing data~\cite{hdf5_ad11}.
 It is designed for flexible and efficient I/O.
@@
 The data model defines mechanisms for creating associations between various information items.
-HDF5 is widely used in the industry and scientic domain, in understanding global climate change, special effects in film production, DNA analysis, weather prediction, financial data management etc.
-(citation here)
-Parallel HDF5 (PHDF5) enables developing high performance, parallel applications using standard technologies like MPI in conjunction with HDF5.
+HDF5 is widely used in the industry and scientic domain, in understanding global climate change, special effects in film production, DNA analysis, weather prediction, financial data management etc.~\cite{hdf5_users}
+Parallel HDF5 (PHDF5)~\cite{phdf5} enables developing high performance, parallel applications using standard technologies like MPI in conjunction with HDF5.
 An HDF5 file created using PHDF5 is compatible with serial HDF5 files and is shareable between different platforms.
-PHDF5 exports a standard parallel I/O interface which then uses MPI's parallel I/O functionality.
+PHDF5 exports a standard parallel I/O interface which itself uses MPI's parallel I/O functionality.
 
@@
 However, this native single-file format has its disadvantages.
-A PHDF5 application has multiple processes accessing a single HDF5 file.
-Many popular parallel file systems are known to behave poorly under these circumstances.
+A PHDF5 application has multiple processes accessing a single HDF5 file (called N-1 access pattern).
+Many popular parallel file systems are known to behave poorly under these circumstances~\cite{lustre_perf}\cite{nersc_io}.
 Secondly, since many HDF5 objects are stored in a single file, this eliminates any possibility of performing useful semantic analysis on objects beyond the scope of an HDF5 application.
 In this paper, we address these shortcomings of HDF5, viz.
-performance issues due to multiple processes accessing a single shared file, and lack of a way to perform useful post-processing on HDF5 objects inherent due to the native file format.
+performance issues inherently arising from an N-1 access pattern, and lack of a way to perform useful post-processing on HDF5 objects due to its native file format.
 
-PLFS is a middleware virtual file system developed at Los Alamos National Lab (LANL). It converts writes to a shared logical file into writes to multiple physical files.  
+PLFS is a middleware virtual file system developed at Los Alamos National Lab (LANL)~\cite{plfs_sc09}. It converts writes to a shared logical file into writes to multiple physical files.  
 Thus, it interposes on application I/O and converts its I/O pattern into one that is more suitable for the underlying parallel file system.
@@
 
-In this work, we have developed a new plugin for HDF5 using its recently introduced Virtual Object Layer (VOL).
+In this work, we have developed a new plugin for HDF5 using its recently introduced Virtual Object Layer (VOL)~\cite{vol}.
 The VOL exports an interface that allows writing plugins for HDF5, where it primarily passes references to HDF5 objects, thereby enabling developers to store objects in a format different from the default HDF5 file format.
 We have written a plugin with two main objectives.
-The plugin stores data in a unique way that enables semantic post-processing on HDF5 objects outside the scope of the HDF5 application.
-Secondly, PLFS converts N-->1 accesses into N-->N accesses, thereby showing a significant improvement in performance as well.
-Preliminary results using HDF5's h5perf performance tool show a Xx performance improvement in I/O overall traditional PHDF5.
+The plugin stores data in a unique way that enables semantic post-processing on HDF5 objects.
+Secondly, PLFS converts N-1 accesses into N-N accesses, thereby resulting in improved I/O performance.
+Preliminary results using HDF5's \textit{h5perf} performance tool show a Xx performance improvement in I/O over traditional PHDF5.
 

### 2012-09-19 hdf5/discs_2012/plugin.tex
Started with evaluation, writing bibliography now, included bib files
Added figures, references. Some corrections.

@@
 The plugin stores data in a format which enables performing semantic analysis on the data. 
-It stores every HDF5 object in a separate file/directory, so that data from different objects is not contained in the same file. 
+Instead of storing all objects in a single file, tt stores every HDF5 object in a separate location, so that data from different objects is not contained in the same file. 
 The central object store in HDF5, the HDF5 file, is represented by a directory. Groups are stored as directories as well, whereas datasets, which contain raw data, are stored in files. 
 The relationship between objects is explained by the relative paths between the objects at the file system level.
-As an example, the HDF5 object D1 which in HDF5's B-tree format can be represented as /G1/G2/D1 is stored on the file system as /path-to-hdf5-file/G1/G2/D1.
-This way , the plugin eliminates the need to store the metadata describing the relationship between objects.
+Consider the example shown in figure~\ref{hdf5_example}.
+Using the plugin, file Sample.h5 is stored as a directory.
+Group G1 is stored as a directory under it, and datasets D1 and D2 are files.
+We can see that the relationship between the objects is represented by their relative paths at the file system as shown in figure~\ref{hdf5_example_plugin}.
+This way, the plugin eliminates the need to store the metadata describing the relationship between objects.
 Metadata about datasets, such as the datatype, extent, dimensions etc., are stored as PLFS Xattributes (xattrs) as explained in the previous section.
-
-Consider the example shown figure 1.
-Using the plugin, file test.h5 is stored as a directory.
-Groups G1, G2, G3 are represented as directories at the paths /tmp/test.h5/G1, /tmp/test.h5/G2 , /tmp/test.h5/G1/G3/ respectively.
-Datasets D1 and D2 are files stored under their respective groups.
+\begin{figure}[!t]
+\centering
+\includegraphics[width=2.5in]{hdf5_example_plugin}
+\caption{HDF5 file created using our PLFS plugin}
+\label{hdf5_example_plugin}
+\end{figure}
 
 This approach gives us the ability to distinguish HDF5 objects at the file system level, without requiring an HDF5 application to provide information about objects and the relationship between them. 
-We can now perform post-processing and semantic analysis on the data, outside the scope of the HDF5 application.
+It enables us to perform post-processing and semantic analysis on the data, outside the scope of the HDF5 application.
+
+Implementing the plugin requires providing an implementation for a "driver" interface provided by HDF5.
+In short, we provide an implementation for all functions that access data on disk.
+This includes functions for file management, dataset creation and access, group creation, to name a few.
 

### 2012-09-21 hdf5/discs_2012/abstract.tex
almost complete. yet to change figures

@@
 Additionally, a single-file format does not allow semantic post processing on individual objects outside the scope of the HDF5 application.
-We have written a new plugin for HDF5 using its Virtual Object Layer (VOL) that provides serves two purposes: 1) it uses PLFS to convert the single-file layout into a data layout that is optimized for the underlying file system, and 2) it stores data in a unique way that enables semantic post-processing on the data, independently of HDF5.
-In this paper, we measure the performance improvements achieved using the plugin and discuss future work leveraging the new semantic post-processing functionality enabled.
+We have developed a new plugin for HDF5 using its Virtual Object Layer that provides serves two purposes: 1) it uses PLFS to convert the single-file layout into a data layout that is optimized for the underlying file system, and 2) it stores data in a unique way that enables semantic post-processing on the data.
+We measure the performance of the plugin and discuss work leveraging the new semantic post-processing functionality enabled.
 We further discuss the applicability of this approach for exascale burst buffer storage systems.

### 2012-09-21 hdf5/discs_2012/background.tex
almost complete. yet to change figures

--- 2012-09-17 hdf5/discs_2012/background.tex
@@
 \subsection{HDF5}
-HDF5 is a data model , library and file format for storing and managing data.
-It defines an information set, which is a container of array variables, groups, and types.
-They are known as datasets, groups and datatype objects in HDF5 respectively.
+HDF5 is a versatile data model containing complex data objects and metadata. 
+Its information set is a collection of datasets, groups, datatypes and metadata objects.
 The data model defines mechanisms for creating associations between various information items.
-The main components of HDF5 are briefly described below.
+The main components of HDF5 are described below.
 
@@
 In the HDF5 data model the container of an HDF5 infoset
-is represented by an HDF5 file. It is a collection of HDF5 objects and represents the relationship between them.
-Every file begins with a root group "/", which serves as the "starting-point" in the object hierarchy. An empty file has atleast the root group in it.
+is represented by a file. It is a collection of objects and also explains the relationship between them.
+Every file begins with a root group "/", which serves as the "starting-point" in the object hierarchy.
 
-Datasets:
+Dataset:
 HDF5 datasets are objects that represent actual data or content.
@@
 A dataset is characterized by a dataspace and a datatype.
-An HDF5 dataspace captures the rank (number of dimensions) , and the current and maximum extent in the respective dimensions.
-An HDF5 datatype describes the type of its data elements.
-HDF5 supports ten classes of datatypes including scalar types like integer, floating-point, string, etc. and compound datatypes.
+The dataspace captures the rank (number of dimensions), and the current and maximum extent in each dimension.
+The datatype describes the type of its data elements.
 
-Groups:
-HDF5 groups are akin to directories in a file system (reference here).
-A group is an explicit association between HDF5 objects.
-A group could contain multiple other groups, datasets or datatypes /textit{within} it.
-Traversing a root group generates the /information set graph/.
+Group:
+A group is an explicit association between HDF5 objects. It is synonymous with directories in a file system. A group could contain multiple other groups, datasets or datatypes \textit{within} it.
 
-Attributes:
-Aattributes are used for annotating datasets, groups, and datatype objects. They are datasets themselves, and are /textit{attached} to the existing objects they annotate.
+Attribute:
+Attributes are used for annotating datasets, groups, and datatype objects. They are datasets themselves, and are \textit{attached} to the existing objects they annotate.
 
-The above definitions are explained in the example shown in figure ~\ref{hdf5_example}. 
-The file "sample.h5" contains a root group which itself contains a group G1 and two datasets, D1 and D2.
+For example, as shown in figure~\ref{hdf5_example}, the file "Sample.h5" contains the root group which itself contains a group G1 and two datasets, D1 and D2.
 Group G1 contains a dataset D3.
 Attribute A1 is linked to dataset D1.
-The datatype in all datasets in an integer.
+The objects and the relationships between them can be represented as a B-tree, which is used internally by HDF5 to index its objects.
 
@@
 
-HDF5 file format:
-An HDF5 file is a self-describing format which combines data and metadata.
-High level objects are stored alongwith metadata that is used to describe various characteristics about the object, alongwith the relationship between various objects.
+%Parallel HDF5 (PHDF5):
+%HDF5 supports parallelism using MPI. Multiple processes can be used to access objects in a file. PHDF5 exports a parallel I/O interface, internally making use of MPI-IO. 
+%Applications typically store multiple objects in a single HDF5 file.
+However, many popularly used parallel file systems like Lustre, Panasas etc.
+are known to perform poorly for workloads where multiple processes access a shared file.
 
-Parallel HDF5 (PHDF5):
-As described previously, HDF5 supports parallelism using MPI.
-Multiple processes can be used to create and store objects in a file.
-PHDF5 exports a parallel I/O interface, where internally makes use of MPI-IO to perform I/O in parallel.
-Applications typically store multiple objects in a single HDF5 file.
-However, many popularly used parallel file systems like Lustre, Panasas etc.
-are known to perform poorly for workloads where multiple processes access a shared file (commonly referred to as the N-->1 access pattern).
+Virtual Object Layer (VOL):
+The Virtual Object Layer (VOL) is a new abstraction layer internal to the HDF5 library~\cite{vol}.
+It is implemented just below the public API.
+The VOL exports an interface that allows writing plugins for HDF5, thereby enabling developers to store objects in a format different from the default HDF5 file format (like native netCDF or HDF4 format).
+Plugin writers provide an implementation for a set of functions that access data on disk.
+These include functions for file management, dataset creation and access, group creation, to name a few.
 
@@
 It is situated between the application and the parallel file system responsible for the actual data storage. 
-It transforms an application's N-->1 access pattern into an N-->N access pattern, where every process participating in I/O writes data to its own , separate file. 
-Thus, it interposes on application I/O and converts its I/O pattern into one that is more suitable for the underlying parallel file system.
-The basic operation of PLFs is as follows. For every logical file, PLFS creates a /texit{container} structure on the underlying parallel file system. 
-This container is a hierarchical directory tree consisting of multiple sub-directories.
-Multiple processes opening the same logical file for writing get a unique data file within the container.
-When the processes write to their file, a record identifying the write is appended to an index file shared between all processes.
-Thus, PLFS maintains sufficient metadata to recreate the shared logical file.
-A FUSE daemon exports this container structure as a single, logical file to the end-users, thereby keeping the PLFS operations transparent from them.
+It transforms N-1 into N-N, where every process participating in I/O writes data to its own, separate file. 
+The basic operation of PLFS is as follows. For every writer to a logical file, PLFS creates a unique physical file on the underlying parallel file system. 
+It also maintains sufficient metadata to recreate the shared logical file. 
+We added a new feature to PLFS called extendible attributes (Xattrs). Xattrs serve as short, extensible metadata stored as key-value pairs.
+They can be used to store user-defined information about data for easy and fast retrieval. 
+%We use xattrs to store metadata about HDF5 datasets. For example, the datatype, number of dimensions and their extent are stored as xattrs.
+
+Users can interface with PLFS directly by using the PLFS API or by using an MPI library that has been integrated with PLFS (ad\_plfs layer).

### 2012-09-21 hdf5/discs_2012/conclusion.tex
almost complete. yet to change figures

Using HDF5's virtual object layer, we have developed a new plugin that makes use of PLFS for improved I/O performance. We deviate from HDF5's native file format and store data in a way that allows us to perform useful post-processing independently on HDF5 objects. 
Such a format can further allow us to incorporate semantic analysis of data for burst buffer type workloads. 
Initial results show that our plugin consistently performs better than regular MPI-IO, whereas MPI-IO when used with PLFS's MPI-IO driver shows the best results.

### 2012-09-21 hdf5/discs_2012/evaluation.tex
almost complete. yet to change figures

@@
 For evaluation purposes, we have used HDF5's \textit{h5perf} performance tool~\cite{h5perf}.
-h5perf is a tool for testing the performance of PHDF5.
 It allows configuring various parameters, such as the number of processes to run the test with, number of datasets to be created in the file, amount of data contributed by each process per dataset, amount of data read/written by a process in a single I/O call (transfer size) etc.
@@
 All tests were performed on the Lustre parallel file system~\cite{lustre} on the Atlas cluster at University of Dresden.
-The file system has 12 OSTs with a stripe size of 1 MB.
+The file system has 12 OSTs with a stripe size of 1MB.
 The file system is connected to the compute nodes using an SDR Infiniband link.
-The cluster has 92 nodes with 64 cores each and 64 to 512 GB memory. All nodes are AMD Opteron 6274 running at 2.2 GHz.
+It should be noted that the file system is in production and subject to demanding and varying workload.
+We present here the average bandwidth values of 3 runs, though in certain cases, we have seen significant variation.
+The cluster has 92 AMD Opteron nodes with 64 cores each and 64 to 512 GB memory.
 
-Our tests were configured such that we spawn a maximum of 4 processes per node. Tests run with less than 4 processes utilized a single node.
+We have performed tests with 1,2,4,8,32 and 64 processes. We spawned a maximum of 4 processes per node.
+Reads and writes are either contiguous or interleaved; processes either access contiguous locations in file or execute a strided pattern.
+We compare the performance of the default MPI-IO driver, our plugin, and the PLFS MPI-IO driver (ad\_plfs).
+
+In Figure~\ref{write_contig}, we show the write performance for a transfer size of 1MB for contiguous writes, for varying number of processes.
+It can be seen that the plugin regularly outperforms MPI-IO.
+However, as we increase the number of processes, its performance falls below that of ad\_plfs.
+Figure~\ref{read_contig} shows the performance of contigous reads for an unaligned transfer size.
+The plugin again outperforms MPI-IO, but the performance of ad\_plfs is the best for 64 processes.
+%Figure~\ref{write_interleaved} shows the performance of unaligned writes for a maximum of 8 processes.
+We set the transfer size here to be 10 bytes more than 1MB.
+Also, the writes here are interleaved as opposed to contiguous.
+We see a similar pattern, where the plugin outperforms MPI-IO, but is not as good as that of ad\_plfs for 8 processes.
+Figure~\ref{write_collective} compares the performance of MPI-IO with that of the plugin for collective, unaligned, interleaved writes for a maximum of 8 processes.
+It is important to note that the plugin outperforms MPI-IO in collective mode.
+We have not been able to obtain results for collective ad\_plfs due to resource restrictions.
+
+Overall, our results show that although the plugin generally shows good performance, it does not scale as well as we increase the number of processes.
+This is due to the fact that since we are directly making calls to the PLFS API, the metadata overhead seems to be high.
+We have all processes participating in file open and close operations, which for large problem sizes can add siginificant overhead.
+Hence, for future work, we plan to use ad\_plfs directly in the plugin which can overcome this performance drawback, since ad\_plfs has collective optimizations in open and close.
+It should however be noted that MPI libraries like Volpex-MPI~\cite{volpex}, which do not have MPI-IO as part of their MPI library, could make use of a plugin that does not rely on MPI-IO.
+
+\begin{figure}[!t]
+\centering
+\includegraphics[width=2.5in]{4PEs_node_1M_contig_ind_writes}
+\caption{Performance of contiguous writes}
+\label{write_contig}
+\end{figure}
+
+\begin{figure}[!t]
+\centering
+\includegraphics[width=2.5in]{4pes_node_1m_ind_read}
+\caption{Performance of contiguous reads}
+\label{read_contig}
+\end{figure}
 
@@
 %\centering
-%\includegraphics[width=2.5in]{4pes_interleaved_1m10_coll_writes}
-%\caption{A sample HDF5 file}
-%\label{writes}
+%\includegraphics[width=2.5in]{interleaved_1M10_4pes_writes_noncoll}
+%\caption{Performance of interleaved writes}
+%\label{write_interleaved}
 %\end{figure}
-%
-%\begin{figure}[!t]
-%\centering
-%\includegraphics[width=2.5in]{4pes_interleaved_1m10_coll_reads}
-%\caption{A sample HDF5 file}
-%\label{reads}
-%\end{figure}
+
+\begin{figure}[!t]
+\centering
+\includegraphics[width=2.5in]{4pes_interleaved_1m10_coll_reads}
+\caption{Performance of collective writes}
+\label{write_collective}
+\end{figure}

### 2012-09-21 hdf5/discs_2012/introduction.tex
almost complete. yet to change figures

--- 2012-09-17 hdf5/discs_2012/introduction.tex
@@
-Hierarchical Data Format (HDF5) is a data model , library and file format for storing and managing data.
-It is designed for flexible and efficient I/O.
-HDF5 defines an information set, which is a container of array variables, groups, and types.
-The data model defines mechanisms for creating associations between various information items.
-HDF5 is widely used in the industry and scientic domain, in understanding global climate change, special effects in film production, DNA analysis, weather prediction, financial data management etc.
-(citation here)
-Parallel HDF5 (PHDF5) enables developing high performance, parallel applications using standard technologies like MPI in conjunction with HDF5.
-An HDF5 file created using PHDF5 is compatible with serial HDF5 files and is shareable between different platforms.
-PHDF5 exports a standard parallel I/O interface which then uses MPI's parallel I/O functionality.
+Hierarchical Data Format (HDF5) is a technology suite for efficient management of large and complex data~\cite{hdf5_ad11}.
+It supports complex relationships between data and dependencies between objects.
+HDF5 is widely used in industry and scientic domains, in understanding global climate change, special effects in film production, DNA analysis, weather prediction, financial data management etc.~\cite{hdf5_users}
+Parallel HDF5 (PHDF5)~\cite{phdf5} enables developing high performance, parallel applications using standard technologies like MPI in conjunction with HDF5. 
+%An HDF5 file created using PHDF5 is compatible with serial HDF5 files and is shareable between different platforms. 
+PHDF5 exports a standard parallel I/O interface which itself uses MPI's parallel I/O functionality. This is used alongwith parallel file systems to achieve high performance I/O.
 
-The central component of HDF5, the file, is a self-describing format which combines data and metadata.
-Users typically store multiple HDF5 objects in a single file, and the library stores metadata alongwith data that describes the relationships between various objects amongst other metadata.
+An HDF5 file is a self-describing format which combines data and metadata. It is a container in which users typically store multiple HDF5 objects alongside their metadata. 
 However, this native single-file format has its disadvantages.
-A PHDF5 application has multiple processes accessing a single HDF5 file.
-Many popular parallel file systems are known to behave poorly under these circumstances.
-Secondly, since many HDF5 objects are stored in a single file, this eliminates any possibility of performing useful semantic analysis on objects beyond the scope of an HDF5 application.
-In this paper, we address these shortcomings of HDF5, viz.
-performance issues due to multiple processes accessing a single shared file, and lack of a way to perform useful post-processing on HDF5 objects inherent due to the native file format.
+A PHDF5 application has multiple processes accessing a single file (i.e.
+N-1 access pattern).
+Unfortunately, many popular parallel file systems are known to behave poorly under these circumstances~\cite{lustre_perf}\cite{nersc_io}.
+Secondly, since many HDF5 objects are stored in a single file, this complicates the possibility of performing useful semantic operations on individual objects.
+In this paper, we enhance the performance of HDF5 by addressing the performance issues inherently arising from poor file system performance on N-1 access patterns, and augment performing  useful post-processing on HDF5 objects.
 
-PLFS is a middleware virtual file system developed at Los Alamos National Lab (LANL). It converts writes to a shared logical file into writes to multiple physical files.  
-Thus, it interposes on application I/O and converts its I/O pattern into one that is more suitable for the underlying parallel file system.
-PLFS is popularly used as a checkpoint file system where applications demonstrating heavy checkpoint workloads are known to benefit significantly from PLFS.
+PLFS (Parallel Log-Structured File System) is a middleware virtual file system developed at Los Alamos National Lab (LANL)~\cite{plfs_sc09}.
+It converts writes to a shared logical file into writes to multiple physical files to overcome the performance bottleneck associated with N-1 writes.
+However, it suffers from an inability to understand the structure of the data that it stores.
 
 In this work, we have developed a new plugin for HDF5 using its recently introduced Virtual Object Layer (VOL).  
-The VOL exports an interface that allows writing plugins for HDF5, where it primarily passes references to HDF5 objects, thereby enabling developers to store objects in a format different from the default HDF5 file format.
-We have written a plugin with two main objectives.
-The plugin stores data in a unique way that enables semantic post-processing on HDF5 objects outside the scope of the HDF5 application.
-Secondly, PLFS converts N-->1 accesses into N-->N accesses, thereby showing a significant improvement in performance as well.
-Preliminary results using HDF5's h5perf performance tool show a Xx performance improvement in I/O overall traditional PHDF5.
+This plugin serves two purposes: it stores data in a unique way that enables semantic post-processing on HDF5 objects, and by using PLFS to convert N-1 accesses into N-N accesses, thereby resulting in improved I/O performance.
 
-The rest of the paper is organized as follows.
-We represent details about HDF5 and PLFS in section 2, describe our plugin design and implementation in section 3, and our evaluation in section 4.
-We present related work in section 5, the current status of the plugin alongwith future work in section 6, and finally we conclude in section 7.
+%The rest of the paper is organized as follows.
+We present some background on HDF5 and PLFS in section 2, our plugin design and implementation in section 3, our evaluation in section 4, and our conclusion in section 5.

### 2012-09-21 hdf5/discs_2012/plugin.tex
almost complete. yet to change figures

--- 2012-09-17 hdf5/discs_2012/plugin.tex
@@
-The plugin stores data in a format which enables performing semantic analysis on the data. 
-It stores every HDF5 object in a separate file/directory, so that data from different objects is not contained in the same file. 
-The central object store in HDF5, the HDF5 file, is represented by a directory. Groups are stored as directories as well, whereas datasets, which contain raw data, are stored in files. 
-The relationship between objects is explained by the relative paths between the objects at the file system level.
-As an example, the HDF5 object D1 which in HDF5's B-tree format can be represented as /G1/G2/D1 is stored on the file system as /path-to-hdf5-file/G1/G2/D1.
-This way , the plugin eliminates the need to store the metadata describing the relationship between objects.
-Metadata about datasets, such as the datatype, extent, dimensions etc., are stored as PLFS Xattributes (xattrs) as explained in the previous section.
+We have developed a plugin which stores data in a format which enables performing semantic analysis on the data and uses PLFS for high I/O throughput. 
+Instead of storing all objects in a single file, it stores every HDF5 object in a separate location, so that data from different objects is not contained in the same file.
+In short, we provide a raw mapping of the HDF5 objects to the file system.
+HDF5 files and groups are stored as directories, whereas datasets and attributes, which contain raw data, are stored in files created using PLFS. 
+Consider the example shown in figure~\ref{hdf5_example}.
+Using the plugin, file Sample.h5 is stored as a directory.
+Group G1 is stored as a directory under it, and datasets D1 and D2 are files.
+Attribute A1 is a file created at the same location as dataset D1 to which it is attached.
+Additionally, the name of the attribute is stored as \textit{dataset name}.\textit{attribute name} to denote the dataset to which the attribute is attached.
+Thus these objects are stored at the following paths:\\
+/Sample.h5/\\
+/Sample.h5/G1\\
+/Sample.h5/G1/D3\\
+/Sample.h5/D1\\
+/Sample.h5/D1.A1\\
+/Sample.h5/D2\\
 
-Consider the example shown figure 1.
-Using the plugin, file test.h5 is stored as a directory.
-Groups G1, G2, G3 are represented as directories at the paths /tmp/test.h5/G1, /tmp/test.h5/G2 , /tmp/test.h5/G1/G3/ respectively.
-Datasets D1 and D2 are files stored under their respective groups.
+We can see that the relationship between the objects is represented by their relative paths at the file system.
+That is, the path /Sample.h5/G1/D3 tells us that D3 is a dataset belonging to Group G1 under the root group of the file.
+This way, the plugin also eliminates the need to store the metadata describing the relationship between objects.
+Metadata about datasets, such as the datatype, extent, dimensions etc.
+are stored as Xattrs.
+%\begin{figure}[!t]
+%\centering
+%\includegraphics[width=2.5in]{hdf5_example_plugin}
+%\caption{HDF5 file created using our PLFS plugin}
+%\label{hdf5_example_plugin}
+%\end{figure}
 
-This approach gives us the ability to distinguish HDF5 objects at the file system level, without requiring an HDF5 application to provide information about objects and the relationship between them. 
-We can now perform post-processing and semantic analysis on the data, outside the scope of the HDF5 application.
+This approach gives us the ability to distinguish HDF5 objects inside the storage system.
+
+Our plugin makes direct calls to the PLFS API.
+For the current version of our plugin, we do not support collective I/O operations.
+It should be noted that HDF5 when used natively with MPI-IO allows users to specify whether collective I/O should be used for reading and writing data.
+
+The above format of storing data allows us to perform atleast two different analysis optimizations.
+To illustrate via an example, imagine storing a three-dimensional ocean model within a PLFS file.
+The storage system sees the file as an opaque linear array of bytes.
+With the structure, however, PLFS can provide \textit{active analysis} as well as \textit{semantic restructuring}.
+
+Active analysis borrows the transducers idea from the Semantic File System~\cite{semantic_fs} which has since been productized in BigTable and Hbase~\cite{google_coprocessors,GFS,apache_hbase}.
+With active analysis, the application can ship a data parser function when it creates the PLFS file.
+As the data is written into PLFS, PLFS can apply the data function on the streaming data.
+The function will output key-value pairs which PLFS can embed in its extensible metadata.
+In this example, one simple function might record the height of the largest wave.
+Due to PLFS's model of storing a logical file across multiple physical files, the PLFS extensible metadata can record the height of the largest wave within each physical file.
+However, given that PLFS now understands the structure of the logical file, these multiple physical files, within the PLFS container, are more accurately thought of as shards.
+In a future burst buffer architecture~\cite{burst_buffers}, these semantic shards will be spread across multiple burst buffer nodes.
+Therefore, subsequent analysis of the ocean model can quickly find the burst buffer containing the shard with the largest wave by searching a small amount of extensible metadata instead of scanning the entire ocean model.
+
+Semantic restructuring is the idea of reorganizing the data into a new set of semantic shards.
+This would be done to speed future analysis routines.
+For example, assume that the ocean model was originally sharded using a row-order organization (i.e.
+across latitude instead of longitude).
+An analysis routine which will explore the model along a column-ordering will suffer poor performance with the row-order organization as its access pattern will result in a large number of small reads from a large set of semantic shards.
+However, by knowing the semantic structure, the analysis routine can request a semantic restructuring which will be a compact, intuitively described request such as "restructure into row-ordering."  Without structural knowledge, a semantic restructuring would be significantly more complicated: the analysis routine would have to send a large list of logical offsets to PLFS to inform it of expected read patterns.
+In an exascale system, the list of logical offsets will be in the order of one billion.
+Semantic restructuring shrinks the size of the request to a small constant value.
 

### 2012-09-21 hdf5/discs_2012/evaluation.tex
one change to evaluation. 2) added gary as co-author

@@
 Hence, for future work, we plan to use ad\_plfs directly in the plugin which can overcome this performance drawback, since ad\_plfs has collective optimizations in open and close.
-It should however be noted that MPI libraries like Volpex-MPI~\cite{volpex}, which do not have MPI-IO as part of their MPI library, could make use of a plugin that does not rely on MPI-IO.
+It should however be noted that there are some MPI libraries tailored to suit a specific purpose and which do not provide an implementation for MPI-IO.
+Such libraries can benefit from using a plugin that does not rely on MPI-IO.
 

### 2012-09-21 hdf5/discs_2012/evaluation.tex
A couple of typos found by running aspell on the tex files

@@
 However, as we increase the number of processes, its performance falls below that of ad\_plfs.
-Figure~\ref{read_contig} shows the performance of contigous reads for an unaligned transfer size.
+Figure~\ref{read_contig} shows the performance of contiguous reads for an unaligned transfer size.
 The plugin again outperforms MPI-IO, but the performance of ad\_plfs is the best for 64 processes.
@@
 This is due to the fact that since we are directly making calls to the PLFS API, the metadata overhead seems to be high.
-We have all processes participating in file open and close operations, which for large problem sizes can add siginificant overhead.
+We have all processes participating in file open and close operations, which for large problem sizes can add significant overhead.
 Hence, for future work, we plan to use ad\_plfs directly in the plugin which can overcome this performance drawback, since ad\_plfs has collective optimizations in open and close.

### 2012-09-24 hdf5/discs_2012/abstract.tex
Final submitted version on 9/22. Evaluation contains plots from old 64 filesize test case. Minor typos, bug fixes everywhere. Removed old figures and added new ones.

@@
 Additionally, a single-file format does not allow semantic post processing on individual objects outside the scope of the HDF5 application.
-We have developed a new plugin for HDF5 using its Virtual Object Layer that provides serves two purposes: 1) it uses PLFS to convert the single-file layout into a data layout that is optimized for the underlying file system, and 2) it stores data in a unique way that enables semantic post-processing on the data.
+We have developed a new plugin for HDF5 using its Virtual Object Layer that serves two purposes: 1) it uses PLFS to convert the single-file layout into a data layout that is optimized for the underlying file system, and 2) it stores data in a unique way that enables semantic post-processing on data.
 We measure the performance of the plugin and discuss work leveraging the new semantic post-processing functionality enabled.

### 2012-09-24 hdf5/discs_2012/background.tex
Final submitted version on 9/22. Evaluation contains plots from old 64 filesize test case. Minor typos, bug fixes everywhere. Removed old figures and added new ones.

@@
-\subsection{HDF5}
+%\subsection{HDF5}
 HDF5 is a versatile data model containing complex data objects and metadata. 
@@
 In the HDF5 data model the container of an HDF5 infoset
-is represented by a file. It is a collection of objects and also explains the relationship between them.
+is represented by a file. It is a collection of objects that also explains the relationship between them.
 Every file begins with a root group "/", which serves as the "starting-point" in the object hierarchy.
@@
 Attribute:
-Attributes are used for annotating datasets, groups, and datatype objects. They are datasets themselves, and are \textit{attached} to the existing objects they annotate.
+Attributes are used for annotating datasets, groups, and datatype objects. They are datasets themselves, and are \textit{attached} to existing objects they annotate.
 
@@
 Virtual Object Layer (VOL):
-The Virtual Object Layer (VOL) is a new abstraction layer internal to the HDF5 library~\cite{vol}.
+VOL is a new abstraction layer internal to the HDF5 library~\cite{vol}.
 It is implemented just below the public API.
@@
 
-\subsection{PLFS}
+%\subsection{PLFS}
 PLFS is a middleware virtual file system that converts writes to a shared logical file into writes to multiple physical files. 
@@
 
-Users can interface with PLFS directly by using the PLFS API or by using an MPI library that has been integrated with PLFS (ad\_plfs layer).
+Users can interface with PLFS directly by using the PLFS API or by using its MPI-IO driver (ad\_plfs).

### 2012-09-24 hdf5/discs_2012/conclusion.tex
Final submitted version on 9/22. Evaluation contains plots from old 64 filesize test case. Minor typos, bug fixes everywhere. Removed old figures and added new ones.

@@
 Such a format can further allow us to incorporate semantic analysis of data for burst buffer type workloads. 
-Initial results show that our plugin consistently performs better than regular MPI-IO, whereas MPI-IO when used with PLFS's MPI-IO driver shows the best results.
+Initial results show that our plugin outperforms MPI-IO in most cases, whereas MPI-IO when used with PLFS's MPI-IO driver shows the best results.

### 2012-09-24 hdf5/discs_2012/evaluation.tex
Final submitted version on 9/22. Evaluation contains plots from old 64 filesize test case. Minor typos, bug fixes everywhere. Removed old figures and added new ones.

@@
 For evaluation purposes, we have used HDF5's \textit{h5perf} performance tool~\cite{h5perf}.
-It allows configuring various parameters, such as the number of processes to run the test with, number of datasets to be created in the file, amount of data contributed by each process per dataset, amount of data read/written by a process in a single I/O call (transfer size) etc.
-For our measurements, we created 10 datasets in the file, and the total file size was 64 GB or more.
+It allows configuring various parameters, such as the number of processes, number of datasets, amount of data read/written by a process in a single I/O call (transfer size) etc.
+For our measurements, we created 10 datasets, and the total file size was 64 GB or more.
+In every run, every process contributed equal amount of data per dataset using the default individual (non-collective) I/O mode of h5perf.
 
-All tests were performed on the Lustre parallel file system~\cite{lustre} on the Atlas cluster at University of Dresden.
+Tests were performed on the Lustre parallel file system~\cite{lustre} on the Atlas cluster at University of Dresden.
 The file system has 12 OSTs with a stripe size of 1MB.
 The file system is connected to the compute nodes using an SDR Infiniband link.
-It should be noted that the file system is in production and subject to demanding and varying workload.
-We present here the average bandwidth values of 3 runs, though in certain cases, we have seen significant variation.
 The cluster has 92 AMD Opteron nodes with 64 cores each and 64 to 512 GB memory.
+Tests were run thrice and we present the average bandwidth values, which does not include the time taken to open and close the file.
 
-We have performed tests with 1,2,4,8,32 and 64 processes. We spawned a maximum of 4 processes per node.
+We have performed tests with 1,2,4,8,32 and 64 processes with a maximum of 4 processes per node.
 Reads and writes are either contiguous or interleaved; processes either access contiguous locations in file or execute a strided pattern.
@@
 
-In Figure~\ref{write_contig}, we show the write performance for a transfer size of 1MB for contiguous writes, for varying number of processes.
-It can be seen that the plugin regularly outperforms MPI-IO.
-However, as we increase the number of processes, its performance falls below that of ad\_plfs.
-Figure~\ref{read_contig} shows the performance of contiguous reads for an unaligned transfer size.
-The plugin again outperforms MPI-IO, but the performance of ad\_plfs is the best for 64 processes.
+In Figure~\ref{write_contig}, we show the write performance for a transfer size of 1MB for contiguous writes.
+It can be seen that the plugin regularly outperforms MPI-IO except for the 64 process case, where the metadata overhead incurred by the plugin is high.
+The performance  of ad\_plfs is the best for higher process counts.
+Figure~\ref{read_contig} shows the performance of contiguous reads.
 %Figure~\ref{write_interleaved} shows the performance of unaligned writes for a maximum of 8 processes.
@@
 We see a similar pattern, where the plugin outperforms MPI-IO, but is not as good as that of ad\_plfs for 8 processes.
-Figure~\ref{write_collective} compares the performance of MPI-IO with that of the plugin for collective, unaligned, interleaved writes for a maximum of 8 processes.
-It is important to note that the plugin outperforms MPI-IO in collective mode.
-We have not been able to obtain results for collective ad\_plfs due to resource restrictions.
+Figures~\ref{write_interleaved} and~\ref{read_interleaved} show the interleaved write and read performance respectively for an unaligned transfer size (1M + 10bytes) for a maximum of 8 processes.
+The write performance of MPI-IO is quite poor in this case.
+The plugin easily outperforms MPI-IO and almost matches the performance of ad\_plfs.
 
-Overall, our results show that although the plugin generally shows good performance, it does not scale as well as we increase the number of processes.
-This is due to the fact that since we are directly making calls to the PLFS API, the metadata overhead seems to be high.
-We have all processes participating in file open and close operations, which for large problem sizes can add significant overhead.
+Overall, results show that the plugin consistently shows good performance, however it does not scale as well as ad\_plfs.
+This is due to the fact that since we are directly making calls to the PLFS API, the metadata overhead is high.
+All processes participate in file open and close operations, which for large problem sizes adds significant overhead.
 Hence, for future work, we plan to use ad\_plfs directly in the plugin which can overcome this performance drawback, since ad\_plfs has collective optimizations in open and close.
-It should however be noted that there are some MPI libraries tailored to suit a specific purpose and which do not provide an implementation for MPI-IO.
+It should be noted that there are some MPI libraries tailored to suit specific types of applications and which do not provide an implementation for MPI-IO.
 Such libraries can benefit from using a plugin that does not rely on MPI-IO.
@@
 \centering
-\includegraphics[width=2.5in]{4PEs_node_1M_contig_ind_writes}
+\includegraphics[width=3.6in,height=3.0in]{contig_w}
 \caption{Performance of contiguous writes}
@@
 \centering
-\includegraphics[width=2.5in]{4pes_node_1m_ind_read}
+\includegraphics[width=3.6in,height=3.0in]{contig_r}
 \caption{Performance of contiguous reads}
@@
 
-%\begin{figure}[!t]
-%\centering
-%\includegraphics[width=2.5in]{interleaved_1M10_4pes_writes_noncoll}
-%\caption{Performance of interleaved writes}
-%\label{write_interleaved}
-%\end{figure}
+\begin{figure}[!t]
+\centering
+\includegraphics[width=3.6in,height=3.0in]{interleaved_w}
+\caption{Performance of interleaved, unaligned writes}
+\label{write_interleaved}
+\end{figure}
 
@@
 \centering
-\includegraphics[width=2.5in]{4pes_interleaved_1m10_coll_reads}
-\caption{Performance of collective writes}
-\label{write_collective}
+\includegraphics[width=3.6in,height=3.0in]{interleaved_r}
+\caption{Performance of interleaved,unaligned reads}
+\label{read_interleaved}
 \end{figure}

### 2012-09-24 hdf5/discs_2012/introduction.tex
Final submitted version on 9/22. Evaluation contains plots from old 64 filesize test case. Minor typos, bug fixes everywhere. Removed old figures and added new ones.

@@
 It supports complex relationships between data and dependencies between objects.
-HDF5 is widely used in industry and scientic domains, in understanding global climate change, special effects in film production, DNA analysis, weather prediction, financial data management etc.~\cite{hdf5_users}
+HDF5 is widely used in industry and scientific domains, in understanding global climate change, special effects in film production, DNA analysis, weather prediction, financial data management etc.~\cite{hdf5_users}
 Parallel HDF5 (PHDF5)~\cite{phdf5} enables developing high performance, parallel applications using standard technologies like MPI in conjunction with HDF5. 
 %An HDF5 file created using PHDF5 is compatible with serial HDF5 files and is shareable between different platforms. 
-PHDF5 exports a standard parallel I/O interface which itself uses MPI's parallel I/O functionality. This is used alongwith parallel file systems to achieve high performance I/O.
+PHDF5 exports a standard parallel I/O interface which itself uses MPI's parallel I/O functionality. This is used along with parallel file systems to achieve high performance I/O.
 
@@
 N-1 access pattern).
-Unfortunately, many popular parallel file systems are known to behave poorly under these circumstances~\cite{lustre_perf}\cite{nersc_io}.
-Secondly, since many HDF5 objects are stored in a single file, this complicates the possibility of performing useful semantic operations on individual objects.
-In this paper, we enhance the performance of HDF5 by addressing the performance issues inherently arising from poor file system performance on N-1 access patterns, and augment performing  useful post-processing on HDF5 objects.
+Unfortunately, many popular parallel file systems are known to behave poorly under these circumstances~\cite{lustre_perf,nersc_io}.
+Secondly, since many HDF5 objects are stored in a single file, this complicates performing useful semantic operations on individual objects.
 
@@
 
-In this work, we have developed a new plugin for HDF5 using its recently introduced Virtual Object Layer (VOL).  
-This plugin serves two purposes: it stores data in a unique way that enables semantic post-processing on HDF5 objects, and by using PLFS to convert N-1 accesses into N-N accesses, thereby resulting in improved I/O performance.
+In this paper, we enhance the performance of HDF5 by addressing the performance issues inherently arising from poor file system performance on N-1 access patterns, and augment performing  useful post-processing on HDF5 objects.
+We have developed a new plugin for HDF5 using its recently introduced Virtual Object Layer (VOL).  
+This plugin stores data in a unique way that enables semantic post-processing on HDF5 objects, and uses PLFS to convert N-1 accesses into N-N accesses, thereby resulting in improved I/O performance.
 

### 2012-09-24 hdf5/discs_2012/plugin.tex
A couple of typos found by running aspell on the tex files
Final submitted version on 9/22. Evaluation contains plots from old 64 filesize test case. Minor typos, bug fixes everywhere. Removed old figures and added new ones.

@@
 Instead of storing all objects in a single file, it stores every HDF5 object in a separate location, so that data from different objects is not contained in the same file.
-In short, we provide a raw mapping of the HDF5 objects to the file system.
+In short, we provide a raw mapping of HDF5 objects to the file system.
 HDF5 files and groups are stored as directories, whereas datasets and attributes, which contain raw data, are stored in files created using PLFS. 
@@
 That is, the path /Sample.h5/G1/D3 tells us that D3 is a dataset belonging to Group G1 under the root group of the file.
-This way, the plugin also eliminates the need to store the metadata describing the relationship between objects.
+This approach gives us the ability to distinguish HDF5 objects inside the storage system.
+Also, the plugin eliminates the need to explicitly store the metadata describing the relationship between objects.
 Metadata about datasets, such as the datatype, extent, dimensions etc.
 are stored as Xattrs.
+
 %\begin{figure}[!t]
@@
 
-This approach gives us the ability to distinguish HDF5 objects inside the storage system.
-
 Our plugin makes direct calls to the PLFS API.
-For the current version of our plugin, we do not support collective I/O operations.
+For the current version, we do not support collective I/O operations.
 It should be noted that HDF5 when used natively with MPI-IO allows users to specify whether collective I/O should be used for reading and writing data.
 
-The above format of storing data allows us to perform atleast two different analysis optimizations.
+The above format of storing data allows us to perform at least two different analysis optimizations.
 To illustrate via an example, imagine storing a three-dimensional ocean model within a PLFS file.
@@
 
-Active analysis borrows the transducers idea from the Semantic File System~\cite{semantic_fs} which has since been productized in BigTable and Hbase~\cite{google_coprocessors,GFS,apache_hbase}.
+Active analysis borrows the transducers idea from the Semantic File System~\cite{semantic_fs} which has since been productized in Google's BigTable and Apache Hbase technologies~\cite{google_coprocessors,GFS,apache_hbase}.
 With active analysis, the application can ship a data parser function when it creates the PLFS file.

### 2012-12-11 hdf5/discs_2012/background.tex
For camera ready version

@@
 VOL is a new abstraction layer internal to the HDF5 library~\cite{vol}.
-It is implemented just below the public API.
+As shown in figure~\ref{hdf5_vol} it is implemented just below the public API.
 The VOL exports an interface that allows writing plugins for HDF5, thereby enabling developers to store objects in a format different from the default HDF5 file format (like native netCDF or HDF4 format).
@@
 These include functions for file management, dataset creation and access, group creation, to name a few.
+
+\begin{figure}[!t]
+\centering
+\includegraphics[width=3.5in,height=2.8in]{hdf5_vol}
+\caption{Virtual Object Layer (VOL)}
+\label{hdf5_vol}
+\end{figure}
 
@@
 Users can interface with PLFS directly by using the PLFS API or by using its MPI-IO driver (ad\_plfs).
+
+As stated before, when multiple processes access a single, shared file, file systems like Lustre are known to exhibit sub-standard I/O performance.
+PLFS has demonstrated benefits of converting this N-1 access pattern into one where every process writes data to its own physical file and sufficient metadata is maintained to recreate the original file in the correct order of writes issued.
+Secondly, the data model of HDF5 stores multiple datasets in a single file on disk.
+As a result, one cannot perform any useful post-processing on individual HDF5 objects.
+As we shall see, storing objects in a different file format allows us to perform optimizations on individual HDF5 objects without affecting the remaining data in the file.

### 2012-12-11 hdf5/discs_2012/evaluation.tex
For camera ready version

@@
+In this section, We present the the I/O performance of our plugin.
+These experiments are not targeted towards performing a semantic analysis of data, such as active analysis or semantic restructuring discussed in the previous section.
 For evaluation purposes, we have used HDF5's \textit{h5perf} performance tool~\cite{h5perf}.
 It allows configuring various parameters, such as the number of processes, number of datasets, amount of data read/written by a process in a single I/O call (transfer size) etc.
-For our measurements, we created 10 datasets, and the total file size was 64 GB or more.
+For our measurements, we created 10 1-dimensional datasets, and the total file size was 64 GB or more.
 In every run, every process contributed equal amount of data per dataset using the default individual (non-collective) I/O mode of h5perf.

### 2012-12-11 hdf5/discs_2012/introduction.tex
For camera ready version

@@
 An HDF5 file is a self-describing format which combines data and metadata. It is a container in which users typically store multiple HDF5 objects alongside their metadata. 
-However, this native single-file format has its disadvantages.
 A PHDF5 application has multiple processes accessing a single file (i.e.
@@
 Unfortunately, many popular parallel file systems are known to behave poorly under these circumstances~\cite{lustre_perf,nersc_io}.
-Secondly, since many HDF5 objects are stored in a single file, this complicates performing useful semantic operations on individual objects.
+Secondly, HDF5 stores multiple data objects in a single file.
+This complicates performing useful semantic operations on individual objects.
 

### 2012-12-11 hdf5/discs_2012/plugin.tex
For camera ready version

@@
 Additionally, the name of the attribute is stored as \textit{dataset name}.\textit{attribute name} to denote the dataset to which the attribute is attached.
-Thus these objects are stored at the following paths:\\
+Thus these objects are stored at the following paths as shown in figure ~\ref{hdf5_example_plugin}:\\
 /Sample.h5/\\
@@
 
-%\begin{figure}[!t]
-%\centering
-%\includegraphics[width=2.5in]{hdf5_example_plugin}
-%\caption{HDF5 file created using our PLFS plugin}
-%\label{hdf5_example_plugin}
-%\end{figure}
+\begin{figure}[!t]
+\centering
+\includegraphics[width=2.5in]{hdf5_example_plugin}
+\caption{HDF5 data stored using the plugin}
+\label{hdf5_example_plugin}
+\end{figure}
 
@@
 As the data is written into PLFS, PLFS can apply the data function on the streaming data.
-The function will output key-value pairs which PLFS can embed in its extensible metadata.
+The function will output key-value pairs which PLFS can embed in its extensible metadata (figure~\ref{active_analysis}).
 In this example, one simple function might record the height of the largest wave.
@@
 
+\begin{figure}[!t]
+\centering
+\includegraphics[width=3.5in,height=3.0in]{active_analysis}
+\caption{Active Analysis of Data}
+\label{active_analysis}
+\end{figure}
+
 Semantic restructuring is the idea of reorganizing the data into a new set of semantic shards.
@@
 An analysis routine which will explore the model along a column-ordering will suffer poor performance with the row-order organization as its access pattern will result in a large number of small reads from a large set of semantic shards.
-However, by knowing the semantic structure, the analysis routine can request a semantic restructuring which will be a compact, intuitively described request such as "restructure into row-ordering."  Without structural knowledge, a semantic restructuring would be significantly more complicated: the analysis routine would have to send a large list of logical offsets to PLFS to inform it of expected read patterns.
+However, by knowing the semantic structure, the analysis routine can request a semantic restructuring (figure~\ref{semantic_restructuring}) which will be a compact, intuitively described request such as "restructure into column-ordering."  Without structural knowledge, a semantic restructuring would be significantly more complicated: the analysis routine would have to send a large list of logical offsets to PLFS to inform it of expected read patterns.
 In an exascale system, the list of logical offsets will be in the order of one billion.
 Semantic restructuring shrinks the size of the request to a small constant value.
+A storage format where datasets are stored in separate files allows us to make modifications to the storage layout of one dataset without affecting other datasets and data.
+
+\begin{figure}[!t]
+\centering
+\includegraphics[width=3.5in,height=3.0in]{semantic_restructuring}
+\caption{Semantic Restructuring of Data}
+\label{semantic_restructuring}
+\end{figure}
 
